---
title: 'OHI `r format(Sys.time(), "%Y")`: Natural Products: weighting data prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

[REFERENCE RMD FILE:
<http://ohi-science.org/ohiprep_v2021/globalprep/np/v2021/STEP2_np_weighting_prep.html>]

# Summary

This analysis converts FAO commodities tonnes data for ornamental fish
trade, SAUP fisheries tonnes data for FOFM, and FAO mariculture tonnes
data for seaweeds into a weighting scheme per each ohi region, based off
of 5 year averages of tonnes and values (in usd) for each product. This
weighting scheme will be applied to the scores to determine how much of
each natural products layer affects the overall NP score when updating
in ohi global.


## Updates from previous year

-   updated syntax to follow tidyverse style when reasonable; updated functions `read.csv()` and `write.csv()` to `read_csv()` and `write_csv()`

-   updated file paths to be more reproducible, use of `here()` package and updated syntax to reduce system-dependent file paths

------------------------------------------------------------------------

# Data Source

## FAO: Commodities

**Reference**:\
<https://www.fao.org/fishery/en/statistics/software/fishstatj> App release date: March 2024 

FAO raw commodities quantity 1976_2021 FAO raw commodities value 1976_2022 [FAO metadata](http://www.fao.org/fishery/statistics/global-commodities-production/en)

- Global aquatic trade - All partners aggregated - Quantities and Values - 1976-2022 (Release date: July 2024)


**Downloaded**: July, 26, 2024

**Description**: Quantity (tonnes) and value (USD) of raw commodities (Exports only) for each country, taxa, year. The FAO data is subset to include commodities in these categories: ornamental fish, fish oil, seaweed and plants (see: raw/commodities2products_weighting.csv for details).

**Time range**: 1976-2022

## Seafood Watch sustainability data

**Reference**: <https://www.seafoodwatch.org/globalassets/sfw/pdf/whats-new/seafood-watch-complete-recommendation-list.pdf> Release date: March 4, 2023

**Downloaded**: May 31, 2023

**Description**: Monterrey Bay Aquarium Seafood Watch aquaculture recommendations. Sustainability scored from 0-10. Rescaled to 0-1. There is only one value for seaweeds in the data... 0.67


## Sea Around Us Concepts

**Reference**: Pauly D., Zeller D., Palomares M.L.D. (Editors), 2020.
Sea Around Us Concepts, Design and Data (seaaroundus.org).

**Downloaded**: September 27, 2022

**Description**: Tons per year and SAUP region with information on
sector type, industry type, fishing entitity, reporting status and
taxonomic information.

**Time range**: 1950 - 2019

**Format**: CSV

**Additional Information**:
[Methods](http://www.seaaroundus.org/sea-around-us-methods-index/)

## Avoiding the ecological limits of forage fish for fed aquaculture

**Reference**: Froehlich, H.E., Jacobsen, N.S., Essington, T.E.,
Clavelle, T., and Halpern, B.S. (2018). Avoiding the ecological limits
of forage fish for fed aquaculture. Nature Sustainability 1, 298.

**Downloaded**: July 7, 2020. Obtained from Melanie Frazier (NCEAS).

**Description**: List of FOFM species from Watson v3 data.

**Native data resolution**:

**Format**: CSV format

## RAM

## RAM Legacy Data

**Reference**: [RAM Legacy Stock Assessment Database](http://ramlegacy.org) v4.65, 06/17/2024

**Downloaded**: 08/07/2024

**Description**: B/Bmsy value by stock and year (other data, which we do not use, are also available in the database)

**Native data resolution**: stock (fish stock, species and region specific)

**Time range**: 1800 - 2023

**Format**: CSV format

**Additional Information**: We use the finalized b/bmsy layer from OHI-global for this data prep. We do not actually read in the raw RAM data here.

## FAO: Production

**Reference**:\
<https://www.fao.org/fishery/statistics-query/en/aquaculture/aquaculture_quantity> FAO Global Aquaculture Production Quantity 1950_2022 FAO metadata found [here](http://www.fao.org/fishery/statistics/global-aquaculture-production/en)

**Downloaded**: July 26th, 2024

**Last updated**: March 29th, 2024

**Description**: Quantity (tonnes) of mariculture for each country, species, year.

**Time range**: 1950-2022

------------------------------------------------------------------------

## Setup

```{r setup, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(eval=FALSE)

# Read in packages 
library(here)
library(tidyverse)

# Set scenario year, reproducible file paths
scen_year_number <- 2024 # update this!!
scenario_year <- as.character(scen_year_number)
v_scen_year <- paste0("v", scenario_year)

current_np_dir <- here::here("globalprep", "np", v_scen_year)
previous_np_dir <- here::here("globalprep", "np", paste0("v", scen_year_number - 1))

# Source scripts/functions
source(here::here("workflow", "R", "common.R")) # directory locations
source(here::here(current_np_dir, "R", "np_fxn.R")) # function for handling FAO commodity data specific to NP

# Archived setup:
#prep_year <- 2023
#prep <- file.path(here(paste0("globalprep/np/v", prep_year, "/")))
```

## Methods

Import prepped FAO commodities data set and find 5 year average \$USD /
Tonne average value per product per region. Gapfill these based on the
UN geopolitical region/global values.

```{r}
# Read in the prepped FAO commodities data set 
harvest_tonnes_usd <- read_csv(here(current_np_dir, "int", "np_harvest_tonnes_usd.csv")) %>%
  filter(rgn_id != 213)

max_year <- max(harvest_tonnes_usd$year)
min_year <- max_year - 4 
years <- c(min_year:max_year)
years_df <- data.frame(year = min_year:max_year) # so that we can get 5 year average

# Load OHI region data
region_data()

# Define subset of OHI EEZ regions
rgns_eez <- rgns_eez %>%
  dplyr::select(rgn_id) %>%
  filter(rgn_id != 213) # filter out antarctica

# Make a data frame with every region and every product
products <- data.frame(product = c("seaweeds", "ornamentals", "fish_oil"))

region_product <- full_join(rgns_eez, products, by = character()) %>%
  full_join(years_df, by = character())


# Calculate $/tonne for each OHI region and product and gapfill based on UN geopolitical region 

harvest_tonnes_usd_geo_rgn <- harvest_tonnes_usd %>%
    full_join(region_product, by = c("rgn_id", "product", "year")) %>%
  add_georegion_id() %>%
  dplyr::filter(year %in% years) %>%
  arrange(rgn_id)
  
# calculate georgn values (for each region, year, and product)
geo_rgn_values_df <- harvest_tonnes_usd_geo_rgn %>%
  dplyr::group_by(georgn_id, year, product) %>%
  dplyr::mutate(values = usd/tonnes) %>%
    mutate(values = ifelse(values == "Inf", NA, values)) %>% ## assign Inf as NA
  summarise(georgn_values = mean(values, na.rm = TRUE)) %>% 
  ungroup() ## now we have a 5 year average value for geo region

# calculate global values (for each product across all regions for each year)
global_values_df <- harvest_tonnes_usd_geo_rgn %>%
  dplyr::group_by(product, year) %>%
  mutate(values = usd/tonnes) %>%
  mutate(values = ifelse(values == "Inf", NA, values)) %>% ## assign Inf as NA
  summarise(global_values = mean(values, na.rm = TRUE)) ## now we have a 5 year average global value


# Create final gapfilled data frame
harvest_tonnes_usd_values_gf <- harvest_tonnes_usd_geo_rgn %>%
  left_join(geo_rgn_values_df, by = c("georgn_id", "product", "year")) %>%
  left_join(global_values_df, by = c("product", "year")) %>%
  mutate(values = usd/tonnes) %>%
  mutate(values = ifelse(values == "Inf", NA, 
                         ifelse(values == "NaN", NA, values))) %>% ## assign Inf (n/0) and NaNs (0/0) as NA
  ## gapfill so that when there is no regional value, take the geo regional value, 
  mutate(values_final = ifelse(is.na(values) & georgn_values != "NaN", georgn_values, 
                               # and if there is no geo regional value, take the global value. Else, keep original values. 
                               ifelse(is.na(values) & georgn_values == "NaN", global_values, values))) %>% 
  ## make note of gapfilling if any occurred in the previous step
  mutate(values_gf_description = ifelse(is.na(values) & georgn_values != "NaN", "geo_region",
                                        ifelse(is.na(values) & georgn_values == "NaN", "global", "none"))) %>%
  mutate(values_gf = ifelse(is.na(values) & georgn_values != "NaN", 1,
                            ifelse(is.na(values) & georgn_values == "NaN", 1, 0))) ## now we have a dataset with gapfilled average values and calculated yearly values... now we will save this, and then calculate the 5 year averages

# Save this data frame to int
readr::write_csv(harvest_tonnes_usd_values_gf, here(current_np_dir, "int", "harvest_tonnes_usd_weighting_gf.csv"))

# Now calculate 5 year averages (and assign year = 2020 <- legacy comment, no longer what the code does)
weighting_usd_values_final <- harvest_tonnes_usd_values_gf %>%
  dplyr::select(rgn_id, product, year, values_final) %>%
  group_by(rgn_id, product) %>%
  summarise(value_per_tonne = mean(values_final)) %>%
  ungroup() 

#save this df
readr::write_csv(weighting_usd_values_final, here(current_np_dir, "int", "harvest_weighting_values.csv"))

```

Now prep the tonnes data from each product to be per each region for the
6 year average (2013-2018).

```{r}
## Read in FOFM tonnes data
fofm_tonnes <- read_csv(here(current_np_dir, "int", "mean_catch_FOFM.csv")) %>%
  dplyr::filter(year %in% years) %>%
  mutate(product = "fish_oil") %>%
  group_by(rgn_id, year, product) %>%
  summarise(tonnes = sum(catch)*0.3) %>% ## Note: multiply by 0.3 to account for water loss when converting fish to fish oil.. about 30% of fish are water and 70% are fish oil
  ungroup() %>%
  group_by(rgn_id, product) %>%
  summarise(tonnes = mean(tonnes)) %>% ##calculate 5 year mean of catch
  ungroup() %>%
  full_join(rgns_eez) %>%
  dplyr::select(1:3) %>%
  mutate(product = "fish_oil") %>%
  mutate(tonnes = ifelse(is.na(tonnes), 0, tonnes)) 


## Read in ornamentals tonnes data
np_tonnes <- read_csv(here(current_np_dir, "int", "np_harvest_tonnes_usd.csv")) %>% 
  filter(rgn_id != 213)

orn_fill_df <- region_product %>%
  filter(product == "ornamentals")
  
orn_tonnes <- np_tonnes %>%
  dplyr::filter(year %in% years, product == "ornamentals") %>%
  full_join(orn_fill_df, by = c("rgn_id", "year", "product")) %>% ## full join to fill in the rest of the regions with 0
  mutate(tonnes = ifelse(is.na(tonnes), 0, tonnes),
         usd = ifelse(is.na(usd), 0, usd)) %>% ## gapfill the NAs to be 0
  select(rgn_id, year, product, tonnes) %>%
  group_by(rgn_id, product) %>%
  summarise(tonnes = mean(tonnes)) %>% ## calculate 5 year mean of catch
  ungroup()
  
## Read in seaweed tonnes data
sw_fill_df <- region_product %>%
  filter(product == "seaweeds")

sw_tonnes_raw <- read_csv(here(current_np_dir, "int", "np_seaweeds_tonnes_weighting.csv")) 

sw_tonnes <- sw_tonnes_raw %>%
  mutate(product = "seaweeds") %>%
  group_by(rgn_id, year, product) %>%
  summarise(tonnes = sum(tonnes, na.rm = TRUE)) %>%
  dplyr::filter(year %in% years) %>%
  full_join(sw_fill_df, by = c("rgn_id", "year", "product")) %>%
    mutate(tonnes = ifelse(is.na(tonnes), 0, tonnes)) %>% ## gapfill the NAs to be 0
  dplyr::select(rgn_id, year, product, tonnes) %>%
  ungroup() %>%
  group_by(rgn_id, product) %>%
  summarise(tonnes = mean(tonnes)) %>% ## calculate 5 year average
  ungroup()


```

Now we will calculate the weights per each product. To do this we need
to multiply our average \$ value for each product \* tonnes of each
product, and then divide by the total per each region. We will also
assign year = (YYYY, e.g., 2022) so that this can be read into OHI-global (YYYY
corresponds to the year that these weights were calculated).

```{r}
harvest_weighting_values <- read_csv(here(current_np_dir, "int", "harvest_weighting_values.csv")) %>%
  dplyr::select(rgn_id, product, value_per_tonne)

prod_weights <- orn_tonnes %>%
  bind_rows(sw_tonnes) %>%
  bind_rows(fofm_tonnes) %>% 
  left_join(harvest_weighting_values, by = c("rgn_id", "product")) %>%
  arrange(rgn_id) %>%
  mutate(usd_product = tonnes*value_per_tonne) %>%
  group_by(rgn_id) %>%
  mutate(total_usd = sum(usd_product)) %>%
  ungroup() %>%
  mutate(weight = usd_product/total_usd) %>%
  dplyr::select(-3, -4, -5, -6) %>%
  mutate(year = scen_year_number) %>%
  dplyr::select(rgn_id, year, product, weight) %>%
  filter(rgn_id != 213)

#write.csv(prod_weights, file.path(prep, "output/np_product_weights.csv"), row.names = FALSE)
readr::write_csv(prod_weights, here(current_np_dir, "output", "np_product_weights.csv"))

```


### Datacheck

Note: we're not comparing data for a specific year, e.g., 2019, as we do in many data checks. The data processing that creates `np_product_weights.csv` fills the `year` column with the scenario year (e.g., 2024), which means that the previous year's `np_product_weights.csv` and the current year's weights cannot be compared by filtering to a specific year. 


```{r fig.height=5, fig.width=8}

# Read in previous and current years' np_product_weights to compare
#old_prod_weights <- read_csv(file.path(prep, paste0("../v", prep_year-3, "/output/np_product_weights.csv")))
#old_prod_weights <- read_csv(file.path(prep, paste0("../v", prep_year-2, "/output/np_product_weights.csv")))
old_prod_weights <- read_csv(here(previous_np_dir, "output", "np_product_weights.csv"))
prod_weights <- read_csv(here(current_np_dir, "output", "np_product_weights.csv"))

check <- prod_weights %>%
  rename("new_weight" = "weight") %>%
  left_join(old_prod_weights, by = c("rgn_id", "product")) %>%
  mutate(diff = new_weight - weight) %>%
  left_join(rgns_eez)

# note -- these are across all years...

plot(check$new_weight, check$weight, main = "All Products")
abline(0,1, col = "red")

check_sw <- check %>%
  filter(product == "seaweeds")

plot(check_sw$new_weight, check_sw$weight, main = "Seaweeds")
abline(0,1, col = "red")

check_orn <- check %>%
  filter(product == "ornamentals") %>%
  mutate(difference = new_weight - weight)

plot(check_orn$new_weight, check_orn$weight, main = "Ornamentals")
abline(0,1, col = "red")


check_fofm <- check %>%
  filter(product == "fish_oil")

plot(check_fofm$new_weight, check_fofm$weight, main = "Fish Oil / Fish Meal (FOFM)")
abline(0, 1, col = "red") ## these change because of new SAUP fisheries catch data... 

# max(check$diff, na.rm = TRUE)
# min(check$diff, na.rm = TRUE)

top_10_diffs <- check %>% arrange(desc(abs(diff))) %>% head(n = 10)
top_10_diffs %>% relocate(diff, .after = "rgn_id")
# rgn 164 = Belize 
    # ~ had a "seaweeds" weight of 0.00 in 2023, now has a weight of 0.98 ≈ 0.9845 diff (v2024)
    # ~ had a "fish_oil" weight of 1.00 in 2023, now has a weight of 0.015 ≈ -0.9845 diff (v2024) 
# 155 = Tonga 
    # ~ had an "ornamentals" weight of 0.97 in 2023, now has a weight of 0.00 ≈ -9.77 (v2024)
# 153 = Cook Islands
# 11 = Marshall Islands

# for Cook Islands and Marshall Islands, almost feels like values for `ornamentals` and `fish_oil` got flipped?

## check saint martin, Niue, Bonaire, Sint Eustasius - should decrease in production
## check djibouti - should increase in production

test <- harvest_tonnes_usd %>%
  filter(rgn_id == 46)

harvest_tonnes_usd_old <- #read_csv(file.path(prep, paste0("../v", prep_year-1, "/int/np_harvest_tonnes_usd.csv"))) %>%
  read_csv(here(previous_np_dir, "int", "np_harvest_tonnes_usd.csv")) %>%
  filter(rgn_id != 213)

test2 <- harvest_tonnes_usd_old %>%
  filter(rgn_id == 46)

# test
# test2
# v2024: Djibouti (rgn 46) does increase in production

#gf_orn <- read_csv(here(current_np_dir, "output", "np_ornamentals_harvest_tonnes_gf.csv"))


```




Old old check:

```{r}
# Read in previous and current years' np_product_weights to compare
#old_prod_weights <- read_csv(file.path(prep, paste0("../v", prep_year-3, "/output/np_product_weights.csv")))
#old_prod_weights <- read_csv(file.path(prep, paste0("../v", prep_year-2, "/output/np_product_weights.csv")))
old_prod_weights <- read_csv(here(previous_np_dir, "output", "np_product_weights.csv"))
older_prod_weights <- read_csv(here("globalprep", "np", "v2022", "output", "np_product_weights.csv"))

check_older <- older_prod_weights %>%
  rename("older_weight" = "weight") %>%
  left_join(old_prod_weights, by = c("rgn_id", "product")) %>%
  mutate(diff = weight - older_weight) %>%
  left_join(rgns_eez)


plot(check_older$weight, check_older$older_weight, main = "v2023 vs v2022"); abline(0,1, col="red")

plot(check$new_weight, check$weight, main = "v2024 vs v2023"); abline(0,1, col="red")


```



```{r}
library(plotly)
library(RColorBrewer)


test_comp <- check_older %>% plotly::plot_ly(x = ~older_weight, y = ~weight,
                                             type = "scatter", mode = "markers") %>% 
  add_trace(
    # text = ~paste("Region ID:", rgn_id #, "<br>Product:", product
    #               ),
   color = ~rgn_id,
   # marker = list(
   #    color = ~product,
   #  ),
   hoverinfo = "text+x+y",
   hovertemplate = paste(
     '<b>Region ID</b>: %{color}',
     '<br><b>Old Weight</b>: %{x:.2f}',
     '<br><b>New Weight</b>: %{y:.2f}<br>'
    ),
    #  showlegend = F
  ) %>% 
  # add AB line
  add_trace(
    x = c(0,1),
    y = c(0,1),
    mode = "lines",
    line = list(color = "red"),
    showlegend = FALSE,
    hoverinfo = "none"
  ) %>%
  
  layout(xaxis = list(hoverformat = '.3f'),
         yaxis = list(hoverformat = '.3f'))

test_comp
```


- RAM lost some stocks, so some stocks get removed in v4.65 (2024 data download); also got some new stocks


### New datacheck: interactive plot

```{r}
# ======== Setup ==========================================================
# (mostly copied from earlier setup with addition of library(plotly))
# Read in packages necessary for visualization (in case you cleared your environment)
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyr)
library(here)

# Set scenario year, reproducible file paths
scen_year_number <- 2024 # update this!!
scenario_year <- as.character(scen_year_number)
v_scen_year <- paste0("v", scenario_year)

current_np_dir <- here::here("globalprep", "np", v_scen_year)
previous_np_dir <- here::here("globalprep", "np", paste0("v", scen_year_number - 1))

# ----- Read in NP weights data  -----
# Read in current, previous, and 2 years previous data
old_prod_weights <- read_csv(here(previous_np_dir, "output", "np_product_weights.csv"))
prod_weights <- read_csv(here(current_np_dir, "output", "np_product_weights.csv"))

check <- prod_weights %>%
  rename("new_weight" = "weight") %>%
  left_join(old_prod_weights, by = c("rgn_id", "product")) %>%
  mutate(diff = new_weight - weight) %>%
  left_join(rgns_eez)


#old_prod_weights <- read_csv(here(previous_np_dir, "output", "np_product_weights.csv"))
older_prod_weights <- read_csv(here("globalprep", "np", paste0("v", scen_year_number - 2), "output", "np_product_weights.csv"))

check_older <- older_prod_weights %>%
  rename("older_weight" = "weight") %>%
  left_join(old_prod_weights, by = c("rgn_id", "product")) %>%
  mutate(diff = weight - older_weight) %>%
  left_join(rgns_eez)


# ======== Plot new NP weights data vs. previous year's data ==================


# ---- Create ggplot object with `text` aes argument to prep for plotly ----- #
g_new <- ggplot(check, aes(x = weight, y = new_weight, color = product,
                             text = paste("<b>Region ID:</b>", rgn_id, "<br><b>Product:</b>", product #,
                                          #"<br>Old weight:", older_weight, "<br>New weight:", weight
                                          ))) +
  geom_point(alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.7,
              color = "red", linetype = "dashed") +
  scale_color_manual(values = c("fish_oil" = "#edae49", "seaweeds" = "#386641", "ornamentals" = "#83c5be")) +
  labs(title = "v2024 vs v2023 Natural Product Weights by Product",
       x = "v2023 Weight",
       y = "v2024 Weight",
       color = "Product") +
  theme_minimal()

# ---- Convert to plotly ------------------------------ #
p_new <- ggplotly(g_new, tooltip = "text") %>%
  layout(
    hovermode = "closest",
    xaxis = list(hoverformat = ".4f"), # customize number formatting 
    yaxis = list(hoverformat = ".4f") # (number of places after decimal) 
  )

# ------- Create custom hover template ---------------- #
p_new$x$data <- lapply(p_new$x$data, function(trace) {
  if (trace$mode == "markers") {
    trace$hovertemplate <- paste(
      "%{text}<br>",
      "<b>Old Weight</b>: %{x:.4f}<br>",
      "<b>New Weight</b>: %{y:.4f}<br>"
      )
  }
  return(trace)
})

# Display customized plot
p_new



# ========= Plot old vs. older NP weights data ================================
# Plot previous year's NP weights data vs. year before previous year's data 

#color_scale <- c("fish_oils" = "#edae49", "seaweeds" = "#00798c", "ornamentals" = "#d1495b")

# ---- Create ggplot object with added `text` field to prep for plotly ------ #
g_old <- ggplot(check_older, aes(
  x = older_weight, y = weight, color = product,
  text = paste("<b>Region ID:</b>", rgn_id, "<br><b>Product:</b>", product #,
               #"<br>Old weight:", older_weight, "<br>New weight:", weight
               )
  )) +
  # add scatterplot points
  geom_point(alpha = 0.8) +
  # add ab line
  geom_abline(intercept = 0, slope = 1, alpha = 0.7,
              color = "red", linetype = "solid") +
  # customize colors
  scale_color_manual(values = c("fish_oil" = "#edae49", "seaweeds" = "#386641", "ornamentals" = "#83c5be")) +
  # update labels
  labs(title = "v2023 vs v2022 Natural Products Weights by Product",
       x = "v2022 Weights",
       y = "v2023 Weights",
       color = "Product") +
  # set base theme
  theme_minimal()

# ---- Convert to plotly ------------------------------ #
p_old <- ggplotly(g_old, tooltip = "text") %>%
  layout(
    hovermode = "closest",
    xaxis = list(hoverformat = ".4f"), # customize number formatting 
    yaxis = list(hoverformat = ".4f")
  )

# ------- Custom hover template ----------------------- #
p_old$x$data <- lapply(p_old$x$data, function(trace) {
  if (trace$mode == "markers") {
    trace$hovertemplate <- paste(
      "%{text}<br>",
      "<b>Old Weight</b>: %{x:.4f}<br>",
      "<b>New Weight</b>: %{y:.4f}<br>"
      )
  }
  return(trace)
})

# Display customized plot!
p_old
```


