---
title: 'OHI 2024 - Mariculture Population Pressure Layers'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
---

[REFERENCE RMD FILE](https://ohi-science.org/ohiprep_v2024/globalprep/mar_prs_population/v2024/mar_prs_pop_dataprep.html)


# Summary

This document describes the steps for obtaining and wrangling the data used to calculate the mariculture population pressure sub-dimension for the 2024 global assessment. The general data preparation calculations are summarized [here](http://ohi-science.org/ohi-global/layers.html#inland_coastal_population). For context and explanation see the mariculture (subgoal of food provision) [model summary](http://ohi-science.org/ohi-global/goals.html#food_provision:_mariculture).

# Updates from previous assessment
New UN population data were updated for v2024. CIESIN data were not updated, since the source data has not been updated.

***

# Data Sources

## Gridded Population of the World (v4.11) by CIESE and CIAT

**Reference**: http://sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev11/data-download

Center for International Earth Science Information Network - CIESIN - Columbia University. 2018. Gridded Population of the World, Version 4 (GPWv4): Population Density Adjusted to Match 2015 Revision UN WPP Country Totals, Revision 11. Palisades, NY: NASA Socioeconomic Data and Applications Center (SEDAC). https://doi.org/10.7927/H4F47M65. Accessed 17 April 2019

**Downloaded**: 17 April 2019

**Description**:
The Gridded Population of the World, Version 4 (GPWv4): Population Density Adjusted to Match 2015 Revision of UN WPP Country Totals, Revision 11 consists of estimates of human population density (number of persons per square kilometer) based on counts consistent with national censuses and population registers with respect to relative spatial distribution, but adjusted to match the 2015 Revision of the United Nation's World Population Prospects (UN WPP) country totals, for the years 2000, 2005, 2011, 2015, and 2020. A proportional allocation gridding algorithm, utilizing approximately 13.5 million national and sub-national administrative units, was used to assign UN WPP-adjusted population counts to 30 arc-second grid cells. The density rasters were created by dividing the UN WPP-adjusted population count raster for a given target year by the land area raster.  Documentation for gridded population of the world is located [here](http://sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev11/docs).

**Native data resolution**: 30 arc-seconds

**Time range**: 2000-2020

**Format**: GeoTiff

## UN Population

UN population data is used in this data prep routine to check population counts; spatially distributed counts derived from the gridded world population are aggregated by region and checked against the UN population estimates.

**Reference**: https://esa.un.org/unpd/wpp/

**Downloaded**: 26 June 2020

**Description**: Population (in thousands) for countries.

**Native data resolution**: Country scores

**Time range**: 1950-2020

**Format**: Excel file

**Data cleaning process**: Values from the "ESTIMATES" tab of "WPP2019_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx" were copied for all countries and years (1950-2020) and pasted into new "UN_pop_clean_v2020.csv".

***

# Setup

Load all relevant libraries including parallel processing packages
```{r setup, message=FALSE, warning=FALSE, verbose=FALSE, eval=FALSE}

## Set options for all chunks in code
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 6, fig.height = 4, fig.path = "figs/")

## Load packages, installing them first where necessary
pkg <- c("raster", "sp", "sf", "fasterize", "tidyverse", "foreach", "parallel","doParallel")
new_pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new_pkg)){install.packages(new_pkg)}
if (!("ohicore" %in% installed.packages())){devtools::install_github("ohi-science/ohicore")}

library(ohicore)

## Spatial libraries
library(raster)
# library(rgdal)
library(sp)
library(fasterize)
library(sf) # Vector spatial data
library(terra) # Raster spatial data
library(tidyterra) # For rasters in GGplot

## Data wrangling libraries
library(tidyverse)
library(here)
library(plotly)

## Parallel processing libraries
library(parallel)
library(foreach)
library(doParallel)

```

Define frequently used pathnames. Change scenario and data years in file pathnames code chunk to reflect the most recent data (d) and current assessment year (v).
```{r file paths, eval=FALSE}

## Source common and spatial common files
source('http://ohi-science.org/ohiprep_v2020/workflow/R/common.R')

## Update these!
scenario_yr <- "v2024" # change to reflect assessment year!
data_yr_gpw <- "d2024" # change to reflect year of most recently downloaded Gridded Population of the World (GPW) data! (no change for v2020)
data_yr_un_pop <- "d2020" # change to reflect year of most recently downloaded UN population data! (updated for v2020)

## Define commonly used file paths (matched to what we did in ico_data_prep.Rmd)
dir_server <- file.path(dir_M, "git-annex/globalprep/_raw_data")
dir_github <- here("/globalprep/mar_prs_population", scenario_yr)

## Checking to see if there is a README in the mar_prs_pop and mar_prs_pop/v20?? folders - all goal prep files will need a README!
if(!file.exists(file.path(dir_github, 'README.md'))) {
  warning(sprintf('No README detected in %s', dir_github))
}
if(!file.exists(file.path(dirname(dir_github), 'README.md'))) {
  warning(sprintf('No README detected in %s', file.path(dirname(dir_github))))
}

```

## Import Raw Data

```{r import raw data, eval=FALSE}

## Read in the raw density data to be reprojected and resampled
raw <- list.files(file.path(dir_server, sprintf("CIESEandCIAT_population/%s", data_yr_gpw)),
                  full.names = TRUE, pattern = "\\.tif$",
                  recursive = TRUE)
raw <- raw[grep("count_", raw)] # keep only the files that include the word "density"
raw # check that this looks correct; can double check with the folder on Mazu server folder
i <- terra::rast(raw[4])

ohi_rasters() # loads rasters
# eez_raster <- zones # sourced from ohi_rasters()

## Import raw, cleaned UN population data to be wrangled then used to confirm gpw-derived spatial population
pop <- read.csv(file.path(dir_server, 
                          sprintf("UnitedNations_population/%s/UN_pop_clean_v2020.csv", data_yr_un_pop)), # change year if updated
                strip.white = TRUE, stringsAsFactors = FALSE)

```

# Methods and Calculations

## Reproject and Resample

Gridded world population densities from CIESE and CIAT are projected into the World Mollweide coordinate reference system and resampled using nearest neighbor method to match the ocean raster which is sourced from spatial_commons.R and has 934.4789 x 934.4789 meter resolution.

```{r manipulate spatial data, eval=FALSE}

## Resample resolution and CRS of population data, convert year to numeric, project rasters and save human_density_year_mol.tifs to server with these updates
# Use temporary file path name (_tmp.tif) so that R isn't trying to load and write the same file within the loop - MAKE SURE TO DELETE THESE FILES AFTER THE LOOP IS COMPLETE! 
# Set up a loop to go through raw data frame, convert to a mollweide CRS:

cl <- makeCluster(5) # 5 clusters for 5 files being targeted
registerDoParallel(cl)
foreach(i = raw, .packages = c("raster", "stringr")) %dopar% {
 i = raw[5] # send this down alone and check to see which file it calls; see if there are any errors in file naming
  yr <- as.numeric(as.character(str_sub(i, -15, -12))) # counts backwards by character in file name to grab the year digits
  i = raster::raster(i) %>% # read in file path to raster
    raster::projectRaster(crs = crs(ocean), res = res(ocean), over = TRUE, method = "bilinear", # convert large grids to small grids using bilinear method
                          filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_density_%s_mol_tmp.tif", scenario_yr, yr)), # write to the server because the files are so large
                          overwrite = TRUE, progress = "text")
  
raster::raster(file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_density_%s_mol_tmp.tif", scenario_yr, yr))) %>% 
  raster::resample(ocean, method = 'ngb', overwrite = TRUE, filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_density_%s_mol.tif", scenario_yr, yr)))
} 

# This takes a while to run: ~30-45 min; will generate some warnings

# Read in the created resampled raster and make sure it matches the eez_25mi_inland raster (projection and extents)

stopCluster(cl)
closeAllConnections()

# test <- raster(file.path("/home/shares/ohi/git-annex/globalprep/mar_prs_population/v2024/int/human_density_2020_mol.tif"))
# plot(test)
# #
# test2 <- raster::select(test)
# plot(test2)
```

## Interpolate between Years

GWPv4 data is for years 2005, 2010, 2015, and 2020. Data for missing years must be generated by interpolation.

```{r functions for interpolation, eval=FALSE}

## Define and apply function to calculate yearly change in population density 

## These are the files that were created and written out above. Dont exactly need them since we are working with just counts at this point 
files <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int"), pattern = "human_density_\\d{4}_mol.tif$", full = TRUE)

## Create function for average yearly change
yearly_diff <- function(year_min, year_max, density_files = files){
  rast_diff <- raster::stack(density_files[grep(year_max, density_files)], density_files[grep(year_min, density_files)]) %>%
    raster::overlay(fun = function(x, y){(x - y)/5},
            filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/yearly_change_%s_to_%s.tif", scenario_yr, year_min, year_max)),
            overwrite = TRUE)
}


## Create four rasters for each set of data
yearly_diff(year_min = 2000, year_max = 2005) # this should save 2001, 2002, 2003, etc to estimate intervening years using a linear model - check to make sure it's being saved in the Mazu directory properly
yearly_diff(year_min = 2005, year_max = 2010)
yearly_diff(year_min = 2010, year_max = 2015)
yearly_diff(year_min = 2015, year_max = 2020) # this last one wouldn't run; copied the tif from v2019 since it's the same data

library(terra)

yearly_diff_terra <- function(year_min, year_max, count_files = raw, scenario_yr, dir_M){
  # Load the raster files
  # contingency for rasters with year = 2015 because the grep dont work (picks up all the files because they all have a 2015 in them)
  if(year_max == 2015){
    contingency <- paste0("rev11_", year_max)
    rast_max <- rast(count_files[grep(contingency, count_files)])
  } else {
    rast_max <- rast(count_files[grep(year_max, count_files)])
  }
  if(year_min == 2015){
    contingency2 <- paste0("rev11_", year_min)
    rast_min <- rast(count_files[grep(contingency2, count_files)])
  } else {
    rast_min <- rast(count_files[grep(year_min, count_files)])
  }
  
  # Calculate the difference and divide by 5
  rast_diff <- (rast_max - rast_min) / 5
  
  # Define the output file path
  output_file <- file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/yearly_change_%s_to_%s.tif", scenario_yr, year_min, year_max))
  
  # Save the result to a file
  terra::writeRaster(rast_diff, filename = output_file, overwrite = TRUE)
}

yearly_diff_terra(2000, 2005, count_files = raw, dir_M = dir_M, scenario_yr = scenario_yr)

# test_file <- file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/yearly_change_%s_to_%s.tif", scenario_yr, year_min, year_max)) # returns the file that I want 

yearly_diff_terra(2005, 2010, count_files = raw, dir_M = dir_M, scenario_yr = scenario_yr)

yearly_diff_terra(2010, 2015, count_files = raw, dir_M = dir_M, scenario_yr = scenario_yr)

yearly_diff_terra(2015, 2020, count_files = raw, dir_M = dir_M, scenario_yr = scenario_yr)


## Define and apply function to interpolate population densities for years between years in dataset
# Apply the difference to each year, overwrite as new rasters
# Can double check this by running just one (i = 1)
yearly_interpolate <- function(year, start_raster, yearly_change){
    for(i in 1:4){
      # i = 1 # send this down to see what the loop produces when i=1
      raster::overlay(
        raster::raster(start_raster),
        raster::raster(yearly_change),
        fun = function(x, y){(x + i*y)},
        filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_density_%s_mol.tif", scenario_yr, (year+i))),
              overwrite = TRUE)
    }
}

############# First function: not working and I think its because of the file naming convention ###########
# File calls <- contingency <- paste0("rev11_", year_max)
  # rast_max <- rast(count_files[grep(contingency,count_files)])
yearly_interpolate_terra <- function(year, start_raster, yearly_change) {
  for(i in 1:4){
    start_raster <- terra::rast(start_raster)
    yearly_change_raster <- terra::rast(yearly_change)
    
    yearly_int = (start_raster) + (i * yearly_change_raster)
    
    output_file <- file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, (year+i)))
  
  # Save the result to a file
  terra::writeRaster(yearly_int, filename = output_file, overwrite = TRUE)
  }
}
library(dplyr)
raw1 <- terra::rast(raw[1]) %>% 
  terra::writeRaster(., filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, 2000)))

raw2 <- terra::rast(raw[2]) %>% 
  terra::writeRaster(., filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, 2005)))

raw3 <- terra::rast(raw[3]) %>% 
  terra::writeRaster(., filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, 2010)))

raw4 <- terra::rast(raw[4]) %>% 
  terra::writeRaster(., filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, 2015)))

raw5 <- terra::rast(raw[5]) %>% 
  terra::writeRaster(., filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, 2020)))

############ Second function with error troubleshooting #############
yearly_interpolate_terra <- function(year, start_raster, yearly_change) {
  for(i in 1:4) {
    # Read rasters
    start_raster <- tryCatch({
      r <- terra::rast(start_raster)
      if (is.null(terra::values(r))) stop("Start raster has no values")
      r
    }, error = function(e) {
      stop(sprintf("Failed to read start raster for year %s: %s", year, e$message))
    })
    
    yearly_change_raster <- tryCatch({
      r <- terra::rast(yearly_change)
      if (is.null(terra::values(r))) stop("Yearly change raster has no values")
      r
    }, error = function(e) {
      stop(sprintf("Failed to read yearly change raster for year %s: %s", year, e$message))
    })
    
    # Calculate the new raster
    yearly_int <- tryCatch({
      (start_raster) + (i * yearly_change_raster)
    }, warning = function(w) {
      warning(sprintf("Warning during raster calculation for year %s: %s", year, w$message))
    }, error = function(e) {
      stop(sprintf("Error during raster calculation for year %s: %s", year, e$message))
    })
    
    # Define output file path
    output_file <- file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, (year+i)))
    
    # Save the result to a file
    tryCatch({
      terra::writeRaster(yearly_int, filename = output_file, overwrite = TRUE)
    }, error = function(e) {
      stop(sprintf("Failed to write raster for year %s: %s", year+i, e$message))
    })
  }
}

## Interpolate missing years
files <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int"), pattern = ".tif$", full = TRUE)

for(i in c(2000, 2005, 2010, 2015)) {
  yearly_interpolate_terra(i, start_raster = count_files[grep(sprintf("rev11_%s",i), count_files)],
                     yearly_change = files[grep(sprintf("change_%s_to_%s", i, i+5), files)])
}

for(i in c(2000, 2005, 2010, 2015)) {
  tryCatch({
    start_raster <- files[grep(sprintf("count_%s_mol", i), files)]
    yearly_change <- files[grep(sprintf("change_%s_to_%s", i, i+5), files)]
    
    if (length(start_raster) == 0) {
      stop(sprintf("Start raster for year %s not found.", i))
    }
    if (length(yearly_change) == 0) {
      stop(sprintf("Yearly change file for year %s to %s not found.", i, i+5))
    }
    
    yearly_interpolate_terra(i, start_raster, yearly_change)
  }, error = function(e) {
    message(sprintf("Error processing year %s: %s", i, e$message))
  })
}
# If getting a "Cannot create a RasterLayer object from this file" error after running this loop, check that previously-created _tmp.tif files have been deleted from the mazu directory

# # Check a random raster from the mazu directory
# density_2018 <- raster(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int/human_density_2012_mol.tif"))
# plot(density_2018)

```

## Convert Density to Population Counts

Density data are converted to population by multiplying density times the area of the cell (934.4789 x 934.4789m) and converting to square km from square meters, to get people per square km. GWPv4 provides both densities and population counts. However, we derive population counts from density because the density data is independent of raster scale, so we can reproject/resample and it will still be correct. Change the scale using count data, and it is necessary to correct by dividing by cell area otherwise the values become meaningless; it is easier to just start with density.

```{r convert density data to population count, eval=FALSE}

den_raster <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int"), pattern = "density_\\d{4}_mol.tif$", full = TRUE)
den_raster <- den_raster[21] # check to make sure this looks correct; should only have human_density tif files for each interpolated year
den_raster
cl <- makeCluster(5)
registerDoParallel(cl)

foreach(i = den_raster, .packages = c("raster", "stringr")) %dopar% {
  # i = den_raster #  - does i generate the 2000 raster?
  nm <- basename(i)
  yr <- stringr::str_sub(nm, -12, -9) # pulls out the year from the file name
  cat(nm)
  tmp <- raster::raster(i) # reads in raster
  raster::calc(tmp, fun = function(x){x * (934.4789 * 934.4789 * 0.000001)}, # convert to m2 from km2
       filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/v2024/int/human_count_%s_mol.tif", yr)),
       overwrite=TRUE) # writes a human count raster to the intermediate folder
}
stopCluster(cl)
closeAllConnections()

```

## Load EEZ + 25-mile Inland Raster

```{r create eez + 25-mile inland raster, eval=FALSE}

# No longer using shape files, raster only

# Raster
eez_25mi_inland <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif")) # last updated v2019
eez_25mi_inland_terra <- terra::rast(file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif"))
## Make sure that the layers make sense!
plot(eez_25mi_inland)# will appear to only feature the Antarctic ocean; can check that gray area actually has data. Because we don't include this region in the OHI, the raster has CCAMLR IDs for points in this region, which appear to have high region ID values.
plot(eez_25mi_inland_terra)

ggplot() + 
  geom_spatraster(data = eez_25mi_inland_terra, aes(fill = ocean_plus25mile_inland))

click(eez_25mi_inland)
# Click on gray areas on the map and region IDs will pop up - the data are there! 

# Want to make sure we're including 10km inland - zoom in on smaller region using zoom()
zoom(eez_25mi_inland)
# Once you select a small area to zoom in on, you can layer the outline of the land and EEZ boundaries to make sure the EEZ goes 10km inland:
regions_shape()
plot(regions[1], color = NA, add = TRUE)

```

## Reference Chunk for working with weird rasters 

```{r}
# rgn_rast = terra::rast(rgn_rast_list[1]) ## Create SpatRaster object 
# rgn_rast  = eez_25mi_inland ## Add zero 
# df = freq(rgn_rast) ## Turns frequency table into a dataframe (should have 3 columns: layer, value, count) 
# rgn_rast2 <- raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif"))##Create RasterLayer object 
# 
# df2 = as.data.frame(rgn_rast2@data@attributes) ## Creates dataframe from attributes (Should have 3 columns: ID, OBJECTID, count) 
# df$difference = df$value - df2$ID ## Creating new column in 'df' that is the difference between 'value' and 'ID'  
# unique(df$difference) ## Should return one value: 0 
```


## Append or Extract by OHI Regions

Stack rasters of population counts for each year, sum within regions (using eez_25mi_inland) for each year, and save this as `coastal_pop_zonal_sums.csv` in the intermediate folder. Gather this into long format dataframe, and save as `output/mar_pop_25mi.csv`.

```{r extract population data from all population layers, eval=FALSE}

pop_rasters <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int"), pattern = "count_\\d{4}_mol.tif$", full = TRUE) # should list all the human_count files in folder
pop_rasters[21]
pop_2020 <- terra::rast(pop_rasters[21])
?terra::rast
pop_2020_test <- raster::raster(pop_rasters[21])
pop_stack <- raster::stack(pop_rasters) # stack human_count rasters 2000-2020

# ####### TESTING ######### ---- 
# pop_raster_test <- file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int/gpw_v4_population_count_rev11_2020_30_sec.tif")
# 
# pop_rast <- raster(pop_raster_test)
# 
# crs(pop_rast) <- crs(eez_25mi_inland)
# 
# extent(pop_rast) <- extent(eez_25mi_inland)
# 
# pop_rast <- raster::resample(pop_rast, eez_25mi_inland, method = "bilinear")
eez_vect <- terra::as.polygons(eez_25mi_inland_terra)
coastal_pop <- raster::zonal(world_count, eez_25mi_inland, fun="sum", progress = "text") # sum within zones (regions as specified by eez_25mi_inland)

coastal_pop <- as.data.frame(coastal_pop)

write_csv(coastal_pop, file.path(here("globalprep/mar_prs_population/v2024/int/coastal_pop_zonal_sums.csv"))) # do we need to save this as intermediate data?

## Organize data to have a population sum for each region and each year
coastal_pop2 <- data.frame(coastal_pop) %>%
  tidyr::gather("year", "coastal_pop_25mi", -1) %>%
  dplyr::select(rgn_id = zone, year = year, popsum = coastal_pop_25mi) %>%
  dplyr::mutate(year = as.numeric(substring(year, 13, 16))) %>% # extract numeric portion of year_####
  dplyr::filter(rgn_id <= 250)

write_csv(coastal_pop2, file.path(here("globalprep/mar_prs_population/v2024/output/mar_pop_25mi.csv")))

```

## Calculate Area of 25mi Inland Buffers

```{r calculate area and mar. pressure, eval=FALSE}

# Load in data calculated in spatial folder 
area <- read_csv(here("globalprep/spatial/v2019/output/area_km2_25mi_inland.csv")) # last updated v2019

```

## Calculate Mariculture Population Pressure

```{r calculate mar. population pressure, eval=FALSE}

## Rescale organized coastal population data
pop_rescaled <- coastal_pop2 %>%
  left_join(area, by="rgn_id") %>%
  mutate(density = popsum/area_km2) %>%
  mutate(ln_density = log(density + 1)) %>%
  mutate(scalar = max(ln_density, na.rm = TRUE)) %>%
  mutate(dens_rescaled = ln_density/scalar) %>%
  mutate(dens_rescaled = ifelse(dens_rescaled > 1, 1, dens_rescaled))

filter(pop_rescaled, is.na(area_km2)) # no NA values for the area column (Antarctica not included)

pressure_data <- pop_rescaled %>%
  dplyr::select(rgn_id, year, pressure_score = dens_rescaled)

write_csv(pressure_data, file.path(here("globalprep/mar_prs_population/v2024/output/prs_pop_density.csv")))

```

## Gapfilling

No gapfilling was completed as part of this data prep methodology. Datasets are saved with "_gf" appended just to indicate they are finalized versions.

```{r save prs_pop_density and mar_pop_25mi layers, eval=FALSE}

prs <- read.csv("globalprep/mar_prs_population/v2020/output/prs_pop_density.csv") %>%
  dplyr::mutate(pressure_score = 0)

write.csv(prs, "globalprep/mar_prs_population/v2020/output/prs_pop_density_gf.csv", row.names = FALSE)

mar <- read.csv("globalprep/mar_prs_population/v2020/output/mar_pop_25mi.csv") %>%
  dplyr::mutate(popsum = 0)

write.csv(mar, "globalprep/mar_prs_population/v2020/output/mar_pop_25mi_gf.csv", row.names = FALSE)

```

## Data Checks and/or Meta-Analysis

### Create `UN_population.csv` for Data-checking

```{r create un_population.csv, eval=FALSE}

pop_gather <- pop %>%
  tidyr::gather("year", "population", starts_with("X")) %>%
  dplyr::mutate(population = gsub(" ", "", population)) %>%
  dplyr::mutate(year = gsub("X", "", year)) %>%
  dplyr::mutate(population = as.numeric(as.character(population)) * 1000)

## Ignore Jersey and Guernsey (Channel Islands) for now
pop_gather_rename <- pop_gather %>%
  dplyr::mutate(country = ifelse(str_detect(country,"C\\Ste d'Ivoire"), "Cote d'Ivoire", country)) %>%
  dplyr::mutate(country = ifelse(str_detect(country,"R\\Sunion"), "Reunion", country)) %>%
  dplyr::mutate(country = ifelse(str_detect(country,"Cura\\Sao"), "Curacao", country)) %>%
  dplyr::mutate(country = ifelse(country=="China, Taiwan Province of China", "Taiwan", country)) %>%
  dplyr::mutate(country = ifelse(country=="Dem. People's Republic of Korea", "North Korea", country))
  
## Organize the data into regions used in OHI, and save
pop_rgn <- name_2_rgn(df_in = pop_gather_rename, 
                      fld_name='country', 
                      flds_unique=c('year'))

pop_rgn <- pop_rgn %>%
  dplyr::group_by(rgn_id, year) %>%
  dplyr::summarize(population = sum(population)) %>%
  data.frame()

write.csv(pop_rgn, "globalprep/mar_prs_population/v2020/output/UN_population.csv", row.names = FALSE)
ohi_rgns <- ohicore::rgn_master
```

### Compare `UN_population` with Calculated Count

```{r check full calculated population against latest UN population data, eval=FALSE}

compare_yr <- 2020 # doing below functions for comparison between raster and UN population data for 2015

## Sum counts regionally for the scenario year
pop_rast <- raster(file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/int/human_count_%s_mol.tif", scenario_yr, compare_yr)))
zones <- raster(file.path(dir_M, "git-annex/globalprep/spatial/v2017/regions_land_ocean.tif")) # pulling from 2017 server data
pop_counts <- zonal(pop_rast, zones, fun = "sum", na.rm = TRUE, progress="text")

pop_UN <- pop_rgn %>%
  dplyr::filter(year == compare_yr) %>%
  dplyr::select(rgn_id, pop_UN = population)

## Join filtered UN population and summed calculated counts
compare_pop <- data.frame(pop_counts) %>%
  dplyr::select(rgn_id = zone, pop_rast = sum) %>%
  dplyr::left_join(pop_UN, by="rgn_id") %>% 
  filter(!is.na(pop_UN))
options(scipen = 999)

compare_pop$diff <- compare_pop$pop_rast - compare_pop$pop_UN

median <- median(compare_pop$diff)
pct_diff <- compare_pop$diff/compare_pop$pop_UN
median_diff <- median(pct_diff)
median_diff
## Comparing to the population data that came from ArcGIS Pro! 

fp_gis <- file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/coastal_pop/WGS_1984_OHI_Coastal_Population_Estimates_TableToExcel.xlsx")

gis_pop <- readxl::read_xlsx(fp_gis, col_names = TRUE) %>% 
  dplyr::select(-c(OBJECTID, COUNT, AREA)) %>% 
  dplyr::rename(rgn_id = Value)

compare_pop <- compare_pop %>% 
  dplyr::left_join(gis_pop, by = "rgn_id")

# Difference between gis data population estimates versus UN population estimates
compare_pop$gis_diff <- compare_pop$SUM - compare_pop$pop_UN

median <- median(compare_pop$gis_diff)
pct_diff <- compare_pop$SUM/compare_pop$pop_UN
median_diff <- median(pct_diff)
median_diff
range <- range(compare_pop$gis_diff)
which.min(compare_pop$gis_diff)
## Check plot - investigate outliers (post to issue first)
pop_compare_plot <- ggplot(compare_pop) +
  geom_point(aes(pop_rast, pop_UN, label = rgn_id), alpha = .6) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('Calculated Population %s', compare_yr),
       y = sprintf('UN Population %s', compare_yr),
       title = 'Population Comparison')
ggplotly(pop_compare_plot)

## Plot log values of same comparison
pop_compare_log <- ggplot(compare_pop) +
  geom_point(aes(log(pop_rast), log(pop_UN), label = rgn_id), alpha = .6) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('Log of Calculated Population %s', compare_yr),
       y = sprintf('Log of UN Population %s', compare_yr),
       title = 'Population Comparison (log values)')
ggplotly(pop_compare_log)

```

### Compare to Previous Year

```{r comparison, eval=FALSE}

prev_scen_yr <- paste0("v", as.numeric(substr(scenario_yr, 2, 5)) -1)

old <-  read_csv(here('globalprep/mar_prs_population/v2019/output/mar_pop_25mi.csv')) %>% 
  dplyr::select(rgn_id, year, popsum_old=popsum)

tmp <- coastal_pop2 %>%
  dplyr::left_join(old, by=c("rgn_id", "year"))

years_compare_plot <- ggplot(tmp) +
  geom_point(aes(log(popsum), log(popsum_old), label = rgn_id), alpha = .6) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('Log of Calculated Population %s', scenario_yr),
       y = sprintf('Log of Calculated Population %s', prev_scen_yr),
       title = 'Population comparison between assessments')
ggplotly(years_compare_plot)

```

```{r check_areas, eval=FALSE}

## Compare area with that calculated in previous year (should look very similar)

# v2018: old_area <- read.csv(paste0(prev_scen_yr, "/rgn_area_inland25mi.csv")) %>%
old_area <- read_csv(here('globalprep/mar_prs_population/v2019/int/area_km2_25mi.csv')) %>%
  rename(old_area_km2 = area_km2) %>%
  dplyr::mutate(old_area_km2 = round(old_area_km2)) %>%
  dplyr::left_join(area, by = "rgn_id")

area_compare_plot <- ggplot(old_area) +
  geom_point(aes(area_km2, old_area_km2, label = rgn_id), alpha = .6) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  labs(x = sprintf('%s Area (km2)', scenario_yr),
       y = sprintf('%s Area (km2)', prev_scen_yr),
       title = 'Area Calculation Comparison')
ggplotly(area_compare_plot)


summary(old_area)

```

Find core country outlines 
Find land outlines and create an internal buffer 
Choose which buffer is most appropriate for coastal populations
Use this for eco tourism an clean water pressure 
Rasterize the buffer to use for future reference

**Find where the OHI regions shapefiles are and potentially use those polygons instead of rasters from the get** 

```{r}

count_rast <- raster::raster(raw[5])

count_rast <- terra::rast(count_rast)

eez_25mi_inland <- raster::raster(file.path(dir_M, "git-annex/globalprep/spatial/v2019/ocean_plus25mile_inland.tif"))

eez_25mi_inland <- terra::rast(eez_25mi_inland)

(eez_25mi_inland)

terra::datatype(eez_25mi_inland) # 'FLT4S'

ocean_rast_int <- terra::as.int(eez_25mi_inland) # Changing from float raster to integer raster 

plot(ocean_rast_int)  # Checking to make sure that it makes sense 

ocean_int_reproj <- terra::project(x = ocean_rast_int, count_rast, method="near") # Reprojecting to population raster projection  

plot(ocean_int_reproj) # Checking to make sure that it makes sense

ocean_fact_reproj <- terra::as.factor(ocean_int_reproj) # Factorizing the regions raster to ensure that it is categorical 

plot(ocean_fact_reproj) # Checking to make sure it makes sense 

pop_intersect <- terra::intersect(count_rast, ocean_fact_reproj) # intersecting the two rasters to obtain TRUE where they overlap and FALSE where they do not 

plot(pop_intersect) # Making sure that I know where R thinks the two overlap. This makes sense!! 

names(ocean_fact_reproj) 
# [1] "ocean_plus25mile_inland"
names(ocean_fact_reproj) = c("zone") 
# [1] "zone"

names(count_rast) 
# [1] "gpw_v4_population_count_adjusted_to_2015_unwpp_country_totals_rev11_2020_30_sec"
names(count_rast) = c("population_count") 
# [1] "population_count"

zonal_stats <- terra::zonal(count_rast, ocean_fact_reproj, "sum", na.rm=TRUE)

zonal_filter <- zonal_stats %>% 
  mutate(zone = as.numeric(zone)) %>% 
  dplyr::filter(zone <= 250)

zonal_nas <- zonal_filter %>% 
  filter(population_count %in% ("NaN"))

count_int <- terra::as.int(count_rast)

zonal_stats2 <- terra::zonal(count_int, ocean_fact_reproj, "sum", na.rm=TRUE)

zonal_filter2 <- zonal_stats2 %>% 
  mutate(zone = as.numeric(zone)) %>% 
  dplyr::filter(zone <= 250)

zonal_nas2 <- zonal_filter2 %>% 
  filter(population_count %in% ("NaN"))
```

```{r}
## Pulling from "files"

raster_2000 <- terra::rast(files[1])

change_1 <- terra::rast(files[30])

i=1
yr = 2000+i
raster_2001 <- raster_2000 + (i * change_1)

terra::writeRaster(raster_2001, filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/human_count_%s_mol.tif", scenario_yr, yr)), overwrite = TRUE)

i=2
yr=2000+i
raster_2002 <- raster_2000 + (i*change_1)

terra::writeRaster(raster_2002, filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/human_count_%s_mol.tif", scenario_yr, yr)), overwrite = TRUE)

i=3
yr=2000+i

raster_2003 <- raster_2000 + (i*change_1)

terra::writeRaster(raster_2003, filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/human_count_%s_mol.tif", scenario_yr, yr)), overwrite = TRUE)

i=4
yr=2000+i

raster_2004 <- raster_2000 + (i*change_1)

terra::writeRaster(raster_2004, filename = file.path(dir_M, sprintf("git-annex/globalprep/mar_prs_population/%s/human_count_%s_mol.tif", scenario_yr, yr)), overwrite = TRUE)

raster_2005 <- files[grep(sprintf("count_%s_mol", 2005), files)]

```


