---
title: 'OHI 2021 - Mariculture Population Pressure Layers'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

[REFERENCE RMD FILE](https://ohi-science.org/ohiprep_v2021/globalprep/mar_prs_population/v2021/mar_prs_pop_dataprep.html)


# Summary

This document describes the steps for obtaining and wrangling the data used to calculate the mariculture population pressure sub-dimension for the 2021 global assessment. The general data preparation calculations are summarized [here](http://ohi-science.org/ohi-global/layers.html#inland_coastal_population). For context and explanation see the mariculture (subgoal of food provision) [model summary](http://ohi-science.org/ohi-global/goals.html#food_provision:_mariculture).

# Updates from previous assessment
UN population data were not updated since 2021. CIES data were not updated but the data obtained was changed from population_density .tif files to population_count .tif files. 

* We are no longer reprojecting and resampling population data to World Mollweide projection, as that affected the population data for "border" regions (i.e. those on the border of where the projection wraps around) far too much. 
  
  * We have instead modified the eez_25mi_inland raster to WGS 1984 projection using "nearest neighbor" method for resampling and not modified the population rasters. 

* In addition, because the previous script used the raster package, this new script has been changed to utilize the terra package instead. 
  
  * Some processes are less iterative in the terra package so must be done in sequence, however the processes that can function iteratively have been made so. 


***

# Data Sources

## Gridded Population of the World (v4.11) by CIESE and CIAT

**Reference**: http://sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev11/data-download

Center for International Earth Science Information Network - CIESIN - Columbia University. 2018. Gridded Population of the World, Version 4 (GPWv4): Population Density Adjusted to Match 2015 Revision UN WPP Country Totals, Revision 11. Palisades, NY: NASA Socioeconomic Data and Applications Center (SEDAC). https://doi.org/10.7927/H4F47M65. Accessed 17 April 2019

**Downloaded**: 17 April 2019

**Description**:
The Gridded Population of the World, Version 4 (GPWv4): Population Density Adjusted to Match 2015 Revision of UN WPP Country Totals, Revision 11 consists of estimates of human population density (number of persons per square kilometer) based on counts consistent with national censuses and population registers with respect to relative spatial distribution, but adjusted to match the 2015 Revision of the United Nation's World Population Prospects (UN WPP) country totals, for the years 2000, 2005, 2011, 2015, and 2020. A proportional allocation gridding algorithm, utilizing approximately 13.5 million national and sub-national administrative units, was used to assign UN WPP-adjusted population counts to 30 arc-second grid cells. The density rasters were created by dividing the UN WPP-adjusted population count raster for a given target year by the land area raster.  Documentation for gridded population of the world is located [here](http://sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev11/docs).

**Native data resolution**: 30 arc-seconds

**Time range**: 2000-2020

**Format**: GeoTiff

## UN Population

UN population data is used in this data prep routine to check population counts; spatially distributed counts derived from the gridded world population are aggregated by region and checked against the UN population estimates.

**Reference**: https://esa.un.org/unpd/wpp/

**Downloaded**: 26 June 2020

**Description**: Population (in thousands) for countries.

**Native data resolution**: Country scores

**Time range**: 1950-2020

**Format**: Excel file

**Data cleaning process**: Values from the "ESTIMATES" tab of "WPP2019_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx" were copied for all countries and years (1950-2020) and pasted into new "UN_pop_clean_v2020.csv".

***

# Setup

Load all relevant libraries including parallel processing packages
```{r setup, message=FALSE, warning=FALSE, verbose=FALSE, eval=FALSE}

## Set options for all chunks in code
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 6, fig.height = 4, fig.path = "figs/")

## Load packages, installing them first where necessary
pkg <- c("raster", "rgdal", "sp", "sf", "fasterize", "tidyverse", "foreach", "parallel","doParallel")
new_pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new_pkg)){install.packages(new_pkg)}
if (!("ohicore" %in% installed.packages())){devtools::install_github("ohi-science/ohicore")}

library(ohicore)

## Spatial libraries
# library(raster)
# # library(rgdal)
# library(sp)
# library(sf)
# library(fasterize)

## Data wrangling libraries
library(tidyverse)
library(here)
library(plotly)

## Parallel processing libraries
library(parallel) 
library(foreach)
library(doParallel)

```

Define frequently used pathnames. Change scenario and data years in file pathnames code chunk to reflect the most recent data (d) and current assessment year (v).
```{r file paths, eval=FALSE}

## Source common and spatial common files
source('http://ohi-science.org/ohiprep_v2020/workflow/R/common.R')

## Update these!
scenario_yr <- "v2020" # change to reflect assessment year!
data_yr_gpw <- "d2019" # change to reflect year of most recently downloaded Gridded Population of the World (GPW) data! (no change for v2020)
data_yr_un_pop <- "d2020" # change to reflect year of most recently downloaded UN population data! (updated for v2020)

## Define commonly used file paths (matched to what we did in ico_data_prep.Rmd)
dir_server <- file.path(dir_M, "git-annex/globalprep/_raw_data")
dir_github <- here("/globalprep/mar_prs_population", scenario_yr)

## Checking to see if there is a README in the mar_prs_pop and mar_prs_pop/v20?? folders - all goal prep files will need a README!
if(!file.exists(file.path(dir_github, 'README.md'))) {
  warning(sprintf('No README detected in %s', dir_github))
}
if(!file.exists(file.path(dirname(dir_github), 'README.md'))) {
  warning(sprintf('No README detected in %s', file.path(dirname(dir_github))))
}

```



```{r}
# Load population count raster files list to run with terra::rast (creates SpatRaster with multiple layers)
files <- list.files(file.path(dir_M, "git-annex/globalprep/mar_prs_population/v2024/int", pattern  ), pattern = "count_")
files # Make sure that they are the correct files

# Create a raster with (should be) 21 layers, one for each year plus the year 2000.
pop_stack <- terra::rast(c(files))

# Load Eez WGS 1984 Raster 
  ## Ayo make sure to run through the other script and write out the WGS 1984 raster so that the next person can just load it from THIS FILEPATH!! 
eez_25mi_inland_wgs1984 <- file.path(dir_M, "git-annex/globalprep/spatial/v2024/ocean_plus25mile_inland_wgs1984.tif")

oceans_rast <- terra::rast(eez_25mi_inland_wgs1984)

zonal_stats <- terra::zonal(x = pop_stack, z = oceans_rast, fun="sum", na.rm=TRUE)

```

