---
title: "OHI `r format(Sys.Date(), '%Y')` - CF Data Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# CF Livelihoods: Proportion of Tourism Jobs per Country/Region per Year

## Overview

* **Data:** Labor Force & Employment Data

  * Labor Force data from World Bank (downloaded June 28. 2024)

      - https://data.worldbank.org/indicator/SL.TLF.TOTL.IN
      
  * OECD
  
  * FAO Yearbook
      
  * ILOSTAT (downloaded July 2, 2024) -- for cf wage data
  
      - https://rshiny.ilo.org/dataexplorer46/?lang=en&id=EAR_4MTH_SEX_ECO_CUR_NB_A 


### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo, # for gapfilling
  pdftools
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# ILO raw data directory
ilo_dir <- here(raw_data_dir, "ILOSTAT", data_dir_version_year)

# FAO raw data directory
fao_dir <- here(raw_data_dir, "FAO", data_dir_version_year)

# output data dir for intermediate data products
int_dir <- here(data_path, "int")
```


## Read in Data

```{r}
#----------OHI regions data--------------------
# read in OHI regions for joining
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  janitor::clean_names() %>% 
  select(-notes)

############################################
# Livelihoods: Number of fisher jobs
############################################

#----------OECD data---------------------------
oecd_raw <- read_csv(file.path(dir_M, "/git-annex/globalprep/_raw_data/OECD/d2023/ocean_economy.csv")) %>% 
  select(Country, Year, VARIABLE, Variable, Unit, Value, PowerCode) 

#----------cf OECD number of fishers data------
#note that this is data for the same category as the FAO fishers data, FAO data will be used to gapfill
oecd_fisher_jobs <- oecd_raw %>% 
  filter(VARIABLE == "FISH_EMP_MARINE_FISH_TOT")  %>% 
  select(- c(PowerCode, VARIABLE)) %>% 
  filter(!Country == "OECD - Total") %>% 
  clean_names() %>% 
  mutate(unit = ("persons (1)"),
         value = (value * 1000),
         sector = "cf",
         data_source = "oecd sustainable economies") %>% 
  select(country, year, value, unit, sector, data_source)

#------cf FAO Yearbook Number of Fishers-------
# turn the number of fishers table from the FAO yearbook into a data frame and clean.
# note: world total includes countries and regions not within the table.  Unsure of where that data comes from; look for methodology.
fao_pdf <- pdf_text(here(fao_dir, "fao_statistics_yearbook_07_03_24.pdf"))

fishers_table <- fao_pdf[[169]] %>% #page number of the pdf
  as.data.frame()

#specify columns 
columns <- c("country", "1995", "2000", "2005", "2010", "2015", "2016", "2017", "2018", "2019", "2020","2021")

#turn the pdf page into a clean dataframe
fao_fisher_jobs <- fishers_table %>%
  separate_rows('.', sep = "\n") %>% #separate rows based on line breaks
  rename(main_column = ".") %>%
  mutate(main_column = str_remove_all(main_column, "\\*")) %>% #remove the * 
  mutate(main_column = str_replace_all(main_column, "E(?!\\p{L})", " ")) %>% #replace all the Es not followed by letters (used to indicate estimate)
  slice(7:69) %>% #remove all rows not in the table
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% #separate into columns based on double spaces
  pivot_longer(cols = -c("country"), names_to = "year") %>% 
  mutate(value = str_squish(value)) %>% #remove extra spaces
  mutate(value = str_remove_all(value, " ")) %>% 
  mutate(value = if_else(value == "…" , NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(unit = "persons (1)",
         sector = "cf",
         data_source = "FAO number of fishers") %>% 
  slice(-c(1:44))

############################################
# Livelihoods: Quality of fisher jobs
############################################

#-------cf ILO wage data (2009 - 2021) -------
# this data has limited entries, so local currency will need to be ppp adjusted to increase the number of observations per geo area
cf_ilo_wage_raw <- read_csv(here(ilo_dir, "EAR_4MTH_SEX_ECO_CUR_NB_A-filtered-2024-07-03-cf-wage.csv"))
```

#### FAO Yearbook Number of Fishers

Turn the number of fishers table from the FAO yearbook into a data frame
and clean.

```{r, eval = FALSE}

```

### Read in Data

```{r}
# ===================== Read in Data ==============================


# Labor force data ----
labor_raw <- readxl::read_xls(here(wb_dir, "worldbank_labor_force_raw.xls"),
                               skip = 3,
                               na = "")

# OHI regions data ----
# read in OHI regions for joining
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") 


# ILO wage data ----
# define file path to ILO wage data 
ilo_fp <- file.path(dir_M,
                        "git-annex/globalprep/_raw_data/ILOSTAT/d2024/EAR_4MTH_SEX_ECO_CUR_NB_A-filtered-2024-07-02-tour-wage.csv")

# read in ILO wage data
ilo_wage_data <- read_csv(ilo_fp)

```

# Livelihood: Number of Jobs -- cf

(put this within previous chunk, read in data etc. for number of tourism jobs, join with labor_regions, make new proportion of tourism jobs column, group by year, do prelim plotting etc. and evaluate data quality etc.)

Cleaning and wrangling data
```{r}
# total number of people employed by the fishing (previously commercial fishing, cf -- see README.md) sector

```

Create the proportion between tourism jobs and the total labor force
```{r}

```

Saving intermediate data
```{r}

```


# Livelihood: Quality of Jobs -- cf

  - note: this data came from many sources and was compiled by ILO

Cleaning and wrangling data
```{r}


```


Gapfilling if necessary
```{r}

```


```{r}
# find regions with only 1 data point (still have NAs in appx_wage column)

```


```{r}
# # preliminary plot 
# line_plot <- plotly::plot_ly(wage_gf, x = ~year, y = ~appx_wage_fill, color = ~ref_area_label, type = "scatter", mode = "lines") %>% 
#   layout(title = "All Regions: monthly Service Wages (USD (2017) PPP-Adjusted)", 
#          xaxis = list(title = "Year"),
#          yaxis = list(title = "monthly Wages (USD)"))
# 
# line_plot
# 
# # htmlwidgets::saveWidget(line_plot, file = "prop_tourism_laborforce.html")
```


Clean up ILO data, join with OHI regions
```{r}
# join with OHI regions

```


Plot
```{r}
# # interactive plot after joining with OHI Regions
# plotly::plot_ly(wage_regions, x = ~year, y = ~appx_wage_fill, color = ~admin_country_name, type = "scatter", mode = "lines") %>% 
#   layout(title = "All Regions: Monthly Service Wages (USD)", 
#          xaxis = list(title = "Year"),
#          yaxis = list(title = "Monthly Wages (USD)"))

```



