---
title: "OHI `r format(Sys.Date(), '%Y')` - CF ECO Revenue Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

# CF Economies: USD Value of Marine (Commercial) Fishing per Country/Region per Year

## Overview

-   FAO Capture Data 

    Data Source FAO Global Capture Production (in metric tonnes)

    This version of the value database was downloaded from the
    Statistical Query Panel. Data from [FAO Global Capture
    Production](https://www.fao.org/fishery/en/collection/capture?lang=en)

    Citation: FAO 2023. Global Capture Production. Fisheries and
    Aquaculture Division
    <https://www.fao.org/fishery/en/collection/capture?lang=en>

    Source information: Navigate to the [online query
    portal](https://www.fao.org/fishery/statistics-query/en/capture/capture_quantity)
    for FAO Global Capture Production Quantity. Deselect all
    pre-selected years. Drag these fields into selected rows: ASFIS
    species name, FAO major fishing area name, ASFIS species ISSCAP
    group name En. ASFIS species Family scientific name, FAO major
    fishing areas, Inland/Marine areas Name en. Click on show data and
    confirm that data is present for 1950- two years prior to current
    year. Click download and select yes to include Null Values.

    Date: September 15th, 2023

    Time range: 1950-2021

    Native data resolution: Country level

    Format: csv

    Description: Global Capture Production Quantity

-   Ex-Vessel Price Data (downloaded August 24, 2023)

    -   Ex-vessel price data is in USD/metric tonne.

        -   ex-vessel-prices: ex-vessel prices for fishery caught
            species from 1976-2019

            -   exvessel_price_database_1976_2019.csv: ex-vessel price
                data gathered from [Melnychuk et al.
                2016](https://doi.org/10.1093/icesjms/fsw169) and
                updated to 2019 using methods described in the
                public-facing [github
                repo](https://github.com/SFG-UCSB/price-db-sfg)
                associated with the Melnychuk et al. 2016 paper

        -   \*\*Citation for paper\*\*  

        -   Melnychuk, M. C., Clavelle, T., Owashi, B., and
            Strauss, K. 2016. Reconstruction of global ex-vessel prices
            of fished species. - ICES Journal of Marine Science.
            doi:10.1093/icesjms/fsw169.
            
            
### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo, # for gapfilling
  pdftools
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

source(here('workflow/R/fao_fxn.R')) # function for cleaning old version of FAO data 

source(here('workflow/R/fao_online_portal_clean.R'))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# ILO raw data directory
ilo_dir <- here(raw_data_dir, "ILOSTAT", data_dir_version_year)

# FAO raw data directory
fao_dir <- here(raw_data_dir, "FAO", data_dir_version_year)

# OECD raw data directory
oecd_dir <- here(raw_data_dir, "OECD", data_dir_version_year)

# FAO capture raw data directory
fao_capture_dir <- here(raw_data_dir, "FAO_capture", data_dir_version_year)

# Ex-vessel raw data directory
ex_vess_dir <- here(raw_data_dir, "ex-vessel-price-database-updated")

# output data dir for intermediate data products
int_dir <- here(data_path, "int")
```

### Read in Data

```{r}
fao_capture_raw <- read_csv(here(fao_capture_dir, "Global_capture_production_Quantity_1950-2021.csv"))

exvessel_prices_raw <- read_csv(here(ex_vess_dir, "price-db-results", "exvessel_price_database_1976_2019.csv"))

# read in cf revenue data from v2023, using FAO Capture data and Ex-vessel data
cf_revenue <- read_csv(here("globalprep", "le", "v2023", "int", "cf_revenue.csv"))
```

### Run ohicore name_2_rgn
```{r}
# check names
# These data were removed for not having any match in the lookup tables:
# 
# bonaire/s.eustatius/saba (need to be disaggregated) british indian ocean ter (DNE)    channel islands (split by eez area: Jersey, larger island)    french southern terr (DNE)
#                        1                        1                        1                        1 
#.   isle of man (cannot be lumped w Britain, maybe Ireland)  netherlands antilles  (curacao)      other nei (DNE)        saint barthélemy (DNE)
#                        1                        1                        1                        1 
# saint helena/asc./trist. (Saint Helena)   saint-martin (french)  (DNE)         yugoslavia sfr (DNE)
#                        1                        1                        1 
# 
# These data were removed for not being of the proper rgn_type (eez,ohi_region) or mismatching region names in the lookup tables:
#            tmp_type
# tmp_name    disputed landlocked
#   ethiopia         0         16
#   palestine       25          0

length(unique(cf_revenue$country)) #203

cf_revenue_fix <- cf_revenue %>%
  mutate(country = case_when(
    country=="Bonaire/S.Eustatius/Saba" ~ "Bonaire", # how to disaggregate instead of aggregate?
    country=="British Indian Ocean Ter" ~ "British Indian Ocean Territory",
    country=="Channel Islands" ~ "Jersey", # would be disaggregated into Guernsey and Jersey
    country=="Saint Helena/Asc./Trist." ~ "Saint Helena", 
    country=="Saint-Martin (French)" ~ "Northern Saint-Martin",
    country=="Netherlands Antilles" ~ "Curacao",
    country=="Puerto Rico" ~ "Puerto Rico and Virgin Islands of the United States",
    country=="China, Hong Kong SAR" ~ "China",
    country=="China, Macao SAR" ~ "China",
    country=="United Republic of Tanzania, Zanzibar" ~ "Tanzania",
    country=="Tanzania, United Rep. of" ~ "Tanzania",
    TRUE ~ country # Everything else, leave it be
  ))
```

## Disaggregation 

Look at rasterized fishing effort, to determine the relative proportion of it within each eez, and distribute catch relative to that proportion.

**Regions to disaggregate:**

Channel Islands --> Guernsey (228) and Jersey (227)

Bonaire/S.Eustatius/Saba --> Bonaire (245), Sint Eustatius (249), and Saba (248)


```{r}
# Function to calculate average catch for certain regions and certain years
calculate_avg_catch <- function(data, region_ids, years) {
  data %>% 
    filter(rgn_id %in% region_ids, year %in% years) %>% 
    group_by(rgn_id, year) %>% 
    summarize(catch = sum(mean_catch, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(rgn_id) %>% 
    mutate(avg_catch = mean(catch)) %>% 
    select(rgn_id, avg_catch) %>%
    distinct() %>% # ensures that there's only one row per region, removing any potential duplicates
    pivot_wider(names_from = "rgn_id", values_from = "avg_catch")
}

# Function to calculate the proportion of fishing done for each region out of the total (aggregated) area
# it sums up whatever is in that column. data[, -1] selects all rows (because there's nothing before the comma) and all columns except the first one (because of the -1 after the comma).
calculate_proportion <- function(data, column) {
  data[[column]] / sum(data[, -1], na.rm = TRUE)
} 

# Load in data on the mean catch for each region, used as a proxy to determine which region has a greater commercial fishing economy
mean_catch_2017 <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2024/gh-pages/globalprep/fis/v2020/int/mean_catch.csv")

# Define parameters to input into the functions
years_to_analyze <- 2014:2017
channel_islands_ids <- c("228", "227")
bes_islands_ids <- c("245", "249", "248")

# Calculate average catch for Channel Islands for the 3 most recent years
mean_catch_ci_agg <- calculate_avg_catch(mean_catch_2017, channel_islands_ids, years_to_analyze) %>%
  rename(jersey = `227`, guernsey = `228`)

# Calculate average catch for Bonaire/S.Eustatius/Saba
mean_catch_bes_agg <- calculate_avg_catch(mean_catch_2017, bes_islands_ids, years_to_analyze) %>%
  rename(bonaire = `245`, sint_eustatius = `249`, saba = `248`)

# Calculate proportions for each disaggregated region
prop_g <- calculate_proportion(mean_catch_ci_agg, "guernsey")
prop_j <- calculate_proportion(mean_catch_ci_agg, "jersey")

prop_b <- calculate_proportion(mean_catch_bes_agg, "bonaire")
prop_e <- calculate_proportion(mean_catch_bes_agg, "sint_eustatius")
prop_s <- calculate_proportion(mean_catch_bes_agg, "saba")
```

```{r}
# use proportions to disaggregate Channel Islands and Bonaire/S.Eustatius/Saba
split_ci <- cf_revenue %>%
  filter(country == "Channel Islands") %>%
  mutate(split_col = "Guernsey;Jersey") %>%
  separate_rows(split_col, sep = ";") %>%
  group_by(country, year, value, unit) %>%
  case_when(
    split_col == "Guernsey" ~ value*prop_g,
    split_col == "Jersey" ~ value*prop_j
  ) %>% 
  ungroup()

sum(split_ci$value) 

```


```{r}
cf_revenue_rgn <- name_2_rgn(df_in = cf_revenue_fix, 
                             fld_name='country',
                             flds_unique = c("year"))

cf_revenue_rgn_agg <- cf_revenue_rgn %>% 
  select(rgn_id, rgn_name, year, value, unit, sector, data_source) %>% 
  group_by(rgn_id, rgn_name, year) %>% # because china needs to be aggregated and has duplicate years
  dplyr::summarize(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>% 
  dplyr::mutate(year = as.numeric(year)) 

rev_duplicates <- cf_revenue_rgn[duplicated(cf_revenue_rgn[, c("country", "year")]),]
rev_ref_duplicates <- unique(rev_duplicates$country)
rev_rgn_duplicates <- unique(rev_duplicates$rgn_name)
rev_diff <- setdiff(rev_rgn_duplicates, rev_ref_duplicates)
rev_diff # none! Move on
```

## CF Revenue Plot
```{r}
cf_rev_plot <- plotly::plot_ly(cf_revenue_rgn_agg, x = ~year, y = ~value, color = ~rgn_name, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Yearly Revenue for the Marine Fishing Sector", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Revenue of all catch (USD)"))
cf_rev_plot

cf_revenue_china <- cf_revenue_rgn_agg %>% 
  filter(rgn_name %in% "China") #before aggregation, 

plotly::plot_ly(cf_revenue_china, x = ~year, y = ~value, color = ~rgn_name, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Yearly Revenue for the Marine Fishing Sector", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Revenue of all catch (USD)"))
```

## Saving intermediate data
```{r}
# columns should be: rgn_id, rgn_name, year, usd (previously value), unit, sector, usd_yr (in this case = year)

cf_revenue_int <- cf_revenue_rgn_agg %>% 
  mutate(usd = value,
         unit = "USD (1)",
         sector = "cf",
         usd_yr = year) %>% 
  select(-value)

# write_csv(cf_revenue_int, here(int_dir, "eco_cf_usd_pre.csv"))
```


