---
title: "OHI Economies Dataprep"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------


# Economies Dataprep (Measures the amount of revenue produced)

## Overview

-   **Data:**

    -   Tourism: UNWTO Dept. of Statistics (UN Tourism)
        -   <https://www.unwto.org/tourism-statistics/economic-contribution-SDG> 
        
    -  GDP: World Bank, in "Constant 2015 US$"
      -   <https://data.worldbank.org/indicator/NY.GDP.MKTP.KD> Downloaded July 9, 2024. Has annual GDP value from 1960-2023. Additional metadata copied and saved as a `.rtf` in the WorldBank_global_annual_GDP_2015_constant_USD subfolder of the _raw_data WorldBank folder for d2024. Last updated (found in the first few rows of the `.csv`): 2024-06-28

### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  countrycodes,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo # for gapfilling
  
)
# remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO raw data directory
unwto_dir <- here(raw_data_dir, "UNWTO", data_dir_version_year)

# World Bank raw data dir (for GDP data)
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# Output data directory -- for intermediate data products
int_dir <- here(data_path, "int")
```


### Tourism

#### Read in Data

```{r}
# ===================== Read in Data ==============================


# Tourism revenue data ----
gdp_prop_raw <- readxl::read_xlsx(here(unwto_dir, "Indicator-8_9_1-2022-UN_Tourism_april2024_update.xlsx"))

# OHI regions data ----
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  select(-Notes)

# World Bank GDP in 2015 Constant USD data ----
wb_gdp_raw <- read_csv(here(wb_dir, "WorldBank_global_annual_GDP_2015_constant_USD", "WorldBank_global_annual_GDP_2015_constant_USD.csv"), skip = 4)
# date range from 1960 to 2023

```


## Tourism GDP %: Data cleaning

Current unit of "total" in dataframe:
"Tourism direct GDP as a proportion of total GDP (%)" 

```{r}
# ======= GDP proportion dataset cleaning ========
# preliminary cleaning/tidying ----
gdp_prop_clean <- gdp_prop_raw %>% 
  janitor::clean_names() %>% 
  select(c(geo_area_name, year = time_period, gdp_prop_percent = total))


# run name_2_rgn -----
gdp_rgn <- name_2_rgn(df_in = gdp_prop_clean, 
                       fld_name = 'geo_area_name',
                       flds_unique = c("year"))

# sorting through mismatch and aggregation ----

# note on Northern Mariana Islands and Guam OHI region: we only have data for Guam, not Northern Mariana Islands... this is problematic because in the World Bank GDP dataset, we only have GDP data for both the Northern Mariana Islands and Guam and need to aggregate them to fit into the OHI regions system. For now, we'll drop the Northern Mariana Islands GDP data in the WB GDP processing, but please note that if you get GDP % data for this region later on, you'll need to fix that in the GDP data processing chunk.

# China: only have data hong kong and macao, no mainland China data...
# Mel told us to just drop them... 

# check duplicates (none)
tour_eco_duplicates <- gdp_rgn[duplicated(gdp_rgn[, c("rgn_name", "year")]),]
unique(tour_eco_duplicates$geo_area_name)
unique(tour_eco_duplicates$rgn_name)

length(unique(gdp_rgn$rgn_name))
# v2024: 93

```


```{r}
# preliminary plotting ----
prelim_plot <- plot_ly(data = gdp_prop_clean, x = ~year, y = ~gdp_prop_percent, color = ~geo_area_name, 
                       type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for All Geo Regions",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

prelim_plot

# htmlwidgets::saveWidget(prelim_plot, file = "gdp_prelim_plot_all_years.html")


plot_ly(data = gdp_rgn, x = ~year, y = ~gdp_prop_percent, color = ~rgn_name, 
                       type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for OHI Regions",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))
```

Really interesting to see how the proportions drop in 202 -- intuition check: makes sense because of Covid.... especially interesting to note how extremely Macau was impacted!


### Gapfilling

More data cleaning & prep for gapfilling, then gapfilling

```{r}
# define maximum year (determined after visually inspecting data)

# filter tourism GDP data for year range (2009-2019) ----
gdp_prop_filter <- gdp_rgn %>% 
  filter(year %in% c(2009:2019))


# some geo areas are missing data for some years ----
years_df <- tibble(rgn_name = region_names$rgn_name,
                   rgn_id = region_names$rgn_id) %>% # copy region identifier columns
  group_by(rgn_name, rgn_id) %>% # for each geo region, 
  summarize(year = seq(2009, 2019)) # have sequence of years from 2009-2019 
# (fragile and R warns you about preferring `reframe()`)


# join years df and tourism GDP df ----
gdp_year_join <- full_join(years_df, gdp_prop_filter, by = c("rgn_name", "rgn_id", "year"))


# tidy or fill in geo area names for NA years ----
gdp_years <- gdp_year_join %>% 
  select(-geo_area_name) # %>%  # alternatively, set years_df to have geo_area_name column, then perform the following steps to fill in geo area names for NA years
  # group_by(rgn_name) %>% 
  # fill(geo_area_name, .direction = "downup") %>% 
  # ungroup()


# gapfilling ----
# estimate tourism GDP proportion for geo areas with missing data
gdp_filled <- gdp_years %>% 
  group_by(rgn_name) %>% 
  # interpolate (fill missing values between 2 values)
  mutate(appx_gdp_pct = zoo::na.approx(gdp_prop_percent, # using values in this column
                                    na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                    #  extrapolate using rule = 2 from approx(),
                                    # which uses closest data extreme to
                                    #  extrapolate for leading and trailing NAs
                                    rule = 2
         )) # %>% 
  
  # fill NA values with the only other observed value (making sure to note the list of countries/regions we did this for) -- ended up not using this column
  # mutate(appx_gdp_fill = zoo::na.aggregate(appx_gdp, by = 1, FUN = mean, na.rm = FALSE))
```
Note: after further discussing with Melanie Frazier, we decided to not include countries with only 1 year of observations in our dataset.

Assess how many countries we are gapfilling more extremely for

```{r}
# no longer useful, as we cut out this column to avoid over-gapfilling
# # find number of regions that only have 1 data point (wouldn't be filled by na.approx if only 1 datapoint, left as NAs)
# na_regions <- gdp_filled %>% 
#   filter(is.na(appx_gdp_fill))
# 
# # print which geo areas only had 1 datapoint
# #unique(na_regions$rgn_name)
# 
# # calculate proportion of 1 point-filled geo areas to total geo areas
# num_na <- length(unique(na_regions$rgn_name))
# num_tot <- length(unique(gdp_filled$rgn_name))
# 
# paste0("Proportion of geo areas with only 1 data point: ", round(((num_na / num_tot) * 100), 3), "%")
```

#### Preliminary Plotting

```{r}
# plot of all geographic areas ---- 
prelim_gf_plot <- plot_ly(data = gdp_filled, x = ~year, y = ~appx_gdp_fill, color = ~geo_area_name,
                          type = "scatter", mode = "lines") 

prelim_gf_plot

# top_tourism_countries <- gdp_filled %>% 
#   group_by(geo_area_name, year) %>% 
#   summarize()


# subset without Macao for plotting ----
gdp_no_macau <- gdp_filled %>% 
  filter(!geo_area_name %in% c("China, Macao Special Administrative Region"))

plot_ly(data = gdp_no_macau, x = ~year, y = ~appx_gdp_fill, color = ~geo_area_name,
                          type = "scatter", mode = "lines") 


# make subset of geo areas with greatest GDP change ----
gdp_diff <- gdp_filled %>% 
  group_by(geo_area_name) %>% 
  summarize(gdp_diff = (abs(max(appx_gdp_fill) - min(appx_gdp_fill)))) %>% 
  arrange(desc(gdp_diff)) %>% 
  mutate(geo_area_name = str_replace_all(geo_area_name,
                                         pattern = "China, Macao Special Administrative Region",
                                         replacement = "Macao"))

gdp_diff

gdp_top_diff <- gdp_diff %>% 
  slice_head(n = 15)

top_diff_areas <- c(gdp_top_diff$geo_area_name)

gdp_filled_diff <- gdp_filled %>% 
  mutate(geo_area_name = str_replace_all(geo_area_name,
                                         pattern = "China, Macao Special Administrative Region",
                                         replacement = "Macao")) %>% 
  filter(geo_area_name %in% c(top_diff_areas))

m <- list(
  l = 80,
  r = 80,
  b = 80,
  t = 80,
  pad = 4
)

top_diff_gdp_plot <- plot_ly(data = gdp_filled_diff, x = ~year, y = ~appx_gdp_fill, color = ~geo_area_name,
                          type = "scatter", mode = "lines") %>% 
  layout(title = "Geo Areas with the Greatest Difference in Tourism GDP (2009-2019)",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%, Gapfilled)")
         ) 

top_diff_gdp_plot


# htmlwidgets::saveWidget(top_diff_gdp_plot, file = "tour_gdp_top_diff_gf_plot.html")
```



## GDP: WorldBank data cleaning



```{r}
# ======== GDP data cleaning ==================

wb_gdp_clean <- wb_gdp_raw %>% 
  clean_names() %>% 
  # remove unwanted columns
  select(-c(indicator_code, x69)) %>% 
  # pivot to put years in 1 column, values in another
  pivot_longer(cols = x1960:x2023, names_to = "year", values_to = "gdp_value") %>% 
  # tidy up year column
  mutate(year = as.numeric(str_remove_all(year, 'x'))) %>%  
  # filter to date range
  filter(year %in% c(2009:2019))


# run name_2_rgn -----
wb_gdp_rgn <- name_2_rgn(df_in = wb_gdp_clean, 
                       fld_name = "country_name",
                       flds_unique = c("year"))

# thankfully, in the list following this line: "These data were removed for not having any match in the lookup tables:", there are mostly just the WorldBank aggregate groups (i.e., "low & middle income, Latin America & Caribbean, etc.), with the exception of Channel Islands and Isle of Man

# COME BACK ---- split Channel Islands -- proportional to EEZ area, Jersey slightly larger than Guernsey
# Isle of Man -- distinct governing body, should not be aggregated with Ireland or the UK

# identify duplicates ----
wb_gdp_duplicates <- wb_gdp_rgn[duplicated(wb_gdp_rgn[, c("rgn_name", "year")]),]
View(wb_gdp_duplicates)
unique(wb_gdp_duplicates$country_name)
unique(wb_gdp_duplicates$rgn_name)

# then look up these & their pairs in the wb_gdp_clean df to see which countries we do or do not have data for to inform aggregation steps.

# [1] "China": we have data on Macao and Hong Kong --> aggregate (can sum) -- but we don't have tourism GDP % data for mainland China... so just leave as basic aggregation (sum) for now


# [2] "Northern Mariana Islands and Guam" --> have GDP data for Northern Mariana Islands, and Guam. Can aggregate this, but we only have tourism % GDP data for Guam (see earlier comment in GDP % data cleaning chunk). 
# For now, we'll just keep Guam GDP and filter out northern mariana islands GDP... ? (COME BACK: double check w Mel)


# [3] "Puerto Rico and Virgin Islands of the United States" --> have data for both separately --> aggregate



# aggregate: sum by rgn_name etc. ----
# name ~ WorldBank GDP region aggregated
wb_gdp_rgn_agg <- wb_gdp_rgn %>%
  filter(!country_name %in% c("Northern Mariana Islands")) %>% # filtering out NMI since we only have Guam GDP % data in our other dataset -- should revisit this and update once we have that data -- double check to see which of these regions we have data for on both datasets in future years of updates!!!
  group_by(rgn_id, rgn_name, year) %>%
  dplyr::summarize(gdp_value_agg = sum(gdp_value, na.rm = TRUE) 
            )%>%
  ungroup() %>% 
  # turn 0s into NAs (appropriate flag for lack of data, as earlier sum(na.rm = TRUE) step obscured NAs)
  mutate(gdp_value_agg = case_when(
    gdp_value_agg == 0 ~ NA,
    .default = gdp_value_agg
  ))

# aggregation check ----
# to double check and see if Northern Marian Islands was properly filtered out:
agg_nmi_vals <- wb_gdp_rgn_agg %>% 
  filter(rgn_name %in% c("Northern Mariana Islands and Guam")) %>%
  select(year, gdp_value_agg)

guam_reference <- wb_gdp_clean %>% 
  filter(country_name %in% c("Guam")) %>% 
  select(year, gdp_value)

agg_nmi_check <- left_join(agg_nmi_vals, guam_reference)

print(paste("Do the Guam GDP values before aggregation match the 'Northern Mariana Islands and Guam' values after aggregation?"))

if (sum(agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value) == length(unique(agg_nmi_check$gdp_value_agg))
) {
  print("Yes")
} else {
  print("No")
}

# If "No", check which don't match
# print(paste((agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value)))

```


Extra: aggregation explanation/example
```{r agg_example, eval = FALSE}
# to see why we're using summarize instead of just mutating and adding a column see this example:
# test <- wb_gdp_rgn %>%
#   filter(!rgn_name %in% c("Northern Mariana Island")) %>%
#   group_by(rgn_id, rgn_name, year) %>%
#   mutate(test = sum(gdp_value, na.rm = TRUE))
# 
# #View(test)
# test %>% filter(rgn_name %in% c("China"))

# yields value column, but still keeps the old country_name column, so there are still duplicates! 

```


## Find Tourism Value 


```{r}
# join GDP Value (constant 2015 US$) with Tourism GDP % ----
gdp_join <- full_join(gdp_filled, wb_gdp_rgn_agg, by = join_by(rgn_name, rgn_id, year))

# gdp_join_rgn <- name_2_rgn(df_in = gdp_join, 
#                        fld_name = "rgn_name",
#                        flds_unique = c("year"))

# calculate tourism value ---- 
tourism_value <- gdp_join %>% 
  # get proportion to use as multiplier
  mutate(gdp_filled_prop = appx_gdp_pct / 100) %>% 
  ungroup() %>% 
  group_by(rgn_id, rgn_name, year) %>% 
  mutate(tour_value = gdp_filled_prop * gdp_value_agg) %>% 
  # fix data types
  mutate(rgn_id = as.character(rgn_id))


# preliminary plotting ----
# make subset
tour_val_subset <- tourism_value %>% 
  filter(!is.na(tour_value))

# preliminary plot to check this out!

ggplot(data = tour_val_subset, 
       aes(x = as.numeric(year), 
           y = tour_value,
           color = rgn_id)) +
  geom_point() +
  geom_line() +
  theme_bw()


m <- list(
  l = 80,
  r = 80,
  b = 80,
  t = 80,
  pad = 4
)

tour_val_plotly <- plot_ly(data = tour_val_subset, x = ~year, y = ~tour_value, color = ~rgn_name,
                          type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism Value (2009-2019)",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Value (2015 US$)")
         ) 
tour_val_plotly
```


```{r}
# check max value
max_tour_val <- max(tourism_value$tour_value, na.rm = TRUE)

tourism_value %>% 
  filter(tour_value == max_tour_val)

prettyNum(max_tour_val,big.mark=",",scientific=FALSE)

# 582 billion in US in 2019... out of 20 Trillion GDP
```


# Saving intermediate data 

```{r}
#--------Prepping data for int saving -------

tour_revenue_int <- tourism_value %>% 
  rename(usd = tour_value) %>% 
  mutate(
    unit = "USD (1)",
    sector = "tour",
    usd_yr = 2015 # because the tourism values are in "Constant 2015 US$"
  ) %>% 
  select(rgn_id, rgn_name, year, usd, unit, sector, usd_yr) %>% 
  ungroup()

#write_csv(tour_revenue_int, here(int_dir, "eco_tour_usd_pre.csv"))
```




### Join with OHI regions

COME BACK -- ask about joining w OHI rgns

```{r}

# tidy up regions dataset ----
region_names <- region_names_raw %>% select(-Notes)


# join regions & tourism GDP prop ----
gdp_regions_join <- left_join(region_names, gdp_filled, by = c("eez_iso3" = "iso3"))
  

head(gdp_regions_join)
str(gdp_regions_join)

```


Note: remember to add
  mutate(sector = "tour", data_source = "UNWTO Dept. of Statistics")

  at the end

