---
title: "OHI `r format(Sys.Date(), '%Y')` - Mariculture Livelihoods & Economies Data Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Livelihoods: Proportion of Tourism Jobs per Country/Region per Year

## Overview

### Mariculture Revenue Data

-   Data came from FishStatJ, FAO's application to obtain different
    fishery-related metrics by country as well as by sector
-   Citation: © FAO 2024. Global Aquaculture Production. In:
    Fisheries and Aquaculture. Rome. [Cited Tuesday, July 9th 2024].
    <https://www.fao.org/fishery/en/collection/aquaculture>

-   **Instructions for download from FishStatJ**
    -   Go to FAO website for download
        [FAO](https://www.fao.org/fishery/en/statistics/software/fishstatj)
    -   Also open the user manual found on that page, linked
        [here](https://www.fao.org/fishery/static/FishStatJ/FishStatJ_4.03.05-Manual.pdf)
    -   Once downloaded, open FishStatJ on your computer and follow the
        instructions to set it up.
    -   Then click file -\> manage workspaces -\> click 'FAO Global
        Fishery and Aquaculture Production Statistics' -\> click
        'Import' -\> Next -\> Next; until it opens
    -   This should import the workspace and allow you to access the
        'FAO Global Fishery and Aquaculture Production Value' data
    -   Once it opens, click 'File' -\> 'Export Selection (CSV File)'
    -   Store it somewhere you can find on your local drive and from
        there move it into the 'FAO_mariculture' folder under
        /home/shares/ohi/git-annex/globalprep/\_raw_data/FAO_mariculture/*your
        data year*

### Mariculture Jobs Data

-   Partially obtained from [FAO Fisheries and Aquaculture Statistical Yearbook](https://openknowledge.fao.org/server/api/core/bitstreams/2be6c2fa-07b1-429d-91c5-80d3d1af46a6/content)
-   Also brought in OECD data from their online [OECD Data Explorer](https://data-explorer.oecd.org/vis?df%5Bds%5D=DisseminateFinalDMZ&df%5Bid%5D=DSD_SOE%40DF_SOE&df%5Bag%5D=OECD.ENV.EPI&dq=.A....&pd=1995%2C2024&to%5BTIME_PERIOD%5D=false&vw=tb)


### Setup

```{r}
# load packages ---- 
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  priceR,
  pdftools,
  threadr,
  tidyverse, 
  httr,
  plotly,
  zoo # for gapfilling
  
)

# source common.R ----
source(here::here("workflow/R/common.R"))

# set year and file path info ----
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year) # Assessment year
data_dir_version_year <- paste0("d", current_year) # Data folder for assessment year
data_path <- here::here("globalprep", "le", version_year) # Path to assessment folder (v202X) 

# Raw data directory (on Mazu) ---- 
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here::here(raw_data_dir, "WorldBank", data_dir_version_year)

# FAO Mariculture raw data directory 
fao_dir <- here::here(raw_data_dir, "FAO_mariculture", data_dir_version_year)

# output data directory for intermediate data products
int_dir <- here::here(data_path, "int")
```

## Read in Data

```{r}
# OHI regions data ----
# read in OHI regions for joining
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  janitor::clean_names() %>% 
  select(-notes)

# FAO Aquaculture data ----
aquaculture_value <- read_csv(file.path(fao_dir, "/FAO_GlobalAquacultureProduction_Value_1984_2022.csv")) %>% 
  janitor::clean_names()

## str to view file structure (it can be very messy)
# str(aquaculture_value)
```

## Clean Aquaculture value data

### Note: This process was done manually because instead of reading in the data with "flag" as the header for each observation's symbol, 2024's year was read in with "symbol" as the header for each observation's symbool

-   As a result, the data cleaning function fao_online_portal_clean.R
    was not able to be run. So what I did was drag components of the
    cleaning function and manually cleaned the data. See v2023
    livelihoods_economies_dataprep.Rmd to see what they did.
    -   With the data in the correct format (drawn from online FAO
        portal), this function should apply

```{r}
#clean the aquaculture value data ----
#this will also be used in the mariculture jobs data section, so need to keep in all environments in the first step

# Cleaning aquaculture value data
aquaculture_value_clean <- aquaculture_value %>% 
   dplyr::rename(country = country_name, # Assigning relevant column names 
                species =  asfis_species_name, 
                area = fao_major_fishing_area_name,
                environment = environment_name) %>%
  dplyr::select(-c(unit_name)) %>% #clean data, note that we are still using a sub_n value of .1 (100 dollars) 
  mutate(row_id = 1:nrow(.))

# This section adjusts columns simultaneously so that you dont have to ---- 

# Check these! These are for the "year" columns
initial_data_year = 1984
last_data_year = 2022

# Testing to make sure the "year" specs follow the same naming format as the df
test = paste0("x", initial_data_year:last_data_year)

# Check these! These are for the "symbol" or "flag" columns (2024 they were in increments of 2)
flag_start = 8
flag_end = 84

# Testing to make sure the symbol or flag specs are the same naming format as the df 
test = paste0("symbol_", seq(flag_start, flag_end, by = 2))

# Pivoting columns to tidy the data ---- 

# This creates a dataframe of just the values in the FAO data 
fao_values <- aquaculture_value_clean %>% 
  dplyr::select(-c(paste0("symbol_", seq(flag_start, flag_end, by = 2)))) %>% # remove "flag" columns
  pivot_longer(cols = paste0("x", initial_data_year:last_data_year), 
               names_to = "year",
               values_to = "value") %>% # Pivot value colunns
  mutate(year = str_remove(year, "x")) # clean_names() put an 'x' in front of years. Removing the x

# This is a little bit more messy than the values because my column headers contained symbol instead of flag 
fao_flags1 <- aquaculture_value_clean %>% 
  dplyr::select(-c(paste0("x", initial_data_year:last_data_year))) # select out year

### MAYBE MOVE THIS TO APPLY TO THE AQUACULTURE_VALUE_CLEAN DF RATHER THAN FAO_FLAGS1

# Rename the columns in the Flag dataframe to actually have "flag" rather than symbol AND** getting the year for each flag so that we can rejoin the flags df with the values df 
colnames(fao_flags1) = c("country", "species", "area", "environment", "unit", paste0("flag", initial_data_year:last_data_year), "row_id")

# Because values were already selected out, this line of code is just pivoting the flags df and cleaning
fao_flags <- fao_flags1 %>% 
  pivot_longer(cols = paste0("flag", initial_data_year:last_data_year),
               names_to = "flag_year",
               values_to = "flag") %>% # Pivot longer 
  mutate(year = str_remove(flag_year, "flag")) %>% # Remove the "flag" from flag year so that we can have year to join by in the final flag df
  dplyr::select(year, row_id, flag) # selecting relevant columns to join by 

# Creating final df ---- 

# Joining fao values dataframe with the fao flags dataframe so that we can account for "N" flag and assign a value to it 
fao_new <- fao_values %>%
  left_join(fao_flags, by = c("row_id", "year")) # This is why we need year from each flag. If we just pivoted the values and then the flags in the aquaculture_values_clean df then it would double the values for each observation. 

# Assign value to substitute "N" flags for 
sub_N = 0.1

# Replace "N" flags 
fao_new <- fao_new %>% 
  mutate(value = case_when((str_detect(flag, "N") & value == 0) ~ sub_N,
                           TRUE ~value)) %>% #replace values that are 0 and have the flag N with sub_N
  select(-c(row_id, flag)) # Getting rid of temporary indexing row and flag row 


# Filter to only marine habitats
 mariculture_revenue <- fao_new %>% 
   filter(environment %in% c("Brackishwater", "Marine"))

# Sum by region and year (we need total revenue)
total_mar_rev <- mariculture_revenue %>%
  group_by(country, year) %>% 
  dplyr::summarize(value = sum(value, na.rm = TRUE)) %>% 
  mutate(sector = "mar", data_source = "FAO aquaculture value", unit = "USD (1)",
          value = value *1000) %>% # Multiply value times 1000 to get dollars
  select(country, year, value, unit, sector, data_source) # Select relevant columns

# Plotting to gut check ---- 
total_mar_rev %>% 
  # filter(!country == "China") %>% # If needed, filter out China because it tends to squish the lower values 
  ggplot(aes(x = as.numeric(year), y = value, color = country)) + 
  geom_line() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scales::label_dollar(scale = 0.000000001, suffix = "B"))
```

## Assigning country names to their respective regions in OHI

-   **NOTE**: This year used the names_2_rgn function to find out which
    country names needed to be hardcoded to align with the reference
    regions.
    -   To do this: Hold cmd (or ctrl) and click name_2_rgn
    -   This should open up the script for the function
    -   From there, hold cmd (or ctrl) and click rgn_synonyms on
        line 31. This will pull up the synonyms for region names so that
        you can check if yours align. If they do not, you can use
        case_when() as seen below to revalue specific countries to their
        relevant synonyms

```{r}
# Need to join with regions and assign each country its respective OHI Region ----

# Checking the unique country names for anything that needs to be hard-coded
unique(total_mar_rev$country)

# This is hard coding regions with to their synonyms so that the names_2_rgn function can assign them region IDs
mar_rev_rgn <- total_mar_rev %>% 
  mutate(Country = case_when(
    country=="Antigua and Barbuda" ~ "Antigua & Barbuda",
    country=="Bonaire, Sint Eustatius and Saba" ~ "Sint Eustasius",
    country=="Bosnia and Herzegovina" ~ "Bosnia Herzegovina",
    country=="China, Hong Kong SAR" ~ "China",
    country=="Democratic People's Republic of Korea" ~ "North Korea",
    country=="Republic of Korea" ~ "South Korea",
    country=="R<e9>union" ~ "Reunion",
    country=="Saint Kitts and Nevis" ~ "Saint Kitts &amp Nevis",
    country=="Saint Pierre and Miquelon" ~ "Saint Pierre and Miqelon",
    country=="Saint Vincent and the Grenadines" ~ "Saint Vincent/Grenadines",
    country=="Türkiye" ~ "Turkey",
    country=="Yugoslavia SFR" ~ "The Former Yugoslav Republic of Macedonia",
    country=="Netherlands Antilles" ~ "Curacao",
    TRUE ~ country
  ))
  
# Using name_2_rgn function to populate values for rgn_id and rgn_name for each of the countries listed in the dataframe
mar_rev_rgn_temp <- name_2_rgn(df_in = mar_rev_rgn, 
                       fld_name='Country', # Check to make sure that the case matches
                      flds_unique = c("year"))

## Received a warning for a duplicated country. Using the following code to display the portion of the dataframe that's duplicated 
    ## Can see that in this case the duplicated country is China, with both a "China" and "China, Hong Kong SAR" both being considered china 
mar_rev_rgn_temp[duplicated(mar_rev_rgn_temp[, c('Country','year')]),  ] 

# Reducing redundancy by creating a new df with the "China" regions so I can combine them (see line 254)
mar_rev_rgn_check <- mar_rev_rgn_temp %>% 
  filter(Country =="China") %>% 
  ungroup()

# Because I combined hong kong and china, I need to remove them from the original df so that I can rbind() my combined values to the original df 
mar_rev_rgn_temp <- mar_rev_rgn_temp %>% 
  filter(!country %in% "China") %>% 
  filter(!country %in% "China, Hong Kong SAR")

# The source code was used to aggregate duplicates by population for the "Social Progress Index", This would likely be useful for another metric of progress, but since we're dealing with hard values for revenue from aquaculture, my thought process is that we can just add them together, and it will represent the combined value from both Hong Kong as well as China

mar_rev_china_temp <- mar_rev_rgn_check %>% 
  mutate(year = as.character(year)) %>% 
  # group_by(year) %>%
  dplyr::summarize(value = sum(value), .by = year)

# Okay, before I find a way to maintain columns this will have to be hard-coded to bring back the columns into the dataframe in the correct order so I can rbind it back to the primary mar_rev df 
mar_rev_china_temp1 <- mar_rev_rgn_check %>% 
  filter(country == "China") # Filtering country == "China" because I only need ONE set of years in my 'combined' df. If I maintain hong kong I wont be able to drag the columns over because the vectors will be different lengths

# Adding columns in (almost) required order AFTER ensuring the df I'm pulling from is the correct length 
mar_rev_china_temp$country = mar_rev_china_temp1$country
mar_rev_china_temp$unit = mar_rev_china_temp1$unit
mar_rev_china_temp$sector = mar_rev_china_temp1$sector
mar_rev_china_temp$data_source = mar_rev_china_temp1$data_source
mar_rev_china_temp$Country = mar_rev_china_temp1$Country
mar_rev_china_temp$rgn_id = mar_rev_china_temp1$rgn_id
mar_rev_china_temp$rgn_name = mar_rev_china_temp1$rgn_name

# Moving the "country" column to the correct position to rbind the china data back onto the entire data 
mar_rev_china_temp <- mar_rev_china_temp %>% 
  relocate(country, .before = year)

# Binding duplicate data and ensuring dataframe is in the CORRECT format to write out to /int folder ----
eco_mar_usd_raw <- rbind.data.frame(mar_rev_rgn_temp, mar_rev_china_temp) %>% 
  relocate(c(rgn_id, rgn_name), .before = country) %>% # Relocating columns in the order specified by README
  ungroup() %>% # Getting rid of any leftover groups 
  select(-c(country, sector, data_source, Country)) %>% # Getting rid of extra columns
  dplyr::rename(usd = value) %>% # Renaming value to specified format name 
  mutate(sector = "mar", # Adding sector
         usd_yr = year, # UPDATE THIS BASED ON THE USD METRIC CONVERSION (should specify in metadata, but for the data for v2024 the data was just converted to that year's USD value, so the usd_yr is just year)
         year = as.numeric(year)) # Making sure its numeric per file structure. Could check this with class()

## Checking how many unique countries are within the revenue data - 142 (2024 Assessment)
# length(unique(eco_mar_usd_raw$rgn_id))

# write_csv(eco_mar_usd_raw, here::here(data_path, "int", "eco_mar_usd_pre.csv"))
```

# JOBS!!!!!!!

## This was copied from Sophia's pdf processing she did for the fao yearbook

```{r}
#------cf FAO Yearbook Number of Fishers-------
# turn the number of fishers table from the FAO yearbook into a data frame and clean.
# note: world total includes countries and regions not within the table.  Unsure of where that data comes from; look for methodology.

# Need To update the fao_dir from above with the pdf of the FAO Fisheries and Aquaculture Statistics Yearbook
fao_dir <- here::here(raw_data_dir, "FAO", data_dir_version_year)

## Loading the pdf in as an object 
fao_pdf <- pdf_text(here::here(fao_dir, "fao_statistics_yearbook_07_03_24.pdf"))

## Selecting pdf table 
fishers_table <- fao_pdf[[170]] %>% # page number of the pdf that the desired table is on 
  as.data.frame()

## Specify columns 
columns <- c("country", "1995", "2000", "2005", "2010", "2015", "2017", "2018", "2019", "2020","2021")

## Setting the columns name because mutate wasn't picking up on the "." as a column name.
colnames(fishers_table) = c("main_column")
## Checking 
# colnames(fishers_table)


## Turn the pdf page into a clean dataframe
fao_mar_jobs <- fishers_table %>%
  separate_rows(main_column, sep = "\n") %>%  #separate rows based on line breaks
  # rename(main_column = '.') %>%  ## Removed because of above mention of column name 
  mutate(main_column = str_remove_all(main_column, "\\*")) %>% # Remove the * 
  mutate(main_column = str_replace_all(main_column, "E(?!\\p{L})", " ")) %>% # Replace all the Es not followed by letters (used to indicate estimate)
  slice(7:69) %>% #remove all rows not in the table
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% #separate into columns based on double spaces
  pivot_longer(cols = -c("country"), names_to = "year") %>% 
  mutate(value = str_squish(value)) %>% #remove extra spaces
  mutate(value = str_remove_all(value, " ")) %>% 
  mutate(value = if_else(value == "…" , NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(unit = "persons (1)",
         sector = "mar/cf",
         data_source = "FAO number of aquaculture jobs") %>% 
  slice(-c(1:44))

## Final cleaning. Removing extraneous rows not necessary 
fao_mar_jobs_clean <- fao_mar_jobs %>% 
  dplyr::slice(-c(1:6, 327:516)) %>% # %>% view() # Add view() to check df to make sure its tidy after 
  mutate(year = as.numeric(year))

########### OECD data on aquaculture employment ###########

oecd.fp <- file.path(dir_M, "git-annex/globalprep/_raw_data/OECD", data_dir_version_year, "OECD.ENV.EPI,DSD_SOE@DF_SOE,+.A.FISH_EMP_AQUACULTURE_TOT....csv")

#### BRING THIS JAWN IN #####

oecd_mar_jobs <- read.csv(oecd.fp) %>% janitor::clean_names()

length(unique(oecd_mar_jobs$reference_area))

oecd_mar_jobs_clean <- oecd_mar_jobs %>% 
  ## Getting rid of all unnecessary columns 
  select(-c(obs_status_2, observation_status_2, price_base, price_base_2, ref_year_price, price_reference_year, observation_value, time_period_2, iucn_management_categories, ret_period, freq, measure, structure, structure_id, action, measure_2, unit_measure, designation, return_periods, decimals, decimals_2, obs_status, observation_status, unit_mult, unit_multiplier)) %>% 
  ## Moving columns to more intuitive order 
  relocate(c(ref_area, reference_area, time_period, obs_value), .before = structure_name) %>% 
  ## Multiplying obs_value by 1000 to obtain raw value of jobs per country 
  mutate(value = obs_value*1000) %>% 
  ## One more relocate
  relocate(value, .after = obs_value) %>% 
  ## Renaming structure_name to source
  dplyr::rename(data_source = structure_name) %>% 
  ## Removing redundant/unneeded columns 
  select(-c(obs_value, frequency_of_observation)) %>% 
  ## Renaming columns to join with other jobs data from FAO
  dplyr::rename(year = time_period,
                country = reference_area) %>% 
  ## Relocating one last time 
  relocate(unit_of_measure, .before=data_source)

oecd_mar_jobs_join <- oecd_mar_jobs_clean %>% 
  select(-c(unit_of_measure, ref_area)) %>% 
  mutate(unit = "persons (1)",
         sector = "mar") %>% 
  relocate(c(unit, sector), .before=data_source)

## With both datasets our total unique regions comes out to 82
length(unique(fao_mar_jobs_clean$country)) + length(unique(oecd_mar_jobs_join$country))

mar_jobs_tot <- rbind(fao_mar_jobs_final, oecd_mar_jobs_join)

## Joined the unique country names comes out to 65
length(unique(mar_jobs_tot$country))

duplicates <- as.data.frame(mar_jobs_tot[duplicated(mar_jobs_tot[, c('country', 'year')]),  ])

length(unique(duplicates$country))

```

## This piece of code is for assigning countries to region id when there are dis-aggregated country names (i.e China, Hong Kong, Macao) that are considered one country for the OHI Regions

```{r}
## Hard coding a few of the country names that are either synonyms of countries or difficult to identify with name_2_rgn() function
mar_jobs_rgn <- fao_mar_jobs_clean %>%
  filter(!country %in% "World total") %>% 
  dplyr::rename(Country = country) %>% 
  mutate(Country = case_when(
   Country=="China,Taiwan" ~ "China",
    Country=="Russian Fed" ~ "Russia",
    Country=="Rep.of Korea" ~ "South Korea",
    Country=="DPR Korea" ~ "North Korea",
    Country=="Türkiye" ~ "Turkey",
   Country=="USA" ~ "United States of America",
    TRUE ~ Country # Everything else, leave it be
    ))

# Channel Islands are Jersey and Guernsey, but these are already in the data

mar_jobs_rgn_temp <- name_2_rgn(df_in = mar_jobs_rgn, 
                       fld_name='Country',
                      flds_unique = c("year"))

## Received a warning for a duplicated country. Using the following code to display the portion of the dataframe that's duplicated 
    ## Can see that in this case it was China, even though they actually didn't have any duplicated values. Interesting 
mar_jobs_rgn_temp[duplicated(mar_jobs_rgn_temp[, c('Country','year')]),  ] 

# Weight the following duplicates by population
# China, Guadeloupe, Guam, Hong Kong, Macao, Martinique, Northern Mariana Islands, Puerto Rico, Virgin Islands (U.S.)... Populations taken from google

# pop_weights <- data.frame(Country = c("China", "Hong Kong", "Macao",
#                  "Guadeloupe", "Martinique",
#                  "Guam", "Northern Mariana Islands",
#                  "Puerto Rico", "Virgin Islands (U.S.)"),
#                  pop = c(1452883897, 7632334, 669037,
#                          400276, 367507,
#                          172069, 49551,
#                          3252407, 103910))

# mar_jobs_rgn_final <- mar_jobs_rgn_temp %>%
#   # left_join(pop_weights, by = "Country") %>%
#   mutate(pop = ifelse(is.na(pop), 1, pop)) %>%
#   group_by(rgn_id, rgn_name, year) %>%
#   dplyr::summarize(score = weighted.mean(score, pop, na.rm=TRUE),
#             method = paste(unique(method), collapse=" "),
#             gapfill = weighted.mean(gapfill, pop, na.rm=TRUE))%>%
#   ungroup() %>% 
#   dplyr::mutate(year = as.numeric(year))

## Only twenty nine countries with data 
# length(unique(mar_jobs_rgn_temp$Country))
```
