---
title: "Adjusting Economies data by Sector for Inflation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
---

## Overview

The individual LE::ECO scripts from other sectors will all be adjusted for inflation within this script. After, they will be combined into one data frame. Finally, the values will be aggregated by region and year across all sectors, which can then be used to calculate the final score.

Initially, there are three scripts, each with different starting units of value, shown below:

+----------------+-----------------------------------------------------------------------+------------------------------------+--------------------------------+
|                | Metadata documentation                                                | Pre-Adjustment Unit                | Post-Adjustment Unit           |
+================+=======================================================================+====================================+================================+
| cf             | [FAO Capture Data](https://www.fao.org/fishery/en/collection/capture) | Final: USD (current year)          | USD inflation adjusted to 2017 |
|                |                                                                       |                                    |                                |
|                | [Ex-Vessel Price Data](https://github.com/SFG-UCSB/price-db-sfg)      | FAO Capture: tonnes                |                                |
|                |                                                                       |                                    |                                |
|                |                                                                       | Ex-Vessel Prices: USD/metric tonne |                                |
+----------------+-----------------------------------------------------------------------+------------------------------------+--------------------------------+
| tour           |                                                                       | USD PPP adjusted to 2015           | USD inflation adjusted to 2017 |
+----------------+-----------------------------------------------------------------------+------------------------------------+--------------------------------+
| mar            |                                                                       | USD (current year)                 | USD inflation adjusted to 2017 |
+----------------+-----------------------------------------------------------------------+------------------------------------+--------------------------------+

: Pre/Post-Inflation Adjustment Units

## Setup

```{r}
knitr::opts_chunk$set(echo = TRUE)

options(scipen=99)
```

```{r}
# ---- load packages ----
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  # general
  here,
  janitor,
  tidyverse,
  ohicore,
  
  priceR, # for inflation function
  
  # for data viz
  RColorBrewer, 
  plotly,
  scales
)


# ---- source functions ----
# common
source(here("workflow", "R", "common.R"))
# inflation function
source(here("workflow", "R", "inflation_adjustment.R"))

# ---- set year and file path info ----
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)

data_path <- here("globalprep", "le", version_year)

# output data dir for intermediate data products
int_dir <- here(data_path, "int")

```

## Read in pre-adjustment data

```{r}
# sector 1: Fishing (cf)
cf_df_pre <- read_csv(here(int_dir, "eco_cf_usd_pre.csv"))

# sector 2: Mariculture (mar)
mar_df_pre <- read_csv(here(int_dir, "eco_mar_usd_pre.csv"))

# sector 3: Tourism (tour)
tour_df_pre <- read_csv(here(int_dir, "eco_tour_usd_pre.csv"))

```

## Implement `inflation_adjustment.R`

```{r}
# ---- Fishing (cf) ----
cf_df_adj <- inflation_adjustment(
  cf_df_pre, # csv with the following columns: rgn_id, rgn_name, year, usd, unit, sector, usd_yr
  conversion_date = 2017, # USD year you want the adjusted values to be in
  country = "US", # country of unit value
  current_year = current_year) # assessment year (to index into vYYYY subfolder in "le" folder)


# ---- Mariculture (mar) ----
mar_df_adj <- inflation_adjustment(mar_df_pre, conversion_date = 2017,
                                   country = "US", current_year = current_year)


# ---- Tourism (tour) ----
tour_df_adj <- inflation_adjustment(tour_df_pre, conversion_date = 2017,
                                    country = "US", current_year = current_year)


```

```{r}
# cf_df_adj_filter <- cf_df_adj %>% 
#   filter(year %in% c())

#
```

## Combine adjusted data frames

```{r}
# bind rows of adjusted data frames
eco_usd_bind <- rbind(cf_df_adj, mar_df_adj, tour_df_adj)

# find maximum and minimum years for each adjusted data frame
min_max_df <- tibble(
  min_year = c(min(cf_df_adj$year), min(mar_df_adj$year), min(tour_df_adj$year)),
  max_year = c(max(cf_df_adj$year), max(mar_df_adj$year), max(tour_df_adj$year))
)

# set year range based on limiting dataset values
start_year <- max(min_max_df$min_year) # take maximum of minimum years
stop_year <- min(min_max_df$max_year) # take minimum of maximum years


# ----- create aggregated adjusted data frame ------------

eco_usd_filter <- eco_usd_bind %>% 
  filter(year %in% c(start_year:stop_year)) # filter to date range


# make all geographic areas have the same date range (and fill missing values with NAs)
eco_years_df <- tibble(rgn_name = eco_usd_filter$rgn_name,
                       rgn_id = eco_usd_filter$rgn_id) %>% 
  group_by(rgn_id, rgn_name) %>% 
  summarize(year = seq(start_year,
                       stop_year))
# note: returns warning to use 'reframe' instead, but that function is experimental as of July 2024

# ---- join with gdp prop data ----
eco_usd_years <- left_join(eco_years_df, eco_usd_filter,
                          by = c("rgn_id", "rgn_name", "year"))
```

Join with OHI regions before aggregating by region for all sectors
```{r}
# ------------ join with OHI Regions to make clear which regions do not have any data -----
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  select(-Notes)

# ---- join with years df ----
rgn_yr_df <- tibble(rgn_name = region_names$rgn_name,
                       rgn_id = region_names$rgn_id) %>% # take the regions from region_names, which is all of the regions, even the ones we dont have data for
  group_by(rgn_id, rgn_name) %>% 
  summarize(year = seq(start_year,
                       stop_year)) # each region should have a row for each year as well now, filled with NA

# ---- final data frame before aggregation, in which all regions and sectors are present ----
eco_usd_rgn_full <- left_join(rgn_yr_df, eco_usd_filter,
                          by = c("rgn_id", "rgn_name", "year")) # works!

```

-   plot heatmap of sectors we have data for within date range

-   find

```{r}
# ---- make NA status column ----
eco_status <- eco_usd_years %>% 
  mutate(status = !is.na(usd)) # TRUE if we DO have value data
# %>% 
#   group_by(rgn_id, rgn_name, year) %>% 
#eco_status

# ---- using the aggregate function from {stats} ----
agg_test <- eco_status %>% 
  # sum of status column for each rgn_name + year + sector combination
  aggregate(status~rgn_name+year+sector, FUN = sum)

#agg_test

# ---- count number of sectors we have data for for each country per year ----
agg_test_sum <- agg_test %>% 
  ungroup() %>% 
  group_by(rgn_name, year) %>% 
  summarize(data_status = sum(status, na.rm = TRUE))

#View(agg_test_sum)

# ---- make heatmap to visualize missing data ----
agg_heatmap_plot <- ggplot(data = agg_test_sum) +
  geom_tile(aes(x = year,
                y = rgn_name,
                fill = as.factor(data_status)), color = "black") +
  scale_fill_manual(values = c("white", "#AACAF9", "#5594F2", "#005FEC", "darkblue"), na.value = "grey") +
  labs(y = "", 
       x = "Year",
       title = "Number of Sectors with Data per Region per Year",
       fill = "Number of Sectors with Data") +
  theme_bw() 

agg_heatmap_plot


plotly::ggplotly(agg_heatmap_plot) 
# %>% 
#   add_trace(
#     x = ~year, 
#     y = ~rgn_name, 
#     fill = ~as.factor(data_status),
#     showlegend = T,
#     hovertemplate = paste('<i>Year</i>: $%{x}',
#                         '<br><b>Region Name</b>: %{y}<br>',
#                         '<b>%{fill}</b>')
#   )
```

```{r}
# calculate aggregate value across all sectors 
eco_usd_adj <- eco_usd_rgn_full %>% 
  group_by(rgn_id, rgn_name, year) %>% # group by places & years
  summarize(value = sum(adj_usd, na.rm = TRUE)) %>%  # sum values
  mutate(value = case_when( 
    value == 0 ~ NA, # replace 0s with NAs (flag of no data that was lost in summing)
    .default = value
  ))

# write data frame to int
# write_csv(eco_usd_adj, here(int_dir, "eco_usd_adj.csv"))
```

## Calculating scores following OHI methodology

-   [Methodology](https://ohi-science.org/ohi-methods/goals/goal-models-data.html#economies-subgoal-of-livelihoods-and-economies)

$$
x_{eco} = \frac { \displaystyle\sum _{ k=1 }^{ N }{ { e }_{ c,k } }  }{ \displaystyle\sum _{ k=1 }^{ N }{ { e }_{ r,k } }  }
$$

where, $e$ is the total adjusted revenue generated directly and indirectly from sector $k$, at current $c$, and reference $r$, time points

Based on this methodology and equation, we created and defined the `value` column earlier as the sum of the adjusted USD value for the three sectors (fishing, mariculture, and tourism). To calculate the score, we will take that combined sum across sectors per region per year and divide the current year's value by the

<https://stackoverflow.com/questions/743812/calculating-moving-average>

<https://www.rdocumentation.org/packages/zoo/versions/1.8-12/topics/rollmean>

```{r}
# pseudo code

# want: current year / average of 5 previous years (sum of 5 previous yrs / 5, na.rm = T)

# for each rgn_id, rgn_name


current_value <- eco_usd_adj[eco_usd_adj$year == stop_year, "value"]

reference_range <- c((stop_year - 5):(stop_year - 1))

reference_point <- (sum(eco_usd_adj[eco_usd_adj$year %in% c(reference_range), "value"], na.rm = TRUE) / length(reference_range))



# ============ stretch goals / want to implement ===========================
# potential: add if statement that calculates average based on most recent year (next available year) if current year is NA


# potential:
# want to change length(reference_range) to length(!is.na(value)) in reference range for each sector -- would need to have pre-aggregated data 
```

Score function: using average of 5 years' previous values as denominator

Note: after doing some pseudo coding and tests, we created the following function to calculate scores given a "current year" (2019 for v2024). We do want to note that we used perplexity.ai to write portions of this function after providing it with the context of the problem.

```{r}
library(dplyr)

calculate_scores <- function(df, stop_year) {
  data %>%
    group_by(rgn_id, rgn_name) %>%
    mutate(
     # current_value = value[year == max(year[year <= stop_year])],
      current_value = df[df$year == stop_year, "value"],
      
      reference_range = list(seq(max(year[year <= stop_year]) - 5, max(year[year <= stop_year]) - 1)),
      
      reference_point = sapply(reference_range, function(x) mean(value[year %in% x], na.rm = TRUE)),
      
      score = current_value / reference_point
    ) %>%
    
    filter(year == max(year[year <= stop_year])) %>%
    
    select(rgn_id, rgn_name, year, current_value, reference_point, score) %>%
    
    ungroup()
}
```

Testing function:

```{r}
test_range <- eco_usd_adj %>% 
  group_by(rgn_id, rgn_name) %>%
  mutate(
    # set current value to value in max year of most recent years (starting at stop_year)
    current_value = value[year == max(year[year <= stop_year])],
    
    # create 5 year time range of most recent years of data
    reference_range = list(seq(max(year[year <= stop_year]) - 5, max(year[year <= stop_year]) - 1)),
    
    # define reference point as the mean of values within the reference range, excluding NA values (which is important)
    reference_point = sapply(reference_range, function(x) mean(value[year %in% x], na.rm = TRUE)),
    
    # create score column as current value divided by reference point
    score = current_value / reference_point
    )
  
# test_range$reference_range
# test_range$reference_point

# Assuming your data frame is named 'eco_usd_adj'
result <- calculate_scores(eco_usd_adj, stop_year = 2019)  # Replace 2019 with your desired stop year
```

Note -- data is kept for countries that have fewer than 5 years if that data is within the time range defined earlier

```{r}
top_score <- result %>% 
  filter(score == max(result$score, na.rm = TRUE))

top_score
```

```{r}
summary <- summary(result)

paste("Summary of `score` column:")
summary(result$score)

top_change <- result %>% 
  slice_max(order_by = score, n = 10) %>% 
  ungroup()

top_change

low_change <- result %>% 
  slice_min(order_by = score, n = 10) %>% 
  ungroup()

low_change
```

Aside from the top 10 countries and bottom 4 countries, the scores of the rest of the countries were roughly in the range of 0.5-1.6

```{r}
top_change$rgn_name <- factor(top_change$rgn_name, levels = unique(top_change$rgn_name)[order(top_change$score, decreasing = TRUE)])

plotly::plot_ly(data = top_change, x = ~score, y = ~rgn_name, color = ~rgn_name, 
                                type = "bar") %>% 
  layout(title = "Economies Score: Top 10",
         xaxis = list(title = "Score"),
         showlegend = FALSE,
         margin = m,
         yaxis = list(title = '')
        # yaxis = list(title = "Score")
        )



plot_df <- result %>% ungroup()

plotly::plot_ly(data = plot_df, x = ~score, y = ~rgn_name, color = ~rgn_name, 
                                type = "bar") %>% 
  layout(title = "Economies Score",
         xaxis = list(title = "Score")
        )

```

## Alternative Score method: value 5 years ago as reference point

```{r}
calculate_scores_2 <- function(data, stop_year) {
  data %>%
    group_by(rgn_id, rgn_name) %>%
    mutate(
      # creating columns so we can perform column-wise operations 
      
      # set current year equal to stop year or else most recent year of data
      current_year = max(year[year <= stop_year]),
      # get value in current or most recent year of data
      current_value = value[year == current_year],
      # set year for reference point to 5 years before current or most recent year
      reference_year = current_year - 5,
      # get value in reference year
      reference_value = value[year == reference_year],
      # calculate score
      score = current_value / reference_value
    ) %>%
    # get year for which score was calculated for each region 
    filter(year == current_year) %>%
    # select relevant columns
    select(rgn_id, rgn_name, year, current_value, reference_value, score) %>%
    ungroup()
}



```

Test new function

```{r}
single_point_scores_2019 <- calculate_scores_2(eco_usd_adj, stop_year = 2019)
single_point_scores_2019
```

```{r}
paste("Single reference point summary:")
summary(single_point_scores_2019$score)


paste("5 year average reference point summary:")
summary(result$score)
```

In 2019, Nauru (and members of the PNA) sold
