---
title: "Adjusting Economies data by Sector for Inflation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
---


## Setup

```{r}
knitr::opts_chunk$set(echo = TRUE)

options(scipen=99)
```


```{r}
# ---- load packages ----
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  # general
  here,
  janitor,
  tidyverse,
  ohicore,
  
  priceR, # for inflation function
  
  # for data viz
  RColorBrewer, 
  plotly,
  scales
)


# ---- source functions ----
# common
source(here("workflow", "R", "common.R"))
# inflation function
source(here("workflow", "R", "inflation_adjustment.R"))

# ---- set year and file path info ----
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)

data_path <- here("globalprep", "le", version_year)

# output data dir for intermediate data products
int_dir <- here(data_path, "int")

```


## Read in pre-adjustment data

```{r}
# sector 1: Fishing (cf)
cf_df_pre <- read_csv(here(int_dir, "eco_cf_usd_pre.csv"))

# sector 2: Mariculture (mar)
mar_df_pre <- read_csv(here(int_dir, "eco_mar_usd_pre.csv"))

# sector 3: Tourism (tour)
tour_df_pre <- read_csv(here(int_dir, "eco_tour_usd_pre.csv"))

```

## Implement `inflation_adjustment.R`

```{r}
# ---- Fishing (cf) ----
cf_df_adj <- inflation_adjustment(
  cf_df_pre, # csv with the following columns: rgn_id, rgn_name, year, usd, unit, sector, usd_yr
  conversion_date = 2017, # USD year you want the adjusted values to be in
  country = "US", # country of unit value
  current_year = current_year) # assessment year (to index into vYYYY subfolder in "le" folder)


# ---- Mariculture (mar) ----
mar_df_adj <- inflation_adjustment(mar_df_pre, conversion_date = 2017,
                                   country = "US", current_year = current_year)


# ---- Tourism (tour) ----
tour_df_adj <- inflation_adjustment(tour_df_pre, conversion_date = 2017,
                                    country = "US", current_year = current_year)


```

```{r}
# cf_df_adj_filter <- cf_df_adj %>% 
#   filter(year %in% c())
```



## Combine adjusted data frames 

```{r}
# bind rows of adjusted data frames
eco_usd_bind <- rbind(cf_df_adj, mar_df_adj, tour_df_adj)

# find maximum and minimum years for each adjusted data frame
min_max_df <- tibble(
  min_year = c(min(cf_df_adj$year), min(mar_df_adj$year), min(tour_df_adj$year)),
  max_year = c(max(cf_df_adj$year), max(mar_df_adj$year), max(tour_df_adj$year))
)

# set year range based on limiting dataset values
start_year <- max(min_max_df$min_year) # take maximum of minimum years
stop_year <- min(min_max_df$max_year) # take minimum of maximum years

# ----- create aggregated adjusted data frame ------------

eco_usd_filter <- eco_usd_bind %>% 
  filter(year %in% c(start_year:stop_year)) # filter to date range
```


- plot heatmap of sectors we have data for within date range 

- find 

```{r}



eco_heatmap <- eco_usd_filter %>% 
  mutate(status = !is.na(usd)) 
# %>% 
#   group_by(rgn_id, rgn_name, year) %>% 
eco_heatmap
agg_test <- eco_heatmap %>% aggregate(status~rgn_name+year+sector, FUN = sum)

agg_test


agg_test_sum <- agg_test %>% 
  ungroup() %>% 
  group_by(rgn_name, year) %>% 
  summarize(data_status = sum(status, na.rm = TRUE))

agg_test_sum


agg_heatmap_plot <- ggplot(data = agg_test_sum) +
  geom_tile(aes(x = year,
                y = rgn_name,
                fill = as.factor(data_status)), color = "black") +
  scale_fill_manual(values = c("white", "#AACAF9", "#5594F2", "#005FEC", "darkblue"), na.value = "grey") +
  labs(y = "", 
       x = "Year",
       title = "Number of Sectors with Data per Region per Year") +
  theme_bw() 

agg_heatmap_plot


plotly::ggplotly(agg_heatmap_plot)
```


```{r}
eco_usd_adj <- eco_usd_filter %>% 
  group_by(rgn_id, rgn_name, year) %>% # group by places & years
  summarize(value = sum(adj_usd, na.rm = TRUE)) %>%  # sum values
  mutate(value = case_when( 
    value == 0 ~ NA, # replace 0s with NAs (flag of no data that was lost in summing)
    .default = value
  ))



# write data frame to int
#write_csv(eco_usd_adj, here(int_dir, "eco_usd_adj.csv"))

```


## Calculating scores following OHI methodology

- [Methodology](https://ohi-science.org/ohi-methods/goals/goal-models-data.html#economies-subgoal-of-livelihoods-and-economies)

$$
x_{eco} = \frac { \displaystyle\sum _{ k=1 }^{ N }{ { e }_{ c,k } }  }{ \displaystyle\sum _{ k=1 }^{ N }{ { e }_{ r,k } }  }
$$

where, $e$ is the total adjusted revenue generated directly and indirectly from sector $k$, at current $c$, and reference $r$, time points

Based on this methodology and equation, we created and defined the `value` column earlier as the sum of the adjusted USD value for the three sectors (fishing, mariculture, and tourism). To calculate the score, we will take that combined sum across sectors per region per year and divide the current year's value by the 

<https://stackoverflow.com/questions/743812/calculating-moving-average>

<https://www.rdocumentation.org/packages/zoo/versions/1.8-12/topics/rollmean>

```{r}
# pseudo code

# want: current year / average of 5 previous years (sum of 5 previous yrs / 5, na.rm = T)

# for each rgn_id, rgn_name


current_value <- eco_usd_adj[eco_usd_adj$year == stop_year, "value"]

reference_range <- c((stop_year - 5):(stop_year - 1))

reference_point <- (sum(eco_usd_adj[eco_usd_adj$year %in% c(reference_range), "value"], na.rm = TRUE) / length(reference_range))






# ============ stretch goals / want to implement ===========================
# potential: add if statement that calculates average based on most recent year (next available year) if current year is NA


# potential:
# want to change length(reference_range) to length(!is.na(value)) in reference range for each sector -- would need to have pre-aggregated data 
```



Test

```{r}
library(dplyr)

calculate_scores <- function(df, stop_year) {
  data %>%
    group_by(rgn_id, rgn_name) %>%
    mutate(
     # current_value = value[year == max(year[year <= stop_year])],
      current_value = df[df$year == stop_year, "value"],
      
      reference_range = list(seq(max(year[year <= stop_year]) - 5, max(year[year <= stop_year]) - 1)),
      
      reference_point = sapply(reference_range, function(x) mean(value[year %in% x], na.rm = TRUE)),
      
      score = current_value / reference_point
    ) %>%
    
    filter(year == max(year[year <= stop_year])) %>%
    
    select(rgn_id, rgn_name, year, current_value, reference_point, score) %>%
    
    ungroup()
}
```


Testing function: 

```{r}
test_range <- eco_usd_adj %>% 
  group_by(rgn_id, rgn_name) %>%
  mutate(
    # set current value to value in max year of most recent years (starting at stop_year)
    current_value = value[year == max(year[year <= stop_year])],
    
    # create 5 year time range of most recent years of data
    reference_range = list(seq(max(year[year <= stop_year]) - 5, max(year[year <= stop_year]) - 1)),
    
    # define reference point as the mean of values within the reference range, excluding NA values (which is important)
    reference_point = sapply(reference_range, function(x) mean(value[year %in% x], na.rm = TRUE)),
    
    # create score column as current value divided by reference point
    score = current_value / reference_point
    )
  
test_range$reference_range
test_range$reference_point

# Assuming your data frame is named 'eco_usd_adj'
result <- calculate_scores(eco_usd_adj, stop_year = 2019)  # Replace 2019 with your desired stop year
```

Note -- data is kept for countries that have fewer than 5 years if that data is within the time range defined earlier

```{r}
top_score <- result %>% 
  filter(score == max(result$score, na.rm = TRUE))

top_score
```


```{r}
summary <- summary(result)

paste("Summary of `score` column:")
paste0(summary[36:42])

top_change <- result %>% 
  slice_max(order_by = score, n = 10) %>% 
  ungroup()

top_change

low_change <- result %>% 
  slice_min(order_by = score, n = 10) %>% 
  ungroup()

low_change
```
Aside from the top 10 countries and bottom 4 countries, the scores of the rest of the countries were roughly in the range of 0.5-1.6


```{r}
top_change$rgn_name <- factor(top_change$rgn_name, levels = unique(top_change$rgn_name)[order(top_change$score, decreasing = TRUE)])

plotly::plot_ly(data = top_change, x = ~score, y = ~rgn_name, color = ~rgn_name, 
                                type = "bar") %>% 
  layout(title = "Economies Score: Top 10",
         xaxis = list(title = "Score"),
         showlegend = FALSE,
         margin = m,
         yaxis = list(title = '')
        # yaxis = list(title = "Score")
        )



plot_df <- result %>% ungroup()

plotly::plot_ly(data = plot_df, x = ~score, y = ~rgn_name, color = ~rgn_name, 
                                type = "bar") %>% 
  layout(title = "Economies Score",
         xaxis = list(title = "Score") #,
        # yaxis = list(title = "Score")
        )

```

