---
title: "OHI `r format(Sys.Date(), '%Y')` - Economies Tourism Dataprep"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------


# Economies Dataprep (Measures the amount of revenue produced)

## Overview

-   **Data:**

### Tourism direct GDP as a proportion of total GDP 

Tourism direct GDP as a proportion of total GDP (indicator 8.9.1): UNWTO Dept. of Statistics (UN Tourism)

-   <https://www.unwto.org/tourism-statistics/economic-contribution-SDG> 
-   A big issue with this data set is that it does not have data for
mainland China. Thus, I was instructed to gapfill using another
data source. I found tourism revenue data on the website for the
National Bureau of Statistics for China

-   China Gapfilling -- °ºNational Data: National Bureau of Statistics of China (NBS) <https://data.stats.gov.cn/english/easyquery.htm?cn=C01>

-   Made an account to download data:
-   Email:
[aramji\@bren.ucsb.edu](mailto:aramji@bren.ucsb.edu){.email}
-   Password: OHI.f3ll0ws!
-   Security Question: who has influenced you the most?
-   Answer: Melanie Frazier

-   Returned to home page, clicked “Annual”, clicked “Tourism” from options on the left, clicked “Year” dropdown menu and selected “LATEST20”, then clicked the download button and selected “csv”. File appeared as “Annual.csv”, which I then renamed to `eco_tour_china_all_metrics_2004-2023.csv` 
-   Added to new file in `_raw_data` on Mazu → NBS_China

-   For this tourism revenue data used for China, I wanted to discern whether or not these numbers included Macao and Hong Kong, as that can sometimes be the case (and would significantly impact my data processing methods). Thankfully, the website where I downloaded the data had information on this: 

> "National statistical indicators involved in the database don't contain data from Hong Kong Special Administrative Regions (SAR), Macao SAR and Taiwan Province except for administrative divisions, land area, forest resources and precipitation. Hong Kong SAR and Macao SAR are a part of the overall national statistics. According to the relevant principles of the PRC Hong Kong Basic Law and Macao SAR Basic Law, Hong Kong, Macao and the mainland are relatively independent statistical regions. Based on their different statistical systems and legal requirements, they conduct statistical work independently." <https://data.stats.gov.cn/english/staticreq.htm?m=aboutctryinfo#:~:text=National%20statistical%20indicators%20involved%20in,area%2C%20forest%20resources%20and%20precipitation.> 

-    while this is a little ambiguous, I think it's safe to interpret this to mean that the tourism data we downloaded from the site does not include data from/about Macao and Hong Kong



### GDP: World Bank, in "constant 2015 US$"

-   <https://data.worldbank.org/indicator/NY.GDP.MKTP.KD> 

    -   Downloaded July 9, 2024. 
    -   Contains country-level annual GDP value (in constant 2015 USD) from 1960-2023.
    -   Additional metadata copied and saved as a `.rtf` in the `WorldBank_global_annual_GDP_2015_constant_USD` subfolder of the `_raw_data`'s `WorldBank` folder for `d2024`. 
    -   Date last updated (found in the first few rows of the `.csv` when opened locally using Numbers): 2024-06-28


### USA Coastal Tourism

Economics: National Ocean Watch (ENOW) Data

 -   Marine Economies: Industries for States and Coastal US
 -   https://coast.noaa.gov/digitalcoast/data/ 
 
Office for Coastal Management, 2024: Time-Series Data on the Ocean and Great Lakes Economy for Counties, States, and the Nation between 2005 and 2021 (Sector Level) from 2008-01-01 to 2019-12-31. NOAA National Centers for Environmental Information, https://www.fisheries.noaa.gov/inport/item/48033. Downloaded 2024-07-18.

> Abstract: Economics: National Ocean Watch (ENOW) contains annual time-series data for over 400 coastal counties, 30 coastal states, 8 regions, and the nation, derived from the Bureau of Labor Statistics and the Bureau of Economic Analysis. It describes six economic sectors that depend on the oceans and Great Lakes and measures four economic indicators: Establishments, Employment, Wages, and Gross Domestic Product (GDP).

    -   Metadata: https://www.fisheries.noaa.gov/inport/item/48033 

-   Range: 2005-2021

Has sector-level data:
-   Includes Ship and Boat Building, Tourism and Recreation, Living Resources, Fish hatcheries and aquaculture, Seafood processing, Seafood markets, Fishing, Marine Construction, Marine Transportation, All Ocean Sectors
-   Has employment, wages, GDP, RealGDP columns (RealGDP = adjusted for inflation -- theoretically to year of most recent update based on the metadata, which would be 2021. However, upon futher investigation, the gdp and real gdp values match in the year 2012, so it seems that RealGDP is GDP adjusted to 2012 USD)
-   Contains -9999 values for data privacy – these values are captured in the totals but not in state-level

- On Mazu in `home/shares/ohi/git-annex/globalprep/_raw_data/ENOW/d2024/ENOW_Industries/ENOW_Industries_2005_2021.csv`

-   Downloaded from NOAA Office for Coastal Management; Digital Coast website: <https://coast.noaa.gov/digitalcoast/data/>.

### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  ohicore, # name_2_rgn
  here,
  countrycode,
  janitor,
  priceR, # for inflation adjustment
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  ggplot2,
  tidyr,
  dplyr,
  readr,
  readxl,
  httr,
  scales,
  plotly,
  zoo # for gapfilling
  
)
# for extra gapfilling (v2024: didn't use this method)
# remotes::install_github("skgrange/threadr") # for na_extrapolate
#library(threadr)


# source 
source(here::here("workflow", "R", "common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here::here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO raw data directory
unwto_dir <- here::here(raw_data_dir, "UNWTO", data_dir_version_year)

# World Bank raw data dir (for GDP data)
wb_dir <- here::here(raw_data_dir, "WorldBank", data_dir_version_year)

# China National Bureau of Statistics (NSB) directory
nbsc_dir <- here::here(raw_data_dir, "NBS_China", data_dir_version_year)

# Economics: National Ocean Watch (ENOW) Data for USA Coastal Tourism
enow_dir <- here::here(raw_data_dir, "ENOW", data_dir_version_year, "ENOW_Industries")

# Output data directory -- for intermediate data products
int_dir <- here::here(data_path, "int")
```


## Read in Data

```{r}
# ===================== Read in Data ==============================

# ---- Tourism revenue data ----
gdp_prop_raw <- readxl::read_xlsx(here::here(unwto_dir,
                                       "Indicator-8_9_1-2022-UN_Tourism_april2024_update.xlsx")) # update this if/when new data is downloaded

# ---- OHI regions data ----
region_names <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  janitor::clean_names() %>% 
  dplyr::select(-notes)

# ---- World Bank GDP in 2015 Constant USD data ----
wb_gdp_raw <- readr::read_csv(here::here(wb_dir,
                                   "WorldBank_global_annual_GDP_2015_constant_USD",
                                   "WorldBank_global_annual_GDP_2015_constant_USD.csv"),
                              skip = 4) # skip first 4 rows
# date range from 1960 to 2023


# ---- NBSC Data (different units for international and domestic tourism) ----
nbsc_raw <- readr::read_csv(here::here(nbsc_dir,
                                       "eco_tour_china_all_metrics_2004-2023.csv"), # update this if/when new data is downloaded
                            skip = 2) # skip first 2 rows

# ---- ENOW Data for USA Coastal Tourism Gapfilling ----
enow_usa_raw <- readr::read_csv(here::here(enow_dir, "ENOW_Industries_2005_2021.csv"), na = c("-9999"))

```


## Tourism GDP % 

### Data cleaning

Current unit of "total" in dataframe:

- "Tourism direct GDP as a proportion of total GDP (%)" 

```{r}
# ---- preliminary cleaning/tidying ----
gdp_prop_clean <- gdp_prop_raw %>% 
  janitor::clean_names() %>% 
  dplyr::select(c(geo_area_name, year = time_period, gdp_prop_percent = total))

# get number of geo areas in dataset
length(unique(gdp_prop_clean$geo_area_name))
# v2024: 117

# make all geographic areas have the same date range (and fill missing values with NAs)
years_df <- tibble(geo_area_name = gdp_prop_clean$geo_area_name) %>% 
  dplyr::group_by(geo_area_name) %>% 
  dplyr::summarize(year = seq(min(gdp_prop_clean$year),
                       max(gdp_prop_clean$year)))

# ---- join with gdp prop data ----
gdp_prop_yrs <- dplyr::left_join(years_df, gdp_prop_clean,
                          by = c("geo_area_name", "year"))

# here we can see that the update has been made:
gdp_prop_yrs[46:60,] # for countries without data in certain years,
# the value column (gdp_prop_percent) is populated with NAs (which is what we want!)
# and all geo areas have the same number of years
```

#### Preliminary plot 1: full dataset
Create preliminary plot of cleaned full dataset (not joined/filtered to OHI regions yet)
```{r}
# ---- preliminary plot ----
# define margins
m <- list(
  l = 80,
  r = 60,
  b = 80,
  t = 80,
  pad = 4
)

# ---- create plotly interactive plot ----
prelim_plot <- plotly::plot_ly(data = gdp_prop_yrs,
                       x = ~year,
                       y = ~gdp_prop_percent,
                       color = ~geo_area_name, 
                       type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  plotly::layout(title = "Tourism direct GDP as a proportion of total GDP (%) for All Geo Regions",
         margin = m, # set margins
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

# display plot
prelim_plot

# optional: 
# htmlwidgets::saveWidget(prelim_plot, file = "tourism_gdp_prop_all_data.html")
```
v2024 notes/observations:

-   Macao has an extremely high percent of tourism contributing to their GDP (max of 59.6% in 2023)
-   all geographic areas appear to have a sharp decrease in tourism contribution to GDP in 2020, which is to be expected due to the COVID-19 Pandemic. From 2019 to 2020, Macao saw a fall by more than half, from 52.57% in 2019 to 21.44% in 2020.



### Filter & Run OHI region name function

```{r}
# check that this is the intended stop/cutoff year
stop_year <- 2019 # update this depending on data updates

# ---- Filter ----
# set date range selection (rationale described in the paragraph below the previous plot)
gdp_filter <- gdp_prop_yrs %>% 
  dplyr::filter(year <= stop_year)

# ---- Run name_2_rgn -----
# note: if we don't namespace from ohicore:: here, it'll use 
# the local name_2_rgn function stored/defined in this file:
# ~/ohiprep_v2024/globalprep/cw_trend_trash/v2016/georgn_gapfill.R
gdp_rgn <- ohicore::name_2_rgn(df_in = gdp_filter, 
                      fld_name = "geo_area_name",
                      flds_unique = c("year"))

# sorting through mismatch and aggregation ----

# v2024: note on Northern Mariana Islands and Guam OHI region: we only have data for Guam, not Northern Mariana Islands... this is problematic because in the World Bank GDP dataset, we only have GDP data for both the Northern Mariana Islands and Guam and need to aggregate them to fit into the OHI regions system. For now, we'll drop the Northern Mariana Islands GDP data in the WB GDP processing, but please note that if you get GDP % data for this region later on, you'll need to fix that in the GDP data processing chunk.

# v2024: China: we only have data on Hong Kong and Macao, no mainland China data...
# Mel told us to just drop them... which makes sense in context, but is unfortunate, as Macao had the highest values. I will look for data on mainland China to supplement if possible.

# check duplicates         # v2024: (none)
tour_eco_duplicates <- gdp_rgn[duplicated(gdp_rgn[, c("rgn_name", "year")]),]
unique(tour_eco_duplicates$geo_area_name)
unique(tour_eco_duplicates$rgn_name)

length(unique(gdp_rgn$rgn_name))
# v2024: 93

# ---- Rename Macao and Hong Kong ----
gdp_prop_agg <- gdp_filter %>% 
  dplyr::mutate(geo_area_name = case_when(
    geo_area_name == "China, Macao Special Administrative Region" ~ "Macao SAR, China",
    geo_area_name == "China, Hong Kong Special Administrative Region" ~ "Hong Kong SAR, China",
    .default = geo_area_name
  ))


# ---- Run name_2_rgn again -----
gdp_rgn_agg <- ohicore::name_2_rgn(df_in = gdp_prop_agg, 
                          fld_name = 'geo_area_name',
                          flds_unique = c("year"))

# great! now nothing has been removed... now we need to check the duplicates 
tour_agg_duplicates <- gdp_rgn_agg[duplicated(gdp_rgn_agg[, c("rgn_name", "year")]),]
unique(tour_agg_duplicates$geo_area_name)
unique(tour_agg_duplicates$rgn_name)
# inspect duplicates
# gdp_rgn_agg %>% filter(rgn_name == "China")

# ----- China subsets ----
# exclusively China regions
macao_hk_subset <- gdp_rgn_agg %>% 
  dplyr::filter(rgn_name == "China")


# excluding China regions
# (we'll join them back together later once we have the tourism values for HK, Macao, and China) 
no_mc_hk_subset <- gdp_rgn_agg %>% 
  dplyr::filter(rgn_name != "China")


```


#### Exploratory plot: OHI regions, based on `rgn_name`, excluding China

Create preliminary plot with the newly added `rgn_name` field

```{r}
# ---- updated prelim plot ----
ohi_rgns_prelim_plot <- plot_ly(data = no_mc_hk_subset, x = ~year, y = ~gdp_prop_percent, color = ~geo_area_name, 
                                type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for OHI Regions",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

ohi_rgns_prelim_plot


# htmlwidgets::saveWidget(ohi_rgns_prelim_plot, file = "eco_tour_ohi_rgns_prelim_plot.html")
```

It is notable how much the y-axis scale changes once we drop Macao. I think that the -0.1 value we see for Fiji could be indicative of the country spending money on tourism infrastructure, development, and advertisement but then not getting enough tourism revenue due to Covid, but this is worth investigating further (I couldn't find anything helpful in the metadata, unfortunately).


### Heatmap: investigating data gaps


```{r}
library(ggplot2)

heatmap_df <- gdp_rgn_agg %>% 
  dplyr::mutate(status = !is.na(gdp_prop_percent)) %>% 
  dplyr::ungroup()

heatmap_plot <- ggplot(data = heatmap_df) +
  geom_tile(aes(x = year,
                y = geo_area_name,
                fill = as.factor(status)), color = "black") +
  scale_fill_manual(values = c("gainsboro", "darkgreen")) +
  labs(y = "",
       x = "Year",
       title = "Data Status per Region per Year",
       fill = "Does the dataset \n have data for \n this region?:") +
  theme_bw() 

plotly::ggplotly(heatmap_plot)

```



```{r}
# make subset showing counts of data per country

raw_gdp_pct_data_summary <- heatmap_df %>% 
  # group by region name and year
  dplyr::group_by(geo_area_name) %>% 
  # return counts of data points per country pe year (essentially number of years we do have data for)
  dplyr::summarize(n_data_points = sum(status)) %>% 
  dplyr::arrange((n_data_points)) # show number of data points that we have for that region across the full time range 

head(raw_gdp_pct_data_summary, n = 20)

# sum(heatmap_df$status)
```


## Country filtering

Rules:

- Must have at least one data point in the last five years &&
- must have > three data points over time


- After inspecting all of the Economies sector datasets, we decided to set our date range from 2008-2019. There are several reasons for this range:
-  Tourism data starts in 2008
-  Fishing data ends in 2019
-  2020 is rife with statistical anomalies -- this would seriously impact gapfilling (for example, a low 2020 data point may cause extremely inaccurate gapfilling for data missing in 2018-2019)

```{r}
# define date range for filtering
start_year <- 2008
stop_year <- 2019 
date_range <- c(start_year:stop_year) # update this depending on your data!

# create date range of the last 5 years (including most recent year)
recent_years <- c(((max(date_range)) - 4):max(date_range))

# check to see if we need to ungroup
is_grouped_df(heatmap_df)

# ---- at least 1 data point in the last 5 years ----
# filter by number of "TRUE" status
five_yr_subset <- heatmap_df %>% 
  dplyr::filter(year %in% c(recent_years)) %>% 
  dplyr::group_by(geo_area_name) %>% 
  # group_by(rgn_id, rgn_name) %>% 
  dplyr::summarize(n_data_points = sum(status)) %>% 
  dplyr::filter(n_data_points >= 1)
# arrange(n_data_points) # for visual inspection

# extract geo area names
five_yr_geos <- c(five_yr_subset$geo_area_name)

# look at loss from first filtering step:
length(five_yr_geos) # v2024: 89 
length(unique(heatmap_df$geo_area_name)) - length(five_yr_geos) # number of geo areas filtered out
# v2024: 6 
# --- more than 3 data points in time range ----
total_data_points <- heatmap_df %>% 
  # filter to date range
  dplyr::filter(year %in% c(date_range)) %>% 
  # group by geo_area_name
  dplyr::group_by(geo_area_name) %>% 
  # summarize n_data_points
  dplyr::summarize(n_data_points = sum(status)) %>% 
  # filter to > 3
  dplyr::filter(n_data_points > 3)
# extract names

over_three_pts_geos <- c(total_data_points$geo_area_name)  

# look at loss from first filtering step:
length(over_three_pts_geos) # v2024: 77 
length(unique(heatmap_df$geo_area_name)) - length(over_three_pts_geos) # number of geo areas filtered out
# v2024: 18


# ---- filter rgn_names subsets ----
# filter to geo area names in five_yr_subset && total_data_points
# filter(geo_area_name %in% c(five_yr_geos) && geo_area_name %in% c(over_three_pts_geos))

# Subset without Macao and Hong Kong
no_china_filter <- no_mc_hk_subset %>% 
  # filter year for specified date range (double-checking)
  dplyr::filter(year %in% c(date_range)) %>% 
  # filter to geo_area_name that fulfills both (AND) conditions/rules 
  dplyr::filter(geo_area_name %in% c(five_yr_geos) & geo_area_name %in% c(over_three_pts_geos)) %>% 
  # add status column back in for plotting
  dplyr::mutate(status = !is.na(gdp_prop_percent)) %>% 
  dplyr::ungroup()

# Macao & Hong Kong subset
macao_hk_filter <- macao_hk_subset %>% 
  # filter year for specified date range (double-checking)
  dplyr::filter(year %in% c(date_range)) %>% 
  # filter to geo_area_name that fulfills both (AND) conditions/rules 
  dplyr::filter(geo_area_name %in% c(five_yr_geos) & geo_area_name %in% c(over_three_pts_geos)) %>% 
  # add status column back in for plotting
  dplyr::mutate(status = !is.na(gdp_prop_percent)) %>% 
  dplyr::ungroup()
```

#### Heatmap: subset without Macao & Hong Kong

```{r no-mc-hk-heatmap}
# ---- heatmap to check if countries we expected to be removed were removed ----

no_china_heatmap <- ggplot(data = no_china_filter) +
  geom_tile(aes(x = year,
                y = geo_area_name,
                fill = as.factor(status)), color = "black") +
  scale_fill_manual(values = c("gainsboro", "darkgreen")) +
  labs(y = "",
       x = "Year",
       title = "Filtered Data Status per Region per Year (excluding Macao & Hong Kong)",
       fill = "Does the dataset \n have data for \n this region?:") +
  theme_bw() 

plotly::ggplotly(no_china_heatmap)
```

After discussing the datagaps after filtering, Melanie advised that we manually remove Argentina (in v2024, we had 4 total data points for this country, but they were all from the past 4 years (2016-2019)). Note: we also removed the United States in this step, so we're using the ENOW data read in earlier to address this data gap.

```{r}
no_china_filter <- no_china_filter %>% 
  dplyr::filter(!geo_area_name %in% c("Argentina"))
```

#### Heatmap: subset of Macao & Hong Kong

```{r macao-hk-heatmap}
macao_hk_heatmap <- ggplot(data = macao_hk_filter) +
  geom_tile(aes(x = year,
                y = geo_area_name,
                fill = as.factor(status)), color = "black") +
  scale_fill_manual(values = c("gainsboro", "darkgreen")) +
  labs(y = "",
       x = "Year",
       title = "Filtered Data Status per Region per Year (Macao & Hong Kong)",
       fill = "Does the dataset \n have data for \n this region?:") +
  theme_bw() 

plotly::ggplotly(macao_hk_heatmap)
```



### Gapfilling

Here we'll use `na.approx()` from {zoo} to interpolate missing values and extrapolate missing extremes (NA values that do not fall between non-NA values) by copying the nearest extreme value. After speaking with Melanie Frazier about our methodology, she recommended to not fill in any values for countries with only 1 data point (or any that don't satisfy the other rules), which is why we're doing this step after filtering. 

```{r}

# ---- Gapfilling ----
# subset excluding China regions ----
# estimate tourism GDP proportion for geo areas with missing data
no_china_gf <- no_china_filter %>% 
  dplyr::group_by(rgn_id, rgn_name) %>% 
  # interpolate (fill missing values between 2 values)
  dplyr::mutate(appx_gdp_pct = zoo::na.approx(gdp_prop_percent, # using values in this column
                                       na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                       #  extrapolate using rule = 2 from approx(),
                                       # which uses closest data extreme to
                                       #  extrapolate for leading and trailing NAs
                                       rule = 2
  )) %>% 
  # rename old status column
  dplyr::rename(raw_status = status) %>% 
  # add gapfilled (GF) status column
  dplyr::mutate(gf_status = !is.na(appx_gdp_pct)) %>% 
  dplyr::ungroup()

# subset with Macao and Hong Kong ----
# estimate tourism GDP proportion for geo areas with missing data
macao_hk_gf <- macao_hk_filter %>% 
  dplyr::group_by(geo_area_name) %>% # not by region name, which would be "China" for both
  # interpolate (fill missing values between 2 values)
  dplyr::mutate(appx_gdp_pct = zoo::na.approx(gdp_prop_percent, # using values in this column
                                       na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                       #  extrapolate using rule = 2 from approx(),
                                       # which uses closest data extreme to
                                       #  extrapolate for leading and trailing NAs
                                       rule = 2
  )) %>% 
  # rename old status column
  dplyr::rename(raw_status = status) %>% 
  # add gapfilled (GF) status column
  dplyr::mutate(gf_status = !is.na(appx_gdp_pct)) %>% 
  dplyr::ungroup()
```


```{r}
# subset without OHI China regions ----
no_china_gf_heatmap <- ggplot(data = no_china_gf) +
  geom_tile(aes(x = year,
                y = geo_area_name,
                fill = as.factor(gf_status)), color = "black") +
  scale_fill_manual(values = c("darkgreen", "gainsboro")) +
  labs(y = "",
       x = "Year",
       title = "Filtered & Gapfilled Data Status per Region per Year (excluding Macao & Hong Kong)",
       fill = "Does the dataset \n have data for \n this region?:") +
  theme_bw() 

plotly::ggplotly(no_china_gf_heatmap)
```


Because of all of the filtering that we did earlier, the result of this gapfilling is a completely filled dataset. In previous iterations when we gapfilled before filtering, there were still many NAs (FALSEs) due to large data gaps in the raw dataset.


#### Exploratory plotting: gapfilled GDP % data

```{r}
# ---- create plotly interactive plot ----
prelim_plot_gf <- plot_ly(data = no_china_gf,
                          x = ~year,
                          y = ~appx_gdp_pct,
                          color = ~geo_area_name, 
                          type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  layout(title = "Tourism direct GDP as a proportion of total GDP (%), Gapfilled \n (Exluding China, Macao & Hong Kong)",
         margin = m, # set margins
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism proportion of total GDP (%)"))

# display plot
prelim_plot_gf
```


In future years, when we have more data for 2020 and beyond, or if novel gapfilling techniques are used (e.g., not linear fitting, as the 2019-2020 trend is an irregular pattern poorly suited to linear models), we recommend not cutting off the data after 2019. Studying the impacts of the pandemic on tourism value and the economies subgoal as a whole is important and seems like a very interesting opportunity for future years' fellows.




## GDP: WorldBank annual GDP (Constant 2015 US$)

-  Global dataset of country-level GDP in constant 2015 USD (they write this as US$ in World Bank documentation)

### Data cleaning


```{r}
# use minimum (start) year in tourism GDP % dataset for filtering:
# start_year 2008
# stop_year 2019 
# date_range 2008:2019

# ---- cleaning WorldBank GDP data ----
wb_gdp_clean <- wb_gdp_raw %>% 
  janitor::clean_names() %>% 
  # remove unwanted columns
  dplyr::select(-c(indicator_code, x69)) %>% 
  # pivot to put years in 1 column, values in another
  tidyr::pivot_longer(cols = x1960:x2023, # update if new data downloaded
                      names_to = "year", values_to = "gdp_value") %>% 
  # tidy up year column
  dplyr::mutate(year = as.numeric(str_remove_all(year, 'x'))) %>%  
  # filter to date range
  dplyr::filter(year %in% c(date_range))

```

### Next steps:

-  create subset of GDP data without China, Hong Kong, or Macao: `no_china_gdp`
-  run `name_2_rgn` function on this, make necessary adjustments: `nc_gdp_rgn`, final post-aggreagtion and adjustments: `nc_gdp_rgn_agg`
-  then join with gapfilled, filtered, no-china dataframe `no_china_gf`; call this `no_china_join`
-  calculate usd tourism value for `no_china_join`. call this `nc_tourism_value`

-  create China subset of GDP data `china_gdp`, join with `macao_hk_gf` (gdp proportion subset of Macao and Hong Kong, gapfilled)
-  calculate value (will be in Constant 2015 US$): `macao_hk_value`

## Tourism Value (no China)

Now we'll prep to join our GDP and % GDP data frames by first making a subset without China, Macao, or Hong Kong. Then, we'll run `name_2_rgn` from {ohicore} and see what else we need to fix before joining.

```{r nc_rgn_agg}
# ---- make subset excluding China ----
no_china_gdp <- wb_gdp_clean %>% 
  # this works because Macao, Hong Kong, and mainland China all have "China" 
  # in their country_name designations in this dataset
  dplyr::filter(!stringr::str_detect(country_name, "China"))

# check
nrow(no_china_gdp %>% dplyr::filter(stringr::str_detect(country_name, "China")))
# should be 0


# ---- run name_2_rgn ----
# nc = no china
nc_gdp_rgn <- ohicore::name_2_rgn(df_in = no_china_gdp, 
                         fld_name = "country_name",
                         flds_unique = c("year"))

# thankfully, in the list following this line: "These data were removed for not having any match in the lookup tables:", there are mostly just the WorldBank aggregate groups (i.e., "low & middle income, Latin America & Caribbean, etc.), with the exception of Channel Islands (has all NA) and Isle of Man

# don't make the mistake I did, and check your tourism dataset to see if it has Channel Islands, Jersey, or Guernsey first! 
# it doesn't, so we can continue on since the name_2_rgn function helpfully removed this data from the GDP data frame already.

# The Isle of Man is a distinct governing body, and should not be aggregated with Ireland or the UK. It is also not in our tourism dataset, so we can continue on with it being omitted from the df by name_2_rgn. 

# ---- identify duplicates ----
nc_gdp_duplicates <- nc_gdp_rgn[duplicated(nc_gdp_rgn[, c("rgn_name", "year")]),]
#View(wb_gdp_duplicates)
unique(nc_gdp_duplicates$country_name)
unique(nc_gdp_duplicates$rgn_name)


# ---- addressing duplicates ----
# [1] "Northern Mariana Islands and Guam" --> we have GDP data for Northern Mariana Islands, and Guam. Can aggregate this, but we only have tourism % GDP data for Guam (see earlier comment in GDP % data cleaning chunk). 
# For now, we'll just keep Guam GDP and filter out Northern Mariana Islands GDP.


# [2] "Puerto Rico and Virgin Islands of the United States" --> we have data for both in the GDP data frame, but only for Puerto Rico in the GDP % data frame. Thus, we'll need to filter out the US Virgin Islands GDP data and only keep Puerto Rico data to get a more representative value (Puerto Rico tourism proportion of GDP * Puerto Rico GDP, rather than taking the Puerto Rico prop * Puerto Rico + US Virgin Islands GDP)



# ---- aggregate ----
# sum by rgn_name etc.
# name ~ No China GDP region aggregated
nc_gdp_rgn_agg <- nc_gdp_rgn %>%
  dplyr::filter(!country_name %in% c("Northern Mariana Islands", # filtering out NMI since we only have Guam GDP % data in our other dataset -- should revisit this and update once we have that data -- double check to see which of these regions we have data for on both datasets in future years of updates!!!
                                     "Virgin Islands (U.S.)") # similarly, filtering out U.S. Virgin Islands. Only do this if the data continues to have the gaps addressed above.
  ) %>% 
  dplyr::group_by(rgn_id, rgn_name, year) %>%
  dplyr::summarize(gdp_value_agg = sum(gdp_value, na.rm = TRUE)) %>%
  dplyr::ungroup() %>% 
  # turn 0s into NAs (appropriate flag for lack of data, as earlier sum(na.rm = TRUE) step obscured NAs)
  dplyr::mutate(gdp_value_agg = case_when(
    gdp_value_agg == 0 ~ NA, # replace 0 with NA
    .default = gdp_value_agg # maintain all other values
  ))

# ---- aggregation check ----
# to double check and see if Northern Marian Islands was properly filtered out:
agg_nmi_vals <- nc_gdp_rgn_agg %>% 
  dplyr::filter(rgn_name %in% c("Northern Mariana Islands and Guam")) %>%
  dplyr::select(year, gdp_value_agg)

guam_reference <- nc_gdp_rgn %>% 
  dplyr::filter(country_name %in% c("Guam")) %>% 
  dplyr::select(year, gdp_value)

agg_nmi_check <- dplyr::left_join(agg_nmi_vals, guam_reference)

print(paste("Do the Guam GDP values before aggregation match the 'Northern Mariana Islands and Guam' values after aggregation?"))

if (sum(agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value) == length(unique(agg_nmi_check$gdp_value_agg))
) {
  print("Yes")
} else {
  print("No")
}

# If "No", check which don't match
# print(paste((agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value)))
# use FALSEs for indexing

```



Great! Now we're ready to join with the tourism as a proportion of total GDP subset that excludes China. 
We'll then multiply the percent of GDP by the GDP to yield a value (in constant 2015 USD)

```{r nc_tourism_value}
# ---- join GDP Value (constant 2015 US$) with Tourism GDP % ----
no_china_join <- dplyr::full_join(nc_gdp_rgn_agg, no_china_gf,
                                  by = join_by("rgn_id", "rgn_name", "year"))
# note that because we preserved the `geo_area_name` column, we can see which country is the true representative of the aggregate based on our steps above (e.g., Puerto Rico is the geo_area_name value for the Puerto Rico and US Virgin Islands rgn_name rows)

# that is mostly only useful in the intermediate data processing though, so we will be dropping this column as we move on

# ---- calculate tourism value ----
nc_tourism_value <- no_china_join %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(rgn_id, rgn_name, year) %>% 
  # get proportion to use as multiplier, multiply by GDP to get value
  dplyr::mutate(tour_value = (appx_gdp_pct / 100) * gdp_value_agg) %>% 
  # fix data types
  dplyr::mutate(rgn_id = as.character(rgn_id)) %>% 
  # select relevant columns for joining
  dplyr::select(c(rgn_id, rgn_name, year, tour_value))

#View(nc_tourism_value)
#summary(nc_tourism_value)
```

## Tourism Value: China (including Macao and Hong Kong)

**Overview**

-   create China subset of GDP data: `china_gdp`
    -   filter to only include Hong Kong and Macao: `macao_hk_gdp`
-   join with `macao_hk_gf` (gapfilled, filtered gdp proportion subset with only Macao and Hong Kong): `macao_hk_join`
-   calculate value (will be in Constant 2015 US$): `macao_hk_value`

-   clean & tidy China gapfill dataset from National Bureau of Statistics of China
  -   pivot from wide to long (years columns --> year column), clean again
  -   convert values to standard from GDP dataset -- constant 2015 USD
-   join `china_usd_value` with `macao_hk_value`, drop irrelevant columns etc. and aggregate across `rgn_name` (China): call this `china_value`


### China subset of WB GDP


Remember that China subset we made out of the Macao and Hong Kong data? Now we're going to join that with a subset of the global country-level GDP data to find the tourism value for these areas. We will then join this data frame with the China tourism value data we read in earlier (which we will clean & tidy first) and finally aggregate the values to yield one value per year for China as a region.

```{r china_mo_hk_value}
# ---- filter to China regions GDP ----
china_gdp <- wb_gdp_clean %>% 
  dplyr::filter(stringr::str_detect(country_name, "China")) # depending on the dataset, you'd need to change this to "Macao", "Hong Kong"

# check to see if the names will work with OHI region synonyms
unique(china_gdp$country_name)
# yep! and they're named the same as Hong Kong and Macao are in macao_hk_subset

# Since we already have the Tourism Value for mainland China (which we'll clean and tidy in the chunks below), we'll filter out the GDP rows for China now and then bring the tourism values for China back in at the end before aggregating.

# ----- Prep for joining with Macao + Hong Kong subset ----
# drop "China" rows
macao_hk_gdp <- china_gdp %>% 
  dplyr::filter(country_name != "China") # don't need this, as our China data is already in tourism value format

# check
unique(macao_hk_gdp$country_name)


# ---- Join Macao + HK subsets ----
macao_hk_join <- dplyr::left_join(macao_hk_gf, macao_hk_gdp,
                           by = c("year", "geo_area_name" = "country_name"))


# ---- Calculate tourism value ----
macao_hk_value <- macao_hk_join %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(geo_area_name, year) %>% 
  dplyr::mutate(tour_value = (appx_gdp_pct / 100) * gdp_value) %>% 
  # select relevant columns
  dplyr::select(geo_area_name, year, tour_value)
# remember: the unit of this value is in constant 2015 US$ 
```

Plot to visualize Macao & Hong Kong Tourism Value

```{r}
# ---- Exploratory plot of Macao & Hong Kong tourism value ----
library(scales)
ggplot(data = macao_hk_value, 
       aes(x = as.numeric(year), 
           y = tour_value,
           color = geo_area_name)) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d(begin = 0.0, end = 0.7) +
  scale_x_continuous(breaks = c(date_range)) +
  # make y-axis more legible
  scale_y_continuous(labels = scales::label_currency(scale = 0.000000001,
                                                     suffix = "B",
                                                     # $30.00B --> $30B
                                                     scale_cut = cut_long_scale())) +
  theme_bw() +
  # update labels
  labs(title = paste0("Tourism Value in Hong Kong & Macao ", "(",
                      start_year,
                      "-", stop_year, ")"),
       subtitle = "Data Sources: UNWTO, World Bank",
       x = "Year",
       y = "Tourism Value (Constant 2015 US$)",
       color = "Region") +
  # customize theme elements
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.9),
    axis.title.x = element_text(margin = margin(t = 1.2, b = 0.2, unit = "lines")),
    axis.title.y = element_text(margin = margin(l = 0.5, r = 0.5, unit = "lines")),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(margin = margin(t = 0.5, b = 0.5, unit = "lines")),
    plot.subtitle = element_text(margin = margin(b = 1, unit = "lines"))
  )
```


### China Tourism Value
Overview: 

-   clean & tidy China Tourism Value dataset 

    -   clean column names
    -   select relevant categories (earnings from international and domestic tourism) and columns (indicators, year columns)
    -   pivot data from wide to long so that `tour(ism)_value` observations are in one column and `year` is in another columns (each row should be 1 observation -- with indicator, year, and earnings)
    -   clean up the new `year` column
    -   filter to date range specified earlier (`start_year` and `stop_year`)

-   convert tourism value (earnings) to a standardized format (we'll go over this in more detail later)
-   prep for joining with China subset (Macao + Hong Kong) of tourism value data 

```{r}
# ---- clean names ----
nbsc_clean <- nbsc_raw %>% 
  janitor::clean_names() 

# get strings for subcategories of tourism data we're looking for (international and domestic tourism earnings)
#unique(nbsc_clean$indicators)
tourism_categories <- c("Foreign Exchange Earnings from International Tourism(USD million)",
                        "Earnings from Domestic Tourism(100 million yuan)")
# here we can see that these values are in different units. 
# We'll want to standardize them to USD, then to Constant 2015 US$ (what the GDP data is in)

# but first, let's tidy up the data! 

# ---- tidy ----
nbsc_tidy <- nbsc_clean %>% 
  dplyr::filter(indicators %in% c(tourism_categories)) %>% 
  tidyr::pivot_longer(cols = x2023:x2004, # update this depending on year range of newly downloaded data
               names_to = "year", values_to = "tour_value") %>% 
  # tidy up year column
  dplyr::mutate(year = as.numeric(str_remove_all(year, 'x'))) %>%  
  # filter to date range defined earlier
  dplyr::filter(year %in% c(date_range))
```

#### Standardize values 

Tried to get the conversion rates in a reproducible way:

- `quantmod::getFX()` documentation: <https://www.rdocumentation.org/packages/quantmod/versions/0.4.26/topics/getFX> -- unfortunately only has rates for the past 180 days (doesn't work for our situation)
- `priceR::historical_exchange_rates()`: requires you to set up an API key -- the free one only gives you 100 requests per month and also requires you to enter credit card information even for the free account...
- documentation/repo for {priceR} : <https://github.com/stevecondylios/priceR?tab=readme-ov-file>


```{r test-conversion-packages, eval=FALSE}
# ---- quantmod ----
# library(quantmod)
# cny_rates_test <- quantmod::getFX("CNY/USD" #, from = paste0(start_year, "-01-01"), to = paste0(stop_year, "-12-31")
#                                   )
# cny_rates_test

# ---- priceR ----
# #tried using the API key posted by the developer
#Sys.setenv("EXCHANGERATEHOST_ACCESS_KEY"="7e5e3140140bd8e4f4650cc41fc772c0")
## get Chinese Yuan to USD conversion rates for time range
# cny_conversion_rate <- priceR::historical_exchange_rates(from = "CNY", to = "USD",
#                           start_date = paste0(start_year, "-01-01"), end_date = paste0(stop_year, "-12-31"))

## returned the following error:
# Error in `pmap()`:
# ℹ In index: 1.
# Caused by error in `dat[[8]]`:
# ! subscript out of bounds
# Backtrace:
#  1. priceR::historical_exchange_rates(...)
#  2. purrr::pmap_dfr(...)
#  3. purrr::pmap(.l, .f, ...)
#  4. purrr:::pmap_("list", .l, .f, ..., .progress = .progress)
#  8. priceR (local) .f(start_date = .l[[1L]][[i]], end_date = .l[[2L]][[i]], ...)
#  9. dat[[8]] %>% length
```

After trying and failing to get a (truly) free, open-source, reproducible way to get conversion rates, we ultimately decided to use the conversion rate of 100 Million Chinese Yuan (CNY) to USD, which at a rate of 1 CNY to 1 USD of 0.137781 as of July 11, 2024 is: 13,778,100.00 (USD)
- sources: [wise](https://wise.com/us/currency-converter/cny-to-usd-rate?amount=100000000), [Google Finance](https://www.google.com/finance/quote/CNY-USD?sa=X&ved=2ahUKEwiwt9CxgaCHAxWKle4BHdMZCxcQmY0JegQIBxAp)


According to the US Bureau of Labor Statistics, \$1 USD in June 2024 has the same buying power as \$0.76 in June of 2015 [conversion link & source](https://data.bls.gov/cgi-bin/cpicalc.pl?cost1=1.00&year1=202406&year2=201506) (they didn't have data for July of this year, and thankfully the 1 month average of CNY to USD was relatively stable around 0.1378). This means that once we convert the historic CNY to 2024 USD, we can convert those 2024 USD values back to the Constant 2015 USD (which the rest of our GDP data is in) using this conversion rate. Alternatively, we could use the `priceR::adjust_for_inflation()` function. However, the cny_to_usd rate is hard-coded, so for now we'll continue to use the hard-coded approach for adjusting the CNY --> USD current --> USD 2015 values. We will use the `priceR::adjust_for_inflation()` to adjust the USD current --> USD 2015 constant values.


```{r standardize_china_value}
# 100 million yuan = 13,778,100.00 (USD)
h_mil_cny_to_usd <- 13778100 # update this appropriately, or ideally find a reproducible conversion rate method and implement

# convert Chinese Yuan to USD
nbsc_usd <- nbsc_tidy %>% 
  dplyr::mutate(tour_value = case_when(
    indicators == tourism_categories[1] ~ tour_value * 1000000, # USD million --> USD (1)
    indicators == tourism_categories[2] ~ tour_value * h_mil_cny_to_usd #(CNY hundred million --> USD (1))
  ))

# now we have domestic tourism value in 2024 USD and 
# international tourism in current USD ($ year = year column value in data frame)

# ---- adjust for inflation, standardize to 2015 USD ----
nbsc_usd_adjusted <- nbsc_usd %>% 
  # optional: tidy up indicator names
  # mutate(indicators = case_when(
  #   indicators == tourism_categories[1] ~ "Foreign Exchange Earnings from International Tourism",
  #   indicators == tourism_categories[2] ~ "Earnings from Domestic Tourism"
  dplyr::mutate(usd = case_when(
    # international: convert Current USD to 2015 Constant US$ 
    indicators == tourism_categories[1] ~ priceR::adjust_for_inflation(price = tour_value, from_date = year, country = "US", to_date = 2015),
    
    # domestic: (2024 USD --> 2015 USD)
    indicators == tourism_categories[2] ~ priceR::adjust_for_inflation(price = tour_value,
                                                                       from_date = current_year, # or hard code this to whatever year's conversion rate you're using for CNY to USD
                                                                       country = "US", to_date = 2015)
    
  ))


# Check: 2015 years should match up
int_2015 <- nbsc_usd_adjusted %>% 
  dplyr::filter(year == 2015, indicators == tourism_categories[1])

int_2015$tour_value == int_2015$usd


# now that everything is in the same units (Constant 2015 US$), we can go ahead and aggregate the data to get a combined tourism value from international and domestic tourism earnings.

# ---- aggregate ----
china_usd_value <- nbsc_usd_adjusted %>% 
  # for each year,
  dplyr::group_by(year) %>% 
  # sum values !
  dplyr::summarize(usd = sum(tour_value, usd, na.rm = T)) %>% 
  # add descriptive columns for joining
  dplyr::mutate(geo_area_name = "China") %>% 
  # order columns
  dplyr::select(geo_area_name, year, usd)

```

#### Plot China Tourism Value

```{r}
library(ggplot2) # to avoid namespacing
ggplot(data = china_usd_value, 
       aes(x = as.numeric(year), 
           y = usd)) +
  # pulled hexcode from Chinese flag
  geom_point(color = "#EE1C25") +
  geom_line(color = "#EE1C25") +
  scale_x_continuous(breaks = c(date_range)) +
  # make y-axis more legible
  scale_y_continuous(labels = scales::label_currency(scale = 0.000000001,
                                                     suffix = "B"
                                                     # $30.00B --> $30B
                                                     # scale_cut = scales::cut_long_scale()
  )) +
  # set base themex
  theme_bw() +
  # adjust labels
  labs(title = paste0("Tourism Value in mainland China ", "(",
                      start_year,
                      "-", stop_year, ")"),
       subtitle = "Data Sources: UNWTO, World Bank",
       x = "Year",
       y = "Tourism Value (Constant 2015 US$)") +
  # tidy up aesthetic elements
  theme(
    axis.text.x = element_text(angle = 40, vjust = 0.8, hjust = 0.9),
    axis.title.x = element_text(margin = margin(t = 1.2, b = 0.2, unit = "lines")),
    axis.title.y = element_text(margin = margin(l = 0.5, r = 0.5, unit = "lines")),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(margin = margin(t = 0.5, b = 0.5, unit = "lines")),
    plot.subtitle = element_text(margin = margin(b = 1, unit = "lines"))
  )
```

Here we can see that China's earnings from tourism far exceed any of the values we've seen before, surpassing 1.6 Trillion by 2019. These numbers will be slightly higher once we add on the values for Macao and Hong Kong (likely no more than an added 50 Billion or so).

#### China Join

Now we can join this newly created China tourism value data frame with the `macao_hk_value` data frame of tourism value (in Constant 2015 USD). Then, we'll run `name_2_rgn` and aggregate values to get combined tourism values for China as an OHI region. We'll call this `china_agg`.

```{r}
# ---- join ----
china_value_join <- dplyr::full_join(x = china_usd_value, y = macao_hk_value,
                                     by = c("geo_area_name", "year",
                                            "usd" = "tour_value"))


# ---- run ohicore name_2_rgn function ----

china_rgn <- ohicore::name_2_rgn(
  df_in = china_value_join, 
  fld_name = "geo_area_name",
  flds_unique = c("year")
)

# as expected, we get a "DUPLICATES found" warning. Now that the rgn_name and rgn_id have been added to our data frame, we can use those to aggregate by region

# ---- aggregate for OHI region cohesion ----
china_agg <- china_rgn %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(rgn_name, rgn_id, year) %>% 
  dplyr::summarize(tour_value = sum(usd, na.rm = T)) %>% 
  dplyr::mutate(rgn_id = as.character(rgn_id)) # fix data type

```

Great! Now we're ready to reunite this china tourism value data with the rest of our tourism value data :) 

### Full Tourism Value

```{r}
# ---- joining china + no china dfs ----
tourism_value <- dplyr::full_join(nc_tourism_value, china_agg,
                           by = join_by(rgn_id, rgn_name, year, tour_value))

#View(tourism_value)
#summary(tourism_value)
```


#### Exploratory plots

**Static plot**

```{r}
library(ggplot2) # so we don't have to namespace
# ---- preliminary plotting: static ----
# make subset
tour_val_subset <- tourism_value %>% 
  filter(!is.na(tour_value))

# preliminary plot to check this out!
ggplot(data = tour_val_subset, 
       aes(x = as.numeric(year), 
           y = tour_value,
           color = rgn_id)) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d() +
  scale_x_continuous(breaks = c(seq(min(tour_val_subset$year),
                                    max(tour_val_subset$year), 1))) +
  scale_y_continuous(labels = scales::label_currency(scale = 0.000000001, suffix = "B")) +
  theme_bw() +
  labs(title = paste0("Tourism Value ", "(",
                      min(tour_val_subset$year),
                      "-", max(tour_val_subset$year), ")"),
       subtitle = "Data Sources: UNWTO, World Bank",
       x = "Year",
       y = "Tourism Value (Constant 2015 US$)") +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.9),
    axis.title.x = element_text(margin = margin(t = 1.2, b = 0.2, unit = "lines")),
    axis.title.y = element_text(margin = margin(l = 0.5, r = 0.5, unit = "lines")),
    panel.grid.minor.x = element_blank()
  )
```


We can see that the scale of tourism value in China dwarfs all other countries' values. This is obviously a fairly rudimentary (and not very helpful or informative) plot.
It is tedious to have to refer back to a table of the region IDs to understand which region is which, and is overall what I would call a "spaghetti plot". 
Let's use an interactive plot to more usefully visualize what's going on with the data in a subset without the USA.


**Interactive plot**

```{r scrapped-china-usa-plot, eval=FALSE}
# ---- exploratory plotting: interactive ----

no_china_value <- tour_val_subset %>% 
  dplyr::filter(!rgn_name %in% c("China")) %>% 
  dplyr::ungroup()

# ---- China subset ----
# with China 
# chn_usa <- tourism_value %>% 
#   filter(rgn_name %in% c("China"
#                          #, "United States" # filtered out earlier due to lack of data
#                          )) %>% 
#   ungroup() # lines have a tendency to not show up in plotly if the data is grouped
# 
# # without China or USA
# no_chn_usa <- tourism_value %>% 
#   filter(!rgn_name %in% c("China"
#                           #, "United States"
#                           )) %>% 
#   ungroup()
# 
# # ---- create updated interactive plot ----
# chn_plotly <- plotly::plot_ly(data = chn_usa,
#                                   x = ~year,
#                                   y = ~tour_value,
#                                   color = ~rgn_name,
#                                   type = "scatter", mode = "lines") %>% 
#   layout(title = "Tourism Value",
#          margin = m,
#          xaxis = list(title = "Year"),
#          yaxis = list(title = "Tourism Value (2015 US$)")
#   ) 
# 
# chn_plotly


# ---- Plot wihtout China ----
no_china_plotly <- plotly::plot_ly(data = no_china_value,
                                     x = ~year,
                                     y = ~tour_value,
                                     color = ~rgn_name,
                                     type = "scatter", mode = "lines") %>% 
  plotly::layout(title = "Tourism Value",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Value (2015 US$)")
  ) 

no_china_plotly
```


```{r}
# check max value
max_tour_val <- max(tourism_value$tour_value, na.rm = TRUE)

tourism_value %>% 
  dplyr::filter(tour_value == max_tour_val)

prettyNum(max_tour_val,big.mark = ",", scientific = FALSE)

# 1.69 trillion in China in 2019 out of 14.28 trillion total GDP
```

# Coastal Tourism

## Finding coastal tourism values from total tourism values

-   read in data on coastal population counts and total population per region 
-   create dataframe of proportion of coastal population per region
-   multiply percentage decimal by total tourism value to get approximation for coastal tourism 


```{r}
# read in coastal population time series data from mar_prs_population v2021 dataprep script
coastal_pop_count <- read_csv(here::here("globalprep", "mar_prs_population", "v2021", "output", "mar_pop_25mi.csv"))

# read in total population time series data from the UN, created in mar_prs_population v2021 
UN_total_pop <- read_csv(here::here("globalprep", "mar_prs_population", "v2021", "output", "UN_population.csv"))

# ---- filter data to date_range (2008-2019 in v2024) ----
coastal_pop_filter <- coastal_pop_count %>% 
  filter(year %in% c(date_range)) %>% 
  rename(coastal_pop = popsum)

total_pop_filter <- UN_total_pop %>% 
  filter(year %in% c(date_range)) %>% 
  rename(total_pop = population)

# note: the UN total pop and coastal pop dfs seem to have some missing regions

pop_rgn_diff <- dplyr::setdiff(coastal_pop_filter$rgn_id, total_pop_filter$rgn_id)

print(paste("Number of regions that don't occur in both datasets (only 1 set):", length(pop_rgn_diff)))

print(paste("Number of unique regions in Coastal Pop dataset:", length(unique(coastal_pop_filter$rgn_id))))

print(paste("Number of unique regions in UN Total Pop dataset:", length(unique(total_pop_filter$rgn_id))))

# ---- join dataframes ----
coastal_pop_join <- dplyr::full_join(coastal_pop_filter, total_pop_filter, by = c("rgn_id", "year"))

# ---- calculate proportion of coastal pop ----
coastal_prop_pct <- coastal_pop_join %>% 
  mutate(coastal_pct = (coastal_pop / total_pop),
         coastal_prop = (coastal_pct / 100)) # proportion multiplier

zero_coastal <- coastal_prop_pct[coastal_prop_pct$coastal_pop == 0, ]
unique(zero_coastal$rgn_id)

# adjust coastal population proportion and percentage multiplier 
# if coastal_pop != 0 and total_pop is NA, set coastal_prop to 1 (assumes that population of small island regions is all coastal)
coastal_pop_adj <- coastal_prop_pct %>% 
  mutate(coastal_prop = case_when(
    coastal_pop != 0 & is.na(total_pop) ~ 1,
    .default = coastal_prop
  ))


# uninhabited OHI regions
uninhabited_rgns <- read_csv(here::here("globalprep", "spatial", "v2017", "output", "rgn_uninhabited_islands.csv"))

# to join with tourism_value
uninhabited_rgns_ch <- uninhabited_rgns %>% 
  mutate(rgn_id = as.character(rgn_id))

# check if uninhabited regions have any pop data
test_pop <- dplyr::inner_join(coastal_pop_adj, uninhabited_rgns, by = "rgn_id")

tour_uninhabited_check <- dplyr::inner_join(tourism_value, uninhabited_rgns_ch, by = "rgn_id")
#View(tour_uninhabited_check) # df is empty! makes sense because uninhabited regions wouldn't have governments to report tourism values... this means we can also antijoin our coastal population data with the uninhabited regions dataset

# ---- remove uninhabited regions ----
coastal_multipliers <- coastal_pop_adj %>% 
  dplyr::anti_join(uninhabited_rgns, by = "rgn_id") %>% 
  mutate(rgn_id = as.character(rgn_id)) # prep rgn_id column for joining with tourism_value df

```


Basic plot to visually inspect data
```{r}
coastal_pct_plot <- plotly::plot_ly(data = coastal_multipliers,
                                     x = ~year,
                                     y = ~coastal_pct,
                                     color = ~rgn_id, 
                                     type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  plotly::layout(title = "Coastal Population (% per region)",
                 margin = m, # set margins
                 xaxis = list(title = "Year"),
                 yaxis = list(title = "Coastal Population %"))

coastal_pct_plot
```

This seems pretty solid -- we'd expect coastal population % to be relatively stable over this time frame.


Now we can join the coastal population proportion dataframe with the tourism value dataframe to calculate the coastal tourism value.

```{r}
# ---- join coastal prop + tourism value ----
coastal_tourism <- dplyr::full_join(tourism_value, coastal_multipliers,
                                    by = join_by("rgn_id", "year"))

# ---- calculate coastal tourism value ----
coastal_tourism_value <- coastal_tourism %>% 
  mutate(coastal_tour_value = (tour_value * coastal_prop))

# making a clean subset so we can keep a full version of the dataframe for reference if needed
coastal_value_clean <- coastal_tourism_value %>% 
  select(rgn_id, rgn_name, year, coastal_tour_value) %>% 
  # drop United States because we already have their values and don't want duplicates
  filter(!rgn_name %in% c("United States"))
```

Basic plot to visually inspect data
```{r}
coastal_val_plot <- plotly::plot_ly(data = (coastal_value_clean %>% ungroup()),
                                     x = ~year,
                                     y = ~coastal_tour_value,
                                     color = ~rgn_name, 
                                     type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  plotly::layout(title = "Coastal Tourism Value",
                 margin = m, # set margins
                 xaxis = list(title = "Year"),
                 yaxis = list(title = "Tourism Value (constant 2015 USD)"))

coastal_val_plot
```



## Coastal Tourism: The United States of America

The data currently includes states that border/contain the Great Lakes, so we'll need to filter those out. To do that, first we'll clean the data and select tourism sector GDP data. Then we'll make two subsets: one that contains only the great lakes states and one that only contains the total coastal aggregate values. We can then subtract the great lakes subset values from the total to obtain the true coastal tourism GDP values.

```{r}
# ---- Preliminary cleaning & tidying ----
enow_usa_clean <- enow_usa_raw %>% 
  janitor::clean_names() %>% 
  dplyr::select(geo_name, geo_scale, year, ocean_sector, ocean_industry, gdp, real_gdp) %>% 
  dplyr::filter(ocean_sector %in% c("Tourism and Recreation")) %>% 
  dplyr::filter(ocean_industry %in% c("Total"))

# check for year of inflation adjustment real_gdp 
enow_usa_clean %>% dplyr::filter(gdp == real_gdp) %>% head()

# now we know that it has been adjusted to 2012!
adjusted_year <- 2012 # update this if the adjusted date changes

# ---- Subsetting: Coastal USA and Great Lakes states ----

# Coastal total subset 
coastal_total <- enow_usa_clean %>% 
  dplyr::filter(geo_name %in% c("Coastal USA")) %>% # alternatively, you could filter to geo_scale == "National"
  # remove redundant columns after double checking that we filtered properly earlier
  dplyr::select(-c(geo_scale, ocean_sector, ocean_industry))

# Great Lakes subset
great_lakes <- enow_usa_clean %>% 
  dplyr::filter(geo_name %in% c("Illinois", "Indiana",
                         "Michigan", "Minnesota",
                         "Ohio", "Wisconsin",
                         "Pennsylvania")) %>% # excluding  "New York" because it is also coastal
  dplyr::select(-c(geo_scale, ocean_sector, ocean_industry))


# ---- Aggregate: get great lakes total tourism value ----
great_lakes_total <- great_lakes %>% 
  dplyr::group_by(year) %>% 
  dplyr::summarize(gl_gdp = sum(gdp),
            gl_real_gdp = sum(real_gdp))

# ---- Calculate true coastal tourism value ----
coastal_lakes_join <- dplyr::full_join(coastal_total, great_lakes_total,
                                       by = "year")

# subtract great lakes values from coastal total values
true_coastal_value <- coastal_lakes_join %>% 
  mutate(coastal_tour_gdp = (gdp - gl_gdp),
         coastal_tour_real_gdp = (real_gdp - gl_real_gdp))

# check:
if (sum((true_coastal_value$coastal_tour_gdp == (true_coastal_value$gdp - true_coastal_value$gl_gdp)) / nrow(true_coastal_value))) {
  print("Subtraction successful!")
} else {
  print("Subtraction failed. New column is not the difference between expected columns. Please double-check your methods.")
}


# ---- Adjust for inflation & Standardize ----
# if we're not fully confident that the dataset's inflation adjustment matches the inflation adjustment methods that we use throughout this subgoal (priceR::adjust_for_inflation()), we can use the GDP column and convert to 2015 constant USD. We could also convert from 2012 adjusted to 2015 adjusted. The following code can be altered for either method.



# standardize inflation adjustment to 2015 constant USD for joining with global tourism value dataframe
usa_coastal_adj <- true_coastal_value %>% 
  dplyr::mutate(coastal_tour_value = priceR::adjust_for_inflation(
    price = coastal_tour_gdp, # or: coastal_tour_real_gdp
    from_date = year, # or: adjusted_year (set to 2012 earlier)
    country = "US",
    to_date = 2015))

# for gdp (not real_gdp) method: check
usa_check_subset <- usa_coastal_adj %>% 
  filter(year == 2015) %>% # or change to whichever year you've adjusted the data to for inflation
  select(year, coastal_tour_gdp, coastal_tour_value)

if (usa_check_subset$coastal_tour_gdp == usa_check_subset$coastal_tour_value) {
  print(paste("USA Coastal tourism values have been successfully adjusted to", usa_check_subset$year, "USD."))
} else {
  print("Adjustment for inflation did not match for the selected year. Please check that the data has been subsetted to the year of adjustment and try again.")
}


# ---- Prep for joining with full coastal tourism dataset ----
usa_coastal_prep <- usa_coastal_adj %>% 
  mutate(country_name = "United States of America") %>% 
  select(country_name, year, coastal_tour_value)

# run ohicore name_2_rgn
usa_coastal_rgn <- ohicore::name_2_rgn(usa_coastal_prep,
                                       fld_name = "country_name",
                                       flds_unique = "year")

# select columns for joining
# full because we haven't filtered to the standardized date range
usa_coastal_tourism_full <- usa_coastal_rgn %>% 
  select(rgn_id, rgn_name, year, coastal_tour_value)
```


Preliminary plot to inspect data

```{r}

ggplot(data = usa_coastal_tourism_full, 
       aes(x = year, y = coastal_tour_value)) +
  # old glory red hex code from flag
  geom_line(color = "#B31942") +
  # old glory blue hex code from flag
  geom_point(color = "#0A3161") + 
  # show_col(viridis::viridis(6))
  #scale_x_continuous(breaks = c(date_range)) +
  # make y-axis more legible
  scale_y_continuous(labels = scales::label_currency(scale = 0.000000001,
                                                     suffix = "B",
                                                     # $30.00B --> $30B
                                                     scale_cut = cut_long_scale())) +
  theme_bw() +
  # update labels
  labs(title = paste0("Coastal Tourism Value in the United States of America ", "(2005-2021)"),
       subtitle = "Data Source: ENOW",
       x = "Year",
       y = "Tourism Value (Constant 2015 US$)",
       color = "Region") +
  # customize theme elements
  theme(
    axis.text.x = element_text(margin = margin(t = 0.8, unit = "lines")),
    axis.title.x = element_text(margin = margin(t = 1.2, b = 0.2, unit = "lines")),
    axis.title.y = element_text(margin = margin(l = 0.5, r = 0.5, unit = "lines")),
    panel.grid.minor.x = element_blank(),
    plot.title = element_text(margin = margin(t = 0.5, b = 0.5, unit = "lines")),
    plot.subtitle = element_text(margin = margin(b = 1, unit = "lines")))
```

This seems to be fairly reflective of the $143 Billion value associated with US coastal tourism according to [NOAA](https://coast.noaa.gov/states/fast-facts/tourism-and-recreation.html), and the sharp fall in 2020 reflects the trend we expect to see due to COVID-19 Pandemic impacts on tourism globally.

## Final Coastal Tourism

Now, let's filter to the date range of the rest of the data and join to get a full coastal tourism value dataset!
```{r}
# filter USA data
usa_coastal_tourism <- usa_coastal_tourism_full %>% 
  filter(year %in% c(date_range)) %>% 
  mutate(rgn_id = as.character(rgn_id))

# optional: check that year filtering worked
#max(usa_coastal_tourism$year) == stop_year


# ---- join coastal tourism datasets ----
coastal_tourism_full <- dplyr::full_join(coastal_value_clean, usa_coastal_tourism,
                                         by = join_by("rgn_id", "rgn_name", "year", "coastal_tour_value")) %>% 
  ungroup()


```


```{r}
# preliminary plotting to investigate

coastal_tour_plot <- plotly::plot_ly(data = coastal_tourism_full,
                                     x = ~year,
                                     y = ~coastal_tour_value,
                                     color = ~rgn_name, 
                                     type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  plotly::layout(title = "Coastal Tourism Value",
                 margin = m, # set margins
                 xaxis = list(title = "Year"),
                 yaxis = list(title = "Tourism Value (constant 2015 USD)"))

coastal_tour_plot



# make non-USA subset
coastal_no_usa <- coastal_tourism_full %>% 
  filter(!rgn_name %in% c("United States"))


no_usa_coastal_plot <- plotly::plot_ly(data = coastal_no_usa,
                                     x = ~year,
                                     y = ~coastal_tour_value,
                                     color = ~rgn_name, 
                                     type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  plotly::layout(title = "Coastal Tourism Value",
                 margin = m, # set margins
                 xaxis = list(title = "Year"),
                 yaxis = list(title = "Tourism Value (constant 2015 USD)"))

no_usa_coastal_plot
```



## Implementing sector multipliers

There are actually no multipliers for the tourism sector. 


# Saving intermediate data 

Prep for saving & save to `int` folder

```{r}
#--------Prepping data for int saving -------

tour_revenue_int <- coastal_tourism_full %>% 
  rename(usd = coastal_tour_value) %>% 
  mutate(
    unit = "USD (1)",
    sector = "tour",
    # used in the inflation adjustment calculation downstream
    usd_yr = 2015 # because the tourism values are in "Constant 2015 US$"
  ) %>% 
  select(rgn_id, rgn_name, year, usd, unit, sector, usd_yr) %>% 
  ungroup()

#write_csv(tour_revenue_int, here::here(int_dir, "eco_tour_usd_pre.csv"))
```


