---
title: "OHI Economies Dataprep"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------


# Economies Dataprep (Measures the amount of revenue produced)

## Overview

-   **Data:**

-   Tourism: UNWTO Dept. of Statistics (UN Tourism)
-   <https://www.unwto.org/tourism-statistics/economic-contribution-SDG> 
  -  A big issue with this data set is that it does not have data for mainland China. I found a resource from CEIC data that has [monthly tourism revenue for China](https://www.ceicdata.com/en/indicator/china/tourism-revenue) (accessed July 11, 2024), but only goes up until the end of 2018 (as of July 2024). If this dataset is updated in the future, it could be a potential resource used to gapfill tourism value for China, which may allow us to preserve the notable Macao tourism data (though you'd need to aggregate the data by taking a weighted sum based on population or some other metric, which would ultimately obscure Macao's trends in the end...). In the future, I hope that we can incorporate Macao as its own region.
  



-  GDP: World Bank, in "Constant 2015 US$"
-   <https://data.worldbank.org/indicator/NY.GDP.MKTP.KD> Downloaded July 9, 2024. Has annual GDP value from 1960-2023. Additional metadata copied and saved as a `.rtf` in the WorldBank_global_annual_GDP_2015_constant_USD subfolder of the _raw_data WorldBank folder for d2024. Last updated (found in the first few rows of the `.csv`): 2024-06-28

### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  ohicore, # name_2_rgn
  here,
  countrycode,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo # for gapfilling
  
)
# for extra gapfilling (v2024: didn't use this method)
# remotes::install_github("skgrange/threadr") # for na_extrapolate
#library(threadr)


# source 
source(here("workflow", "R", "common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO raw data directory
unwto_dir <- here::here(raw_data_dir, "UNWTO", data_dir_version_year)

# World Bank raw data dir (for GDP data)
wb_dir <- here::here(raw_data_dir, "WorldBank", data_dir_version_year)

# China National Bureau of Statistics (NSB) directory
nbsc_dir <- here(raw_data_dir, "NBS_China", data_dir_version_year)

# Output data directory -- for intermediate data products
int_dir <- here(data_path, "int")
```


## Read in Data

```{r}
# ===================== Read in Data ==============================


# ---- Tourism revenue data ----
gdp_prop_raw <- readxl::read_xlsx(here(unwto_dir, "Indicator-8_9_1-2022-UN_Tourism_april2024_update.xlsx"))

# ---- OHI regions data ----
region_names <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  select(-Notes)

# ---- World Bank GDP in 2015 Constant USD data ----
wb_gdp_raw <- readr::read_csv(here(wb_dir, "WorldBank_global_annual_GDP_2015_constant_USD", "WorldBank_global_annual_GDP_2015_constant_USD.csv"), skip = 4)
# date range from 1960 to 2023


# ---- NBSC Data (different units for international and domestic tourism) ----
nbsc_raw <- readr::read_csv(here::here(nbsc_dir, "eco_tour_china_all_metrics_2004-2023.csv"), skip = 2)

```


## Tourism GDP % 

### Data cleaning

Current unit of "total" in dataframe:

- "Tourism direct GDP as a proportion of total GDP (%)" 

```{r}
# ---- preliminary cleaning/tidying ----
gdp_prop_clean <- gdp_prop_raw %>% 
  janitor::clean_names() %>% 
  select(c(geo_area_name, year = time_period, gdp_prop_percent = total))

length(unique(gdp_prop_clean$geo_area_name))

# make all geographic areas have the same date range (and fill missing values with NAs)
years_df <- tibble(geo_area_name = gdp_prop_clean$geo_area_name) %>% 
  group_by(geo_area_name) %>% 
  summarize(year = seq(min(gdp_prop_clean$year), max(gdp_prop_clean$year)))

# join with gdp prop data
gdp_prop_yrs <- left_join(years_df, gdp_prop_clean, by = c("geo_area_name", "year"))


```

Create preliminary plot of cleaned full dataset (not joined/filtered to OHI regions yet)
```{r}
# ---- preliminary plot ----
# define margins
m <- list(
  l = 80,
  r = 60,
  b = 80,
  t = 80,
  pad = 4
)

# ---- create plotly interactive plot ----
prelim_plot <- plot_ly(data = gdp_prop_yrs,
                       x = ~year,
                       y = ~gdp_prop_percent,
                       color = ~geo_area_name, 
                       type = "scatter", mode = "lines") %>% 
  # update layout: text, margins
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for All Geo Regions",
         margin = m, # set margins
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

# display plot
prelim_plot

# htmlwidgets::saveWidget(prelim_plot, file = "gdp_prelim_plot_all_years.html")
```
v2024:
Macao has an extremely high percent of tourism contributing to their GDP (max of 59.6% in 2023), and all geographic areas appear to have a sharp decrease in tourism contribution to GDP in 2020, which is to be expected due to the COVID-19 Pandemic. Macao saw a fall by more than half, from 52.57% in 2019 to 21.44% in 2020.


### Run OHI region name function

```{r}
# ---- run name_2_rgn -----
gdp_rgn <- name_2_rgn(df_in = gdp_prop_yrs, 
                      fld_name = 'geo_area_name',
                      flds_unique = c("year"))

# sorting through mismatch and aggregation ----

# v2024: note on Northern Mariana Islands and Guam OHI region: we only have data for Guam, not Northern Mariana Islands... this is problematic because in the World Bank GDP dataset, we only have GDP data for both the Northern Mariana Islands and Guam and need to aggregate them to fit into the OHI regions system. For now, we'll drop the Northern Mariana Islands GDP data in the WB GDP processing, but please note that if you get GDP % data for this region later on, you'll need to fix that in the GDP data processing chunk.

# v2024: China: we only have data on Hong Kong and Macao, no mainland China data...
# Mel told us to just drop them... which makes sense in context, but is unfortunate, as Macao had the highest values. I will look for data on mainland China to supplement if possible.

# check duplicates         # v2024: (none)
tour_eco_duplicates <- gdp_rgn[duplicated(gdp_rgn[, c("rgn_name", "year")]),]
unique(tour_eco_duplicates$geo_area_name)
unique(tour_eco_duplicates$rgn_name)

length(unique(gdp_rgn$rgn_name))
# v2024: 93

# ---- Rename Macao and Hong Kong ----
gdp_prop_agg <- gdp_prop_yrs %>% 
  mutate(geo_area_name = case_when(
    geo_area_name == "China, Macao Special Administrative Region" ~ "China, Macao SAR",
    geo_area_name == "China, Hong Kong Special Administrative Region" ~ "China, Hong Kong SAR",
    .default = geo_area_name
  ))


# ---- Run name_2_rgn again -----
gdp_rgn_agg <- name_2_rgn(df_in = gdp_prop_agg, 
                      fld_name = 'geo_area_name',
                      flds_unique = c("year"))

# great! now nothing has been removed... now we need to check the duplicates 
tour_agg_duplicates <- gdp_rgn_agg[duplicated(gdp_rgn_agg[, c("rgn_name", "year")]),]
unique(tour_agg_duplicates$geo_area_name)
unique(tour_agg_duplicates$rgn_name)


# ----- China subset ----
macao_hk_subset <- gdp_rgn_agg %>% filter(rgn_name == "China")


# ---- excluding China subset ----
# (we'll rbind later once we have the tourism values for HK, Macao, and China) 
no_mc_hk_subset <- gdp_rgn_agg %>% filter(rgn_name != "China")


```



Create preliminary plot with updated `rgn_name` field

```{r}
# ---- updated prelim plot ----
ohi_rgns_prelim_plot <- plot_ly(data = no_mc_hk_subset, x = ~year, y = ~gdp_prop_percent, color = ~rgn_name, 
                                type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for OHI Regions",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

ohi_rgns_prelim_plot


# htmlwidgets::saveWidget(ohi_rgns_prelim_plot, file = "eco_tour_ohi_rgns_prelim_plot.html")
```

It is notable how much the y-axis scale changes once we drop Macao. I think that the -0.1 value we see for Fiji could be indicative of the country spending money on tourism infrastructure, development, and advertisement but then not getting enough tourism revenue due to Covid, but this is worth investigating further (I couldn't find anything helpful in the metadata, unfortunately).



### China subset

Remember that China subset we made out of the Macao and Hong Kong data? Now we're going to join that with a subset of the global country-level GDP data to find the tourism value. We will then join this data frame with the China tourism value data we read in earlier (we'll need to clean it first).


### Gapfilling

More data cleaning & prep for gapfilling, then gapfilling

```{r}
# define maximum year (determined after visually inspecting data)
max_year <- 2019 # update this depending on how many regions we have data for -- v2024: set to 2019 due to huge data drop off starting in 2020, likely related to covid pandemic
min_year <- min(gdp_rgn$year)


# some geo areas are missing data for some years ----
years_df <- tibble(rgn_name = region_names$rgn_name,
                   rgn_id = region_names$rgn_id) %>% # copy region identifier columns
  group_by(rgn_name, rgn_id) %>% # for each geo region, 
  summarize(year = seq(min_year, max_year)) # have sequence of years from 2009-2019 
# (fragile and R warns you about preferring `reframe()`)

# join years df and tourism GDP df ----
gdp_year_join <- full_join(years_df, gdp_rgn, by = c("rgn_id", "rgn_name", "year"))


# tidy or fill in geo area names for NA years ----
gdp_years <- gdp_year_join %>% 
  select(-geo_area_name) # %>%  # alternatively, set years_df to have geo_area_name column, then perform the following steps to fill in geo area names for NA years
# group_by(rgn_name) %>% 
# fill(geo_area_name, .direction = "downup") %>% 
# ungroup()


# gapfilling ----
# estimate tourism GDP proportion for geo areas with missing data
gdp_filled <- gdp_years %>% 
  group_by(rgn_name) %>% 
  # interpolate (fill missing values between 2 values)
  mutate(appx_gdp_pct = zoo::na.approx(gdp_prop_percent, # using values in this column
                                       na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                       #  extrapolate using rule = 2 from approx(),
                                       # which uses closest data extreme to
                                       #  extrapolate for leading and trailing NAs
                                       rule = 2
  )) # %>% 


```
Note: after further discussing with Melanie Frazier, we decided to not include countries with only 1 year of observations in our dataset.



```{r, eval = FALSE}
# ---- Assess how many countries we are gapfilling more extremely for --------
# no longer useful, as we cut out this column to avoid over-gapfilling
# # find number of regions that only have 1 data point (wouldn't be filled by na.approx if only 1 datapoint, left as NAs)
# na_regions <- gdp_filled %>% 
#   filter(is.na(appx_gdp_fill))
# 
# # print which geo areas only had 1 datapoint
# #unique(na_regions$rgn_name)
# 
# # calculate proportion of 1 point-filled geo areas to total geo areas
# num_na <- length(unique(na_regions$rgn_name))
# num_tot <- length(unique(gdp_filled$rgn_name))
# 
# paste0("Proportion of geo areas with only 1 data point: ", round(((num_na / num_tot) * 100), 3), "%")
```

#### Preliminary Plotting

```{r}
# ---- plot of all geographic areas ---- 

# define margins
m <- list(
  l = 80,
  r = 80,
  b = 80,
  t = 80,
  pad = 4
)

# create plotly interactive plot
prelim_gf_plot <- plot_ly(data = gdp_filled, x = ~year, y = ~appx_gdp_pct, color = ~rgn_name,
                          type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism Contribution to GDP across all OHI Regions (2009-2019)",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%, Gapfilled)")
  ) 
# display plot
prelim_gf_plot

# note: before runing name_2_rgn, Macao was one of the top countries in terms of tourism contribution to GDP. After running the region function, Macao was dropped as there was no data for mainland China, and this interesting detail of the data became obscured.

# ---- make subset of geo areas with greatest GDP change ----
gdp_diff <- gdp_filled %>% 
  group_by(rgn_name) %>% 
  # find difference
  summarize(gdp_diff = (abs(max(appx_gdp_pct) - min(appx_gdp_pct)))) %>% 
  arrange(desc(gdp_diff)) # for easier visual inspection

head(gdp_diff)

gdp_top_diff <- gdp_diff %>% 
  slice_head(n = 15)

top_diff_areas <- c(gdp_top_diff$rgn_name)

gdp_filled_diff <- gdp_filled %>% 
  filter(rgn_name %in% c(top_diff_areas))


top_diff_gdp_plot <- plot_ly(data = gdp_filled_diff, x = ~year, y = ~appx_gdp_pct, color = ~rgn_name,
                             type = "scatter", mode = "lines") %>% 
  layout(title = "OHI Regions with the Greatest Difference in Tourism GDP (2009-2019)",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%, Gapfilled)")
  ) 

top_diff_gdp_plot


# htmlwidgets::saveWidget(top_diff_gdp_plot, file = "tour_gdp_top_diff_gf_plot.html")
```



## GDP: WorldBank annual GDP (Constant 2015 US$)


### Data cleaning


```{r}
# ---- cleaning WorldBank GDP data ----
wb_gdp_clean <- wb_gdp_raw %>% 
  clean_names() %>% 
  # remove unwanted columns
  select(-c(indicator_code, x69)) %>% 
  # pivot to put years in 1 column, values in another
  pivot_longer(cols = x1960:x2023, names_to = "year", values_to = "gdp_value") %>% 
  # tidy up year column
  mutate(year = as.numeric(str_remove_all(year, 'x'))) %>%  
  # filter to date range
  filter(year %in% c(2009:2019))
```


### Run OHI region name function

```{r}
# ---- run name_2_rgn -----
wb_gdp_rgn <- name_2_rgn(df_in = wb_gdp_clean, 
                         fld_name = "country_name",
                         flds_unique = c("year"))

# thankfully, in the list following this line: "These data were removed for not having any match in the lookup tables:", there are mostly just the WorldBank aggregate groups (i.e., "low & middle income, Latin America & Caribbean, etc.), with the exception of Channel Islands and Isle of Man

# COME BACK ---- split Channel Islands -- proportional to EEZ area, Jersey slightly larger than Guernsey
# Isle of Man -- distinct governing body, should not be aggregated with Ireland or the UK

# ---- identify duplicates ----
wb_gdp_duplicates <- wb_gdp_rgn[duplicated(wb_gdp_rgn[, c("rgn_name", "year")]),]
#View(wb_gdp_duplicates)
unique(wb_gdp_duplicates$country_name)
unique(wb_gdp_duplicates$rgn_name)

# then look up these & their pairs in the wb_gdp_clean df to see which countries we do or do not have data for to inform aggregation steps.

# [1] "China": we have data on Macao and Hong Kong --> aggregate (can sum) -- but we don't have tourism GDP % data for mainland China... so just leave as basic aggregation (sum) for now


# [2] "Northern Mariana Islands and Guam" --> have GDP data for Northern Mariana Islands, and Guam. Can aggregate this, but we only have tourism % GDP data for Guam (see earlier comment in GDP % data cleaning chunk). 
# For now, we'll just keep Guam GDP and filter out northern mariana islands GDP... ? (COME BACK: double check w Mel)


# [3] "Puerto Rico and Virgin Islands of the United States" --> have data for both separately --> aggregate



# ---- aggregate ----
# sum by rgn_name etc.
# name ~ WorldBank GDP region aggregated
wb_gdp_rgn_agg <- wb_gdp_rgn %>%
  filter(!country_name %in% c("Northern Mariana Islands")) %>% # filtering out NMI since we only have Guam GDP % data in our other dataset -- should revisit this and update once we have that data -- double check to see which of these regions we have data for on both datasets in future years of updates!!!
  group_by(rgn_id, rgn_name, year) %>%
  dplyr::summarize(gdp_value_agg = sum(gdp_value, na.rm = TRUE) 
  )%>%
  ungroup() %>% 
  # turn 0s into NAs (appropriate flag for lack of data, as earlier sum(na.rm = TRUE) step obscured NAs)
  mutate(gdp_value_agg = case_when(
    gdp_value_agg == 0 ~ NA,
    .default = gdp_value_agg
  ))

# ---- aggregation check ----
# to double check and see if Northern Marian Islands was properly filtered out:
agg_nmi_vals <- wb_gdp_rgn_agg %>% 
  filter(rgn_name %in% c("Northern Mariana Islands and Guam")) %>%
  select(year, gdp_value_agg)

guam_reference <- wb_gdp_clean %>% 
  filter(country_name %in% c("Guam")) %>% 
  select(year, gdp_value)

agg_nmi_check <- left_join(agg_nmi_vals, guam_reference)

print(paste("Do the Guam GDP values before aggregation match the 'Northern Mariana Islands and Guam' values after aggregation?"))

if (sum(agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value) == length(unique(agg_nmi_check$gdp_value_agg))
) {
  print("Yes")
} else {
  print("No")
}

# If "No", check which don't match
# print(paste((agg_nmi_check$gdp_value_agg == agg_nmi_check$gdp_value)))

```


Extra: aggregation explanation/example
```{r agg_example, eval = FALSE}
# to see why we're using summarize instead of just mutating and adding a column see this example:
# test <- wb_gdp_rgn %>%
#   filter(!rgn_name %in% c("Northern Mariana Island")) %>%
#   group_by(rgn_id, rgn_name, year) %>%
#   mutate(test = sum(gdp_value, na.rm = TRUE))
# 
# #View(test)
# test %>% filter(rgn_name %in% c("China"))

# yields value column, but still keeps the old country_name column, so there are still duplicates! 

```


## Find Tourism Value 

- join Tourism as % GDP with global GDP

- create proportion multiplier

- find USD value of tourism per country


```{r}
# ---- join GDP Value (constant 2015 US$) with Tourism GDP % ----
gdp_join <- full_join(gdp_filled, wb_gdp_rgn_agg, by = join_by(rgn_name, rgn_id, year))

# gdp_join_rgn <- name_2_rgn(df_in = gdp_join, 
#                        fld_name = "rgn_name",
#                        flds_unique = c("year"))

# ---- calculate tourism value ---- 
tourism_value <- gdp_join %>% 
  # get proportion to use as multiplier
  mutate(gdp_filled_prop = appx_gdp_pct / 100) %>% 
  ungroup() %>% 
  group_by(rgn_id, rgn_name, year) %>% 
  mutate(tour_value = gdp_filled_prop * gdp_value_agg) %>% 
  # fix data types
  mutate(rgn_id = as.character(rgn_id))

```


#### Preliminary Plotting

**Static plot**

```{r}
# ---- preliminary plotting: static ----
# make subset
tour_val_subset <- tourism_value %>% 
  filter(!is.na(tour_value))

# preliminary plot to check this out!
ggplot(data = tour_val_subset, 
       aes(x = as.numeric(year), 
           y = tour_value,
           color = rgn_id)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = c(seq(min(tour_val_subset$year),
                                    max(tour_val_subset$year), 1))) +
  scale_y_continuous(labels = scales::label_currency(scale = 0.000000001, suffix = "B")) +
  theme_bw() +
  labs(title = paste0("Tourism Value ", "(",
                      min(tour_val_subset$year),
                      "-", max(tour_val_subset$year), ")"),
       subtitle = "Data Sources: UNWTO, World Bank",
       x = "Year",
       y = "Tourism Value (Constant 2015 US$)") +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.9),
    axis.title.x = element_text(margin = margin(t = 1.2, b = 0.2, unit = "lines")),
    axis.title.y = element_text(margin = margin(l = 0.5, r = 0.5, unit = "lines")),
    panel.grid.minor.x = element_blank()
  )
```

This is obviously a very rudimentary (and not very helpful or informative) plot. It is  tedious to have to refer back to a table of the region IDs to understand which region is which, and is overall what I would call a "spaghetti plot". Let's use an interactive plot to more usefully visualize what's going on with the data!


**Interactive plot**

```{r}
# ---- preliminary plotting: interactive ----

# set margins
m <- list(
  l = 80,
  r = 80,
  b = 80,
  t = 80,
  pad = 4
)

# ---- create updated interactive plot ----
# COME BACK -- not sure why plot lines aren't showig
tour_val_plotly <- plot_ly(data = tour_val_subset,
                           x = ~year,
                           y = ~tour_value,
                           color = ~rgn_name,
                           type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism Value",
         margin = m,
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Value (2015 US$)")
  ) 

tour_val_plotly
```

COME BACK -- lines not showing up....




```{r}
# check max value
max_tour_val <- max(tourism_value$tour_value, na.rm = TRUE)

tourism_value %>% 
  filter(tour_value == max_tour_val)

prettyNum(max_tour_val,big.mark = ",", scientific = FALSE)

# 582 billion in US in 2019... out of 20 Trillion GDP
```


# Saving intermediate data 

```{r}
#--------Prepping data for int saving -------

tour_revenue_int <- tourism_value %>% 
  rename(usd = tour_value) %>% 
  mutate(
    unit = "USD (1)",
    sector = "tour",
    usd_yr = 2015 # because the tourism values are in "Constant 2015 US$"
  ) %>% 
  select(rgn_id, rgn_name, year, usd, unit, sector, usd_yr) %>% 
  ungroup()

#write_csv(tour_revenue_int, here(int_dir, "eco_tour_usd_pre.csv"))
```




### Join with OHI regions

COME BACK -- ask about joining w OHI rgns

```{r}

# # tidy up regions dataset ----
# region_names <- region_names_raw %>% select(-Notes)
# 
# 
# # join regions & tourism GDP prop ----
# gdp_regions_join <- left_join(region_names, gdp_filled, by = c("eez_iso3" = "iso3"))
#   
# 
# head(gdp_regions_join)
# str(gdp_regions_join)

```



