---
title: "OHI `r format(Sys.Date(), '%Y')` - CF ECO Revenue Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

# CF Economies: USD Value of Marine (Commercial) Fishing per Country/Region per Year

## Overview

-   FAO Capture Data 

    Data Source FAO Global Capture Production (in metric tonnes)

    This version of the value database was downloaded from the
    Statistical Query Panel. Data from [FAO Global Capture
    Production](https://www.fao.org/fishery/en/collection/capture?lang=en)

    Citation: FAO 2023. Global Capture Production. Fisheries and
    Aquaculture Division
    <https://www.fao.org/fishery/en/collection/capture?lang=en>

    Source information: Navigate to the [online query
    portal](https://www.fao.org/fishery/statistics-query/en/capture/capture_quantity)
    for FAO Global Capture Production Quantity. Deselect all
    pre-selected years. Drag these fields into selected rows: ASFIS
    species name, FAO major fishing area name, ASFIS species ISSCAP
    group name En. ASFIS species Family scientific name, FAO major
    fishing areas, Inland/Marine areas Name en. Click on show data and
    confirm that data is present for 1950- two years prior to current
    year. Click download and select yes to include Null Values.

    Date: September 15th, 2023

    Time range: 1950-2021

    Native data resolution: Country level

    Format: csv

    Description: Global Capture Production Quantity

-   Ex-Vessel Price Data (downloaded August 24, 2023)

    -   Ex-vessel price data is in USD/metric tonne.

        -   ex-vessel-prices: ex-vessel prices for fishery caught
            species from 1976-2019

            -   exvessel_price_database_1976_2019.csv: ex-vessel price
                data gathered from [Melnychuk et al.
                2016](https://doi.org/10.1093/icesjms/fsw169) and
                updated to 2019 using methods described in the
                public-facing [github
                repo](https://github.com/SFG-UCSB/price-db-sfg)
                associated with the Melnychuk et al. 2016 paper

        -   \*\*Citation for paper\*\*  

        -   Melnychuk, M. C., Clavelle, T., Owashi, B., and
            Strauss, K. 2016. Reconstruction of global ex-vessel prices
            of fished species. - ICES Journal of Marine Science.
            doi:10.1093/icesjms/fsw169.
            
            
### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo, # for gapfilling
  pdftools,
  ohicore,
  ggplot2
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr) # for gapfilling, if necessary


# sources! 
source(here("workflow/R/common.R")) # file creates objects to process data

source(here('workflow/R/fao_fxn.R')) # fxn for cleaning old version of FAO data 

source(here('workflow/R/fao_online_portal_clean.R')) # fxn for obtaining FAO data

# set year and file path info
current_year <- 2024 # Update this in the future!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# ILO raw data directory
ilo_dir <- here(raw_data_dir, "ILOSTAT", data_dir_version_year)

# FAO raw data directory
fao_dir <- here(raw_data_dir, "FAO", data_dir_version_year)

# OECD raw data directory
oecd_dir <- here(raw_data_dir, "OECD", data_dir_version_year)

# FAO capture raw data directory
fao_capture_dir <- here(raw_data_dir, "FAO_capture", data_dir_version_year)

# Ex-vessel raw data directory
ex_vess_dir <- here(raw_data_dir, "ex-vessel-price-database-updated")

# output data dir for intermediate data products
int_dir <- here(data_path, "int")
```

### Read in Data

```{r}
# FAO Capture raw data
fao_capture_raw <- read_csv(here(fao_capture_dir, "Global_capture_production_Quantity_1950-2021.csv"))

# Ex-Vessel raw data, from EmLab, who projected Melnycuk's data (see README)
exvessel_prices_raw <- read_csv(here(ex_vess_dir, "price-db-results", "exvessel_price_database_1976_2019.csv"))

# read in cf revenue data from v2023, using FAO Capture data and Ex-vessel data
cf_revenue <- read_csv(here("globalprep", "le", "v2023", "int", "cf_revenue.csv"))
```

### Run ohicore name_2_rgn
```{r}
# check how many countries are available from the v2023 .csv before running names_2_rgn
length(unique(cf_revenue$country)) #203

# run names_2_rgn, which identifies country names from the cf_revenue and determines which OHI region it belongs to
cf_revenue_rgn <- ohicore::name_2_rgn(df_in = cf_revenue, 
                             fld_name='country',
                             flds_unique = c("year"))

# to determine the correct OHI regions and compare the function's actions:
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% select(-Notes)
```

When names_2_rgn has been run, there are some countries from cf_revenue that are removed due to not having any match in the lookup tables.  This includes: bonaire/s.eustatius/saba, british indian ocean ter, channel islands, french southern terr, isle of man, netherlands antilles, other nei, saint barthélemy, saint helena/asc./trist., saint-martin (french), yugoslavia sfr.

Some could be considered OHI regions, but they are aggregated within another name.  This includes: Bonaire/S.Eustatius/Saba, which can be split into Bonaire, Sint Eustatius, and Saba; and the Channel Islands, which should be split into Jersey and Guernsey.

Others can be dropped because they do not belong to an OHI region.  This includes yugoslavia sfr, saint barthélemy, other nei, isle of man (self-governed), and french southern terr.  

```{r}
# fix the country names so they can be recognized by names_2_rgn
cf_revenue_fix <- cf_revenue %>%
  mutate(country = case_when(
    country=="British Indian Ocean Ter" ~ "British Indian Ocean Territory",
    country=="Saint Helena/Asc./Trist." ~ "Saint Helena", 
    country=="Saint-Martin (French)" ~ "Northern Saint-Martin",
    country=="Netherlands Antilles" ~ "Curacao", #Sint Maarten has already been covered in a separate country name, so this must be referring to Curacao
    country=="Puerto Rico" ~ "Puerto Rico and Virgin Islands of the United States", #to ensure the fxn aggregated correctly
    country=="US Virgin Islands"~"Puerto Rico and Virgin Islands of the United States", #to ensure the fxn aggregated correctly
    country=="China, Hong Kong SAR" ~ "China", #to ensure the fxn aggregated correctly
    country=="China, Macao SAR" ~ "China", #to ensure the fxn aggregated correctly
    country=="United Republic of Tanzania, Zanzibar" ~ "Tanzania", #to ensure the fxn aggregated correctly
    country=="Tanzania, United Rep. of" ~ "Tanzania", #to ensure the fxn aggregated correctly
    TRUE ~ country # Everything else, leave it be
  ))
```

Re-run names_2_rgn after fixing names
```{r}
# re-run names_2_rgn, check which countries are still dropped
cf_revenue_rgn_fix <- ohicore::name_2_rgn(df_in = cf_revenue_fix, 
                             fld_name='country',
                             flds_unique = c("year"))
```


## Disaggregation 

**Regions to disaggregate:**

Channel Islands --> Guernsey (228) and Jersey (227)

Bonaire/S.Eustatius/Saba --> Bonaire (245), Sint Eustatius (249), and Saba (248)

To do this, look at the rasterized fishing effort calculated in 2017 within the FIS subgoal, to determine the relative proportion of mean_catch within each region's eez. Afterwards, distribute revenue relative to that proportion.

```{r}
# load in data on the mean catch for each region, used as a proxy to determine which region has a greater commercial fishing economy
mean_catch_2017 <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2024/gh-pages/globalprep/fis/v2020/int/mean_catch.csv")

# function to calculate average catch for certain regions and certain years
calculate_avg_catch <- function(data, region_ids, years) {
  data %>% 
    select(-stock_id_taxonkey) %>% 
    filter(rgn_id %in% region_ids, year %in% years) %>% # for the specified region id and year:
    group_by(rgn_id, year) %>%
    summarize(catch = sum(mean_catch, na.rm = TRUE)) %>% # summarize the mean catch for all species per rgn_id and year within the sequence of years
    ungroup() %>%
    group_by(rgn_id) %>% #for all of the years per region, find the average catch using the mean catch of each year
    summarize(avg_catch = mean(catch)) %>%
    select(rgn_id, avg_catch) %>% 
    dplyr::distinct() %>%  # ensures that there's only one row per region, removing any potential duplicates
    pivot_wider(names_from = "rgn_id", values_from = "avg_catch") # to call on specific observations more easily
}

# define parameters to input into the functions
years_to_analyze <- 2014:2017
channel_islands_ids <- c("228", "227")
bes_islands_ids <- c("245", "249", "248")

# calculate proportion of average catch for Channel Island regions for the 3 most recent years
mean_catch_ci_agg <- calculate_avg_catch(mean_catch_2017, channel_islands_ids, years_to_analyze) %>%
  rename(jersey = `227`, guernsey = `228`) %>% 
  pivot_longer(cols = c(jersey, guernsey),
             names_to = "rgn_id",
             values_to = "avg_catch") %>% 
  mutate(prop = avg_catch/sum(avg_catch, na.rm = TRUE))

# ccalculate proportion of average catch for Bonaire/S.Eustatius/Saba
mean_catch_bes_agg <- calculate_avg_catch(mean_catch_2017, bes_islands_ids, years_to_analyze) %>%
  rename(bonaire = `245`, sint_eustatius = `249`, saba = `248`) %>% 
  pivot_longer(cols = c(bonaire, sint_eustatius, saba),
             names_to = "rgn_id",
             values_to = "avg_catch") %>% 
  mutate(prop = avg_catch/sum(avg_catch, na.rm = TRUE))

# calculate proportions for each disaggregated region
## Channel Islands
prop_j <- mean_catch_ci_agg$prop[1]
prop_g <- mean_catch_ci_agg$prop[2]

## Bonaire/S. Eustatius/Saba
prop_b <- mean_catch_bes_agg$prop[1]
prop_e <- mean_catch_bes_agg$prop[2]
prop_s <- mean_catch_bes_agg$prop[3]
```

```{r}
# use proportions to disaggregate Channel Islands 
split_ci <- cf_revenue %>%
  filter(country == "Channel Islands") %>%
  mutate(split_col = "Guernsey;Jersey") %>%
  separate_rows(split_col, sep = ";") %>%
  group_by(country, year, value, unit) %>%
  mutate(new_value = case_when(
    split_col == "Guernsey" ~ value * prop_g,
    split_col == "Jersey" ~ value * prop_j,
    TRUE ~ NA_real_
  )) %>% 
  ungroup() %>%
  mutate(country = split_col) %>%
  select(-split_col)

# split for Bonaire/S.Eustatius/Saba
split_bes <- cf_revenue %>%
  filter(country == "Bonaire/S.Eustatius/Saba") %>%
  mutate(split_col = "Bonaire;Sint Eustatius;Saba") %>%
  separate_rows(split_col, sep = ";") %>%
  group_by(country, year, value, unit) %>%
  mutate(new_value = case_when(
    split_col == "Bonaire" ~ value * prop_b,
    split_col == "Sint Eustatius" ~ value * prop_e,
    split_col == "Saba" ~ value * prop_s,
    TRUE ~ NA_real_
  )) %>% 
  ungroup() %>%
  mutate(country = split_col) %>%
  select(-split_col)

# merge with cf_revenue
cf_revenue_disaggregated <- cf_revenue_fix %>%
  bind_rows(split_ci, split_bes) %>%
  mutate(value = case_when(
    country %in% c("Guernsey", "Jersey") ~ new_value,
    country %in% c("Bonaire", "Sint Eustatius", "Saba") ~ new_value,
    TRUE ~ value
  )) %>%
  select(-new_value)
```

Final names_2_rgn after disaggregation
```{r}
#using disaggregated revenue data: looks good!! 
cf_revenue_rgn_final <- name_2_rgn(df_in = cf_revenue_disaggregated, 
                             fld_name='country',
                             flds_unique = c("year"))

#========== Indexing to check fxn ===============

# to see how many duplicates there are after names_2_rgn
rev_duplicates <- cf_revenue_rgn_final[duplicated(cf_revenue_rgn_final[, c("country", "year")]),]

# save the unique duplicates for country and rgn_name to see if the function worked well
rev_ref_duplicates <- unique(rev_duplicates$country)
rev_rgn_duplicates <- unique(rev_duplicates$rgn_name)

# setdiff to determine which observations are not the same in the country and rgn_id duplicates
rev_diff <- setdiff(rev_rgn_duplicates, rev_ref_duplicates)
rev_diff # none! Move on

#=========== Aggregate by rgn_id and year ==============
cf_revenue_rgn_agg <- cf_revenue_rgn_final %>% 
  select(rgn_id, rgn_name, year, value, unit, sector, data_source) %>% 
  group_by(rgn_id, rgn_name, year) %>% # because china needs to be aggregated and has duplicate years
  dplyr::summarize(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>% 
  dplyr::mutate(year = as.numeric(year)) 
```

## CF Revenue Plot
```{r}
cf_rev_plot <- plotly::plot_ly(cf_revenue_rgn_agg, x = ~year, y = ~value, color = ~rgn_name, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Yearly Revenue for the Marine Fishing Sector", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Revenue of all catch (USD)"))
cf_rev_plot

# looks great! China has a very large revenue, which makes sense.  Indonesia and Peru are the have the next highest revenue in 2019.  Interesting to see a decrease in the early 2000s for most regions.
```

Heatmap
```{r}
# make all geographic areas have the same date range (and fill missing values with NAs) to avoid problems downstream
years_df <- tibble(rgn_name = cf_revenue_rgn_agg$rgn_name) %>% 
  mutate(rgn_id = cf_revenue_rgn_agg$rgn_id) %>% 
  group_by(rgn_id, rgn_name) %>% 
  summarize(year = seq(min(cf_revenue_rgn_agg$year),
                       max(cf_revenue_rgn_agg$year)))

# ======== join with cf revenue aggregated data ========
cf_revenue_yrs <- left_join(years_df, cf_revenue_rgn_agg,
                          by = c("rgn_id", "rgn_name", "year"))

```


## Saving intermediate data
```{r}
# columns should be: rgn_id, rgn_name, year, usd (previously value), unit, sector, usd_yr (in this case = year)

cf_revenue_int <- cf_revenue_yrs %>% 
  mutate(usd = value,
         unit = "USD (1)",
         sector = "cf",
         usd_yr = year) %>% 
  select(-value)

# write_csv(cf_revenue_int, here(int_dir, "eco_cf_usd_pre.csv"))
```


