---
title: "OHI Economies Dataprep"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------


# Economies Dataprep (Measures the amount of revenue produced)

## Overview

-   **Data:**

    -   Tourism: UNWTO Dept. of Statistics (UN Tourism)
        -   https://www.unwto.org/tourism-statistics/economic-contribution-SDG 

### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  countrycodes,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo # for gapfilling
  
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- paste0("globalprep/le/", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO raw data directory
unwto_dir <- here(raw_data_dir, "UNWTO", data_dir_version_year)
```


### Tourism

#### Read in Data

```{r}
# ===================== Read in Data ==============================


# Tourism revenue data ----
gdp_prop_raw <- readxl::read_xlsx(here(unwto_dir, "Indicator-8_9_1-2022-UN_Tourism_april2024_update.xlsx"))

# OHI regions data ----
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") 
```


## Data cleaning

"Tourism direct GDP as a proportion of total GDP (%)"


```{r}
# make countrycode -- iso3 function ----
# following this example on Rpubs: https://rpubs.com/Teal_Emery/cleaning_intl_data_tips_and_tricks 
country_regex_to_iso3c <- function(country_string) {
  country_string %>%
    countrycode::countrycode(origin = "country.name", destination = "iso3c", origin_regex = TRUE)
}

# GDP proportion dataset cleaning ----
gdp_prop_clean <- gdp_prop_raw %>% 
  janitor::clean_names() %>% 
  select(c(geo_area_name, year = time_period, gdp_prop_percent = total)) %>% 
  # add iso3 column
  mutate(iso3 = country_regex_to_iso3c(geo_area_name))


# preliminary plotting ----
prelim_plot <- plot_ly(data = gdp_prop_clean, x = ~year, y = ~gdp_prop_percent, color = ~geo_area_name, 
                       type = "scatter", mode = "lines") %>% 
  layout(title = "Tourism direct GDP as a proportion of total GDP (%) for All Geo Regions",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Tourism Contribution to GDP (%)"))

prelim_plot

htmlwidgets::saveWidget(prelim_plot, file = "gdp_prelim_plot_all_years.html")
```

Really interesting to see how the proportions drop in 202 -- intuition check: makes sense because of Covid.... especially interesting to note how extremely Macau was impacted!


## Gapfilling

More data cleaning & prep for gapfilling, then gapfilling

```{r}

# filter tourism GDP data for year range (2009-2019) ----
gdp_prop_filter <- gdp_prop_clean %>% 
  filter(year %in% c(2009:2019))


# some geo areas are missing data for some years ----
years_df <- tibble(iso3 = gdp_prop_filter$iso3) %>% # copy iso3 column
  group_by(iso3) %>% # for each geo region, 
  summarize(year = seq(2009, 2019)) # have sequence of years from 2009-2019 
# (fragile and R warns you about preferring `reframe()`)


# join years df and tourism GDP df ----
gdp_year_join <- left_join(years_df, gdp_prop_filter, by = c("iso3", "year"))


# fill in geo area names for NA years ----
gdp_years <- gdp_year_join %>% 
  group_by(iso3) %>% 
  fill(geo_area_name, .direction = "downup") %>% 
  ungroup()


# gapfilling ----
# estimate tourism GDP proportion for geo areas with missing data
gdp_filled <- gdp_years %>% 
  group_by(geo_area_name) %>% 
  # interpolate (fill missing values between 2 values)
  mutate(appx_gdp = zoo::na.approx(gdp_prop_percent, # using values in this column
                                    na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                    #  extrapolate using rule = 2 from approx(),
                                    # uses closest data extreme to extrapolate 
                                    # for leading and trailing NAs
                                    rule = 2)) %>% 
  
  # fill NA values with the only other observed value (making sure to note the list of countries/regions we did this for)
  mutate(appx_gdp_fill = zoo::na.aggregate(appx_gdp, by = 1, FUN = mean, na.rm = FALSE))
```


Assess how many countries we are gapfilling more extremely for

```{r}
# find number of regions that only have 1 data point (wouldn't be filled by na.approx if only 1 datapoint, left as NAs)
na_regions <- gdp_filled %>% 
  filter(is.na(appx_gdp))

# print which geo areas only had 1 datapoint
unique(na_regions$geo_area_name)

# calculate proportion of 1 point-filled geo areas to total geo areas
num_na <- length(unique(na_regions$geo_area_name))
num_tot <- length(unique(gdp_filled$geo_area_name))

paste0("Proportion of geo areas with only 1 data point: ", round(((num_na / num_tot) * 100), 3), "%")
```

#### Preliminary Plotting

```{r}
prelim_gf_plot <- plot_ly(data = gdp_filled, x = ~year, y = ~appx_gdp_fill, color = ~geo_area_name,
                          type = "scatter", mode = "lines") 


top_tourism_countries <- 

prelim_gf_plot
```



### Join with OHI regions


```{r}

# tidy up regions dataset ----
region_names <- region_names %>% select(-Notes)


# join regions & tourism GDP prop ----
gdp_regions_join <- left_join(region_names, gdp_filled, by = c("eez_iso3" = "iso3"))





```


Note: remember to add
  mutate(sector = "tour", data_source = "UNWTO Dept. of Statistics")
  
  at the end