---
title: "OHI `r format(Sys.Date(), '%Y')` - CF Livelihoods Job Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# CF Livelihoods: Proportion of Tourism Jobs per Country/Region per Year

## Overview

* **Data:** Labor Force & Employment Data

  * Labor Force data from World Bank (downloaded June 28. 2024)

      - https://data.worldbank.org/indicator/SL.TLF.TOTL.IN
      
  * OECD (Employment in fisheries, aquaculture and processing, 2009 - 2021) (downloaded July 2, 2024) -- for cf job data
  
    - <https://data-explorer.oecd.org/vis?fs[0]=Topic%2C1%7CAgriculture%20and%20fisheries%23AGR%23%7CFisheries%20and%20aquaculture%23AGR_FSA%23&pg=0&fc=Topic&bp=true&snb=6&vw=ov&df[ds]=dsDisseminateFinalDMZ&df[id]=DSD_FISH_EMP%40DF_FISH_EMPL&df[ag]=OECD.TAD.ARP&df[vs]=1.0&dq=.A...._T._T&pd=2009%2C&to[TIME_PERIOD]=false&ly[cl]=TIME_PERIOD&ly[rs]=REF_AREA&ly[rw]=DOMAIN>
    
  * FAO Yearbook (downloaded July 2, 2024) -- for cf job gapfilling if needed
  
    - <https://openknowledge.fao.org/server/api/core/bitstreams/2be6c2fa-07b1-429d-91c5-80d3d1af46a6/content>
      
  * ILOSTAT (downloaded July 2, 2024) -- for cf wage data
  
      - https://rshiny.ilo.org/dataexplorer46/?lang=en&id=EAR_4MTH_SEX_ECO_CUR_NB_A 
      - select Rev 3.1.B: Fishing
      
  * Ex-Vessel Price Data (downloaded August 24, 2023) -- for cf wage data
      - from emlab, see README.md


### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo, # for gapfilling
  pdftools
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

source(here('workflow/R/fao_fxn.R')) # function for cleaning old version of FAO data 

source(here('workflow/R/fao_online_portal_clean.R'))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# ILO raw data directory
ilo_dir <- here(raw_data_dir, "ILOSTAT", data_dir_version_year)

# FAO raw data directory
fao_dir <- here(raw_data_dir, "FAO", data_dir_version_year)

# OECD raw data directory
oecd_dir <- here(raw_data_dir, "OECD", data_dir_version_year)

# FAO capture raw data directory
fao_capture_dir <- here(raw_data_dir, "FAO_capture", data_dir_version_year)

# Ex-vessel raw data directory
ex_vess_dir <- here(raw_data_dir, "ex-vessel-price-database-updated")

# output data dir for intermediate data products
int_dir <- here(data_path, "int")
```


Create country code --> iso3 function
```{r}
library(countrycode)

country_regex_to_iso3c <- function(country_string) {
  country_string %>%
    countrycode::countrycode(origin = "country.name", destination = "iso3c", origin_regex = TRUE)
}
```

## Read in Data

```{r}
#----------OHI regions data--------------------
# read in OHI regions for joining
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") %>% 
  janitor::clean_names() %>% 
  select(-notes)

#----------Labor Force Data (from earlier)-----
labor_force <- read_csv(here(data_path, "int", "labor_force.csv"))

############################################
# Livelihoods: Number of fisher jobs
############################################

#---------- cf OECD data---------------------------
oecd_cf_job_raw <- read_csv(here(oecd_dir, "OECD_2009-2021_employment_fish_aqua_processing_raw.csv"))

#------cf FAO Yearbook Number of Fishers-------
# turn the number of fishers table from the FAO yearbook into a data frame and clean.
# note: world total includes countries and regions not within the table.  Unsure of where that data comes from; look for methodology.
fao_pdf <- pdf_text(here(fao_dir, "fao_statistics_yearbook_07_03_24.pdf"))

fishers_table <- fao_pdf[[169]] %>% #page number of the pdf
  as.data.frame()

#specify columns 
columns <- c("country", "1995", "2000", "2005", "2010", "2015", "2016", "2017", "2018", "2019", "2020","2021")

#turn the pdf page into a clean dataframe
fao_fisher_jobs <- fishers_table %>%
  separate_rows('.', sep = "\n") %>% #separate rows based on line breaks
  rename(main_column = ".") %>%
  mutate(main_column = str_remove_all(main_column, "\\*")) %>% #remove the * 
  mutate(main_column = str_replace_all(main_column, "E(?!\\p{L})", " ")) %>% #replace all the Es not followed by letters (used to indicate estimate)
  slice(7:69) %>% #remove all rows not in the table
  separate(main_column, into = columns, sep = " {2,}", extra = "merge") %>% #separate into columns based on double spaces
  pivot_longer(cols = -c("country"), names_to = "year") %>% 
  mutate(value = str_squish(value)) %>% #remove extra spaces
  mutate(value = str_remove_all(value, " ")) %>% 
  mutate(value = if_else(value == "…" , NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  mutate(unit = "persons (1)",
         sector = "cf",
         data_source = "FAO number of fishers") %>% 
  slice(-c(1:44))
```


