---
title: "OHI `r format(Sys.Date(), '%Y')` - Livelilihoods Data Preparation"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Livelihoods: Proportion of Tourism Jobs per Country/Region per Year

## Overview

* **Data:** Labor Force & Employment Data

  * Labor Force data from World Bank (downloaded June 28. 2024)

- https://data.worldbank.org/indicator/SL.TLF.TOTL.IN


### Setup

```{r}
# load packages
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  here,
  janitor,
  terra,
  readxl,
  RColorBrewer,
  foreach,
  doParallel, # for using multiple cores
  tidyverse, 
  httr,
  plotly,
  zoo # for gapfilling
  
)
#remotes::install_github("skgrange/threadr") # for na_extrapolate
library(threadr)


# source 
source(here("workflow/R/common.R"))

# set year and file path info
current_year <- 2024 # Update this!!

version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
data_path <- here("globalprep", "le", version_year)

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# world bank raw data directory
wb_dir <- here(raw_data_dir, "WorldBank", data_dir_version_year)

# output data dir for intermediate data products
int_dir <- here(data_path, "int")
```

### Read in Data

```{r}
# ===================== Read in Data ==============================


# Labor force data ----
labor_raw <- readxl::read_xls(here(wb_dir, "worldbank_labor_force_raw.xls"),
                               skip = 3,
                               na = "")

# OHI regions data ----
# read in OHI regions for joining
region_names <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/eez/spatial/regions_list.csv") 


# ILO wage data ----
# define file path to ILO wage data 
ilo_fp <- file.path(dir_M,
                        "git-annex/globalprep/_raw_data/ILOSTAT/d2024/EAR_4MTH_SEX_ECO_CUR_NB_A-filtered-2024-07-02-tour-wage.csv")

# read in ILO wage data
ilo_wage_data <- read_csv(ilo_fp)


# ==================== Tidy Data ===================================

# Labor force data -----------------------------------
labor_clean <- labor_raw %>% 
  # conver to lower_snake_case
  janitor::clean_names() %>% 
  # tidy data from wide to long (years are currently columns)
  pivot_longer(cols = 5:length(names(labor_raw)),
               names_to = "year", 
               values_to = "thousand_jobs") %>% 
  # remove unnecessary columns
  select(-c(indicator_name, indicator_code)) %>% 
  # clean up year column (currently in the form of xYYYY)
  mutate(year = str_remove_all(year, pattern = "x")) %>% 
  # fix data types
  mutate(year = as.numeric(year),
         country_name = as.factor(country_name),
         thousand_jobs = as.numeric(thousand_jobs))
  

# clean regions data --------------
region_clean <- region_names %>% 
  janitor::clean_names() %>% 
  # drop notes column
  select(-c(("notes")))



# ============== Join labor force + OHI regions =============================

labor_regions_join <- left_join(region_clean, labor_clean,
                                by = c("eez_iso3" = "country_code"))
# note: many-to-many warning is not a cause for concern here because for countries in the labor dataset that match admin countries (but not region names) in the region dataset, it'll populate the rows of region-admin country pairs where there is only a match between admin and labor country names with that country's admin-country-level labor data. It will be important to mention this in the data gaps or analysis, however.


labor_regions <- labor_regions_join %>% 
  select(-c(country_name)) %>% 
  filter(year >= 2009) # starting 5 years back from first year of evaluation

```

# Employment Data

(put this within previous chunk, read in data etc. for number of tourism jobs, join with labor_regions, make new proportion of tourism jobs column, group by year, do prelim plotting etc. and evaluate data quality etc.)

```{r}
# total number of people employed by the tourism sector, in thousands (remember this!)
un_job_data <- file.path(dir_M, "git-annex/globalprep/_raw_data/UNWTO/d2024/unwto-all-data-download_0.xlsx")

# take in .xlsx and read it in, clean it
un_jobs_data_raw <- read_xlsx(un_job_data, skip = 2, sheet = "Employment", col_names = TRUE, na = c("","..")) %>%
  janitor::clean_names() %>%
  select(-c(1:3)) %>%
  fill(1, .direction = "down") %>%
  select(-c(x5, x7:units, notes, x38)) %>%
  filter(x6 %in% "Total") %>%
  select(-x6)

# pivot longer so that we can see how many jobs per year, per country!
un_jobs_data_piv <- un_jobs_data_raw %>%
  pivot_longer(cols = 2:28, names_to = "year", values_to = "jobs") %>%
  mutate(year = str_remove_all(year, 'x')) %>%
  rename(country = basic_data_and_indicators) %>%
  mutate()

# install.packages("countrycode")
library(countrycode)

country_regex_to_iso3c <- function(country_string) {
  country_string %>%
    countrycode::countrycode(origin = "country.name", destination = "iso3c", origin_regex = TRUE)
}

# adding iso3 codes, converting number to thousands to be consistent and prep for the left_join with labor_regions
un_jobs_cc <- un_jobs_data_piv %>% 
  mutate(iso3c = country_regex_to_iso3c(country)) %>%
  mutate(tourism_jobs = jobs*1000) %>%
  mutate(country = str_to_title(country)) %>%
  select(-jobs) %>%
  mutate(year = as.numeric(year))
# warning: ! Some values were not matched unambiguously: BONAIRE, SABA, SERBIA AND MONTENEGRO, SINT EUSTATIUS

# join w OHI regions
tourism_un_jobs_ohi <- left_join(region_clean, un_jobs_cc, by = c("eez_iso3" = "iso3c"))


# ----------Join with Labor Force Data--------------
tourism_labor_join <- left_join(labor_regions, tourism_un_jobs_ohi, by = c("rgn_id", "year")) 
```

Create the proportion between tourism jobs and the total labor force
```{r}
tourism_job_proportion <- tourism_labor_join %>%
  dplyr::relocate(tourism_jobs, .before = thousand_jobs) %>%
  mutate(tourism_proportion = tourism_jobs/thousand_jobs) %>%
  rename(labor_force_total = thousand_jobs) %>%
  select(rgn_id, rgn_name.x, country, year, labor_force_total, tourism_jobs, tourism_proportion)

# plot it, interactively!
line_plot <- plotly::plot_ly(tourism_job_proportion, x = ~year, y = ~tourism_proportion, color = ~country, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Proportional Tourism Employment Within Total Labor Force", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Percent of people by region employed by the tourism sector"))

line_plot

# htmlwidgets::saveWidget(line_plot, file = "prop_tourism_laborforce.html")
```

### Saving intermediate data

```{r}
#write_csv(tourism_job_proportion, here(int_dir, "tour_jobs.csv"))
```


## ILO wage data
  - note: this data came from many sources and was compiled by ILO

```{r}
# preliminary cleanings
wage_data_clean <- ilo_wage_data %>% 
  # lower_snake_case
  janitor::clean_names() %>% 
  # more intuitive value name
  rename(monthly_wage = obs_value) %>% 
  # filter to PPP adjusted data (this accounts for inflation, diff in currencies globally, cost of living etc.)
  filter(classif2_label == "Currency: 2017 PPP $") %>% 
  # group by country/region
  group_by(ref_area_label) %>% 
  # set column name to year
  rename(year = time) %>% 
  # filter to the cap date range from the jobs data -- 2019
  filter(year %in% c(2009:2019)) %>%  
  # add iso3 column, tourism sector label
  mutate(iso3 = country_regex_to_iso3c(ref_area_label),
         sector = "tour") %>% 
  select(c(ref_area_label, iso3, year, monthly_wage, classif2_label))
  


# make sequence of years for gapfilling ----
# note: data does not have wage data for some years in some countries -- we want to account for that by leaving them as NAs, then gapfilling with the average between the pre and post years

years_df <- tibble(iso3 = wage_data_clean$iso3) %>% 
  group_by(iso3) %>% 
  summarize(year = seq(2009, 2019))


wage_data_years <- left_join(years_df, wage_data_clean, by = c("iso3", "year"))

# filling in country names when possible --- 
wage_years_filled <- wage_data_years %>% 
  group_by(iso3) %>% 
  fill(ref_area_label, .direction = "downup") %>% 
  ungroup()

# =================

# test gapfilling
# gap_fill_test <- wage_years_filled %>% 
#   mutate(ref_area_label = as.factor(ref_area_label)) %>% 
#   mutate(lm_est = list(lm(monthly_wage ~ year + ref_area_label)))
# 
# 
# lm_test <- lm(monthly_wage ~ year + ref_area_label, data = wage_years_filled)
# summary(lm_test)
# 
# lm_test$coefficients 

# wage_years_filled$lm_values <- lm_test$fitted.values

# gap filling  ----
wage_gf <- wage_years_filled %>% 
  group_by(ref_area_label) %>% 
  # interpolate (fill missing values between 2 values)
  mutate(appx_wage = zoo::na.approx(monthly_wage, # using values in this column
                                    na.rm = FALSE, # don't replace (internal) NAs in new column that can't be approximated
                                    #  extrapolate using rule = 2 from approx(),
                                    # uses closest data extreme to extrapolate 
                                    # for leading and trailing NAs
                                    rule = 2)) %>% 
  
  # fill NA values with the only other observed value (making sure to note the list of countries/regions we did this for)
  mutate(appx_wage_fill = zoo::na.aggregate(appx_wage, by = 1, FUN = mean, na.rm = FALSE))
# note: made this a new column so that we can still identify each country that we are using this fill (na.aggregate function) method for.
  
# other notes: 
# need more than 1 data point to do any approximation. 
# also, extremes (leading and trailing) are just copied data points from nearest extreme

```



```{r}
# find regions with only 1 data point (still have NAs in appx_wage column)

na_regions <- wage_gf %>% 
  filter(is.na(appx_wage))


unique(na_regions$ref_area_label)
num_na <- length(unique(na_regions$ref_area_label))
num_tot <- length(unique(wage_gf$ref_area_label))

paste0("proportion of countries/regions with only 1 data point: ", round(((num_na / num_tot) * 100), 3), "%")
```




```{r}
# preliminary plot 
line_plot <- plotly::plot_ly(wage_gf, x = ~year, y = ~appx_wage_fill, color = ~ref_area_label, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: monthly Service Wages (USD (2017) PPP-Adjusted)", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "monthly Wages (USD)"))

line_plot

# htmlwidgets::saveWidget(line_plot, file = "prop_tourism_laborforce.html")

```




In this plot we can see that Sao Tome and Principe is a clear high outlier with no variation -- it is one of the countries for which we gapfilled with copied values from a single observation. We will continue to join the data with the OHI regions, then plot again and reinvestigate.


Clean up ILO data, join with OHI regions

```{r}
# join with OHI regions
wage_region_join <- left_join(region_clean, wage_gf, by = c("eez_iso3" = "iso3"))

# come back ---- stopping point for Mel to take a look!

# Currency: 2017 PPP $

wage_regions <- wage_region_join %>% 
  mutate(unit = "Currency: 2017 PPP $") %>% 
  select(-classif2_label) %>% 
  mutate(sector = "tour",
         data_source = "ILO")

```


Plot

```{r}
# interactive plot after joining with OHI Regions
plotly::plot_ly(wage_regions, x = ~year, y = ~appx_wage_fill, color = ~admin_country_name, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Monthly Service Wages (USD)", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Monthly Wages (USD)"))

```

We can see that Sao Tome and Principe is still included and is still this high, unvarying outlier. 
This lack of data combined with how high of an outlier is led us to reconsider our methods for this country and consider dropping it from this intermediate data product. 

Let's drop & plot what that looks like. 

```{r}
wage_regions_no_sp <- wage_regions %>% 
  filter(!ref_area_label %in% c("Sao Tome and Principe"))

# plot to see what it looks like after dropping Sao Tome and Principe
plotly::plot_ly(wage_regions_no_sp, x = ~year, y = ~monthly_wage, color = ~admin_country_name, type = "scatter", mode = "lines") %>% 
  layout(title = "All Regions: Monthly Service Wages (USD)", 
         xaxis = list(title = "Year"),
         yaxis = list(title = "Monthly Wages (USD)"))

```

Dropping Sao Tome and Principe (STP) adjusted the y-axis scale significantly to better visualize the monthly service wage data (USD) for other geographic areas.  Investigating further, it looks as though the Sao Tome and Principe observation from 2017 (133814.9 USD) may have been inputted incorrectly, measured as annual income instead of monthly, or there was an error when it was adjusted by PPP.  Dustin noticed that the data source from this observation was "HIES - Household Budget Survey".  This was the only observation from this data source. Upon further investigation, we discovered the average annual income in STP is ~2,400 USD [WorldData]("https://www.worlddata.info/africa/sao-tome-and-principe/index.php#:~:text=With%20an%20average%20annual%20income,the%20lower%20middle%2Dincome%20countries") and a GDP per capita of ~2,817 USD [World Bank]("https://www.worldbank.org/en/country/saotome/overview"). Therefore, we have decided to drop it for the intermediate data set.


## Saving intermediate data products

```{r}
# write_csv(wage_regions_no_sp, here(int_dir, "tour_quality.csv"))

```

