---
title: 'OHI `r format(Sys.Date(), "%Y")` - Tourism and Recreation '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html' 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

[REFERENCE RMD FILE](http://ohi-science.org/ohiprep_v2024/globalprep/tr/v2024/tr_data_prep.html)

# Summary

This document describes the steps for obtaining the data used to calculate the tourism and recreation goal for the 2024 global assessment.

The general calculation is: tr = Ap \* Sr and Xtr = tr/90th quantile across regions

-   Ap = Proportion of overnight tourist arrivals to total arrivals
-   Sr = (S-1)/5; Sustainability of tourism

*The following data are used:*

-   Numbers of tourist arrivals, used in the calculation of the proportion of arrivals to area of coastline to population: obtained through the [UNWTO](https://www.unwto.org/tourism-statistics/key-tourism-statistics) (in the form of thousands of people). More info on tourism terms [here](https://www.unwto.org/glossary-tourism-terms). Range: 1995-2021 (only 2008-2021 is used)
-   Tourism sustainability: World Economic Forum. The Travel & Tourism Development Index 2021 dataset (version 24 May 2022). 2022. [TTDI](https://www.weforum.org/reports/travel-and-tourism-development-index-2021/downloads-510eb47e12#report-nav)
-   Per capita GDP: (World Bank with gaps filled using CIA data), used to gapfill missing values in Tourism sustainability

# Updates from previous assessment

**We were able to update the following data:**

-   Proportion of tourist arrivals to area of coastline to population - UNWTO data on thousands of tourist arrivals, reported until 2021 (downloaded [here](https://www.unwto.org/tourism-statistics/key-tourism-statistics) (dataset: "Total arrivals" under Inbound Tourism; we use the "Overnights visitors (tourists)" categorization as arrivals where possible) on 8/10/2023.


# v2024 updates

Building upon updates made in v2023 for Tourism & Recreation.

### Layers

- tr_arrivals_props_tourism (arrivals to a country, staying overnight)
- tr_sustainability (Tourism sustainability index)

### Data Sources
Using Erika and Gage's methodology (v2023):

- UNWTO data on thousands of tourist arrivals, reported until 2021 (downloaded here (dataset: "Total arrivals" under Inbound Tourism; we use the "Overnights visitors (tourists)" categorization as arrivals where possible) on 8/10/2023. Use Overnights from this, and gapfill Overnights with Total minus Same-day if possible.
- Sum of inland (1 km) and offshore (3 nm) area (calculated from the csvs used in LSP)
- Population data
  - primarily from the WDI() function which retrieves data from World Bank and gapfill missing populations that also have arrivals data (I do not gapfill all missing populations, just ones that match with regions that don't have NAs in the arrivals data)
  - secondarily with Our World in Data data; in this data, Saba, Sint Eustatius, and Bonaire are not broken up, so I found broken up values separately for them from Statista. These only go back to 2011 instead of the earliest data year for this layer which is 2008, so I gapfill 2008-2010 with each of their respective 2011 data. This causes all OHI regions with arrivals data to now have both inland + offshore data and population data.
- v2024: The (tourism sustainability index dataset)[https://www3.weforum.org/docs/WEF_Travel_and_Tourism_Development_Index_2024.pdf] (TTDI) has been updated! 

### Next Steps

- Explore updated UNWTO data. Determine if Inbound Tourism Arrivals Overnight stays can be aggregated with Domestic Arrivals Overnight stays. - See where there are gaps, what would need to be filled.
- If there are too many gaps, look at employment data instead.
- Integrate the fixed (globalprep/mar_prs_population/v2024/output/mar_pop_25mi.csv) coastal population data for a new method of calculating "Ap" (Arrival proportion):

> Ap = multiply arrivals by coastal population/total population, then divide by coastal area (sum of inland (1 km) and offshore (3 nm) area)

- Rescale using the 90th quartile as the reference point, or 80th (reevaluate when we get to this step)
- Consider dropping low population areas, like they did in v2023
- Calculate `Sr = (S-1)/5`; Sustainability of tourism
- Finally, use the equation: `tr = Ap * Sr` for final scores before updating ohi-global

## Initial set-up code

```{r setup, message=FALSE, warning=FALSE, results="hide"}
# library(devtools)
# devtools::install_github("ohi-science/ohicore@dev") # dont worry about devtools
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI)
library(here)
library(janitor)
library(plotly)
library(readxl)
library(naniar)
library(countrycode)
library(pdftools)

# ---- sources! ----
source(here("workflow", "R", "common.R")) # file creates objects to process data

region_data() # for rgns_all and rgns_eez

regions_shape() # returns spatial shape object named 'regions' which includes land, eez, highseas, and antarctica regions

#source(here(paste0("globalprep/tr/v", version_year, "/R/tr_fxns.R"))) # not used presently

# ---- set year and file path info ----
current_year <- 2024 # Update this in the future!!
version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
prev_ver_yr <- paste0("d", (current_year - 1))

# ---- data directories ----

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO (UN World Tourism) raw data directory
unwto_dir <- here(raw_data_dir, "UNWTO", data_dir_version_year)

# intermediate data outputs dir
int_dir <- here("globalprep","tr", version_year, "intermediate")

# final output dir
output_dir <- here("globalprep","tr", version_year, "output")
```

# Ap: Proportion of tourist arrivals, multiplied by coastal/total population, to area of coastline to population

We use international arrivals data from the [United Nations World Tourism Organization (UNWTO)](https://www.unwto.org/). Up through v2022, we accessed data from the [World Travel & Tourism Council (WTTC)](http://www.wttc.org/), but this was no longer a viable option in v2023. In v2024, we added Domestic Tourism Arrivals.  International and Domestic were downloaded [here](<https://www.unwto.org/tourism-statistics/key-tourism-statistics>) August 16th and 19th, 2024, respectively.

*International data:* Click Inbound Arrivals, scroll down to "Total arrivals", click download data.

*Domestic data:* Click Domestic Tourism, scroll down to "Total trips", click download.  Then download from "Accommodations".

**Gapfilling:**

To address missing values in international arrivals, specifically referring to "Overnight visitors (tourists)," we employ a two-step process. First, we attempt to fill the gaps by subtracting "Same-day visitors (excursionists)" from "Total arrivals" if the latter is available. If this is not feasible, we resort to interpolating or extrapolating based on historical data spanning from 1995 to 2019, employing a linear model to estimate increases or decreases on a regional level.

However, in light of the Covid-19 pandemic, we have adopted a distinct approach for the years 2020 and 2021. We calculate the global average proportionate change from the preceding year, apply this percentage change to the previous year's arrivals or total values, and then add the result to the corresponding previous year's arrivals or total value. - So 2020 was gapfilled with the global average decrease proportion of \~-0.70 for both arrivals and totals. Meaning that we took the 2019 value and multiplied it by -0.7 and then addd that to the 2019 value. (2019 x -0.7) + 2019 = 2020 - 2021 was gapfilled with the global average increase proportion of \~0.2. Meaning that we took the 2020 value and multiplied it by 0.2 and then addd that to the 2020 value. (2020 x 0.2) + 2020 = 2021

## International UNWTO Inbound Tourism Arrivals

```{r}
file_path_unwto_international <- here::here(unwto_dir, "unwto-inbound-arrivals-data.xlsx")
unwto_arrivals_int <- readxl::read_xlsx(file_path_unwto_international, skip = 4) # read in the raw data

unwto_clean <- unwto_arrivals_int %>% 
  select(country = `Basic data and indicators`, total_arrivals = `...6`, subdivision_1 = `...7`, subdivision_2 = `...8`, `1995`:`2021`) %>% # select relevant columns
  fill(country, .direction = "down") %>% # add country name to all data associated with that country
  pivot_longer(cols = c("total_arrivals", "subdivision_1", "subdivision_2"),
               values_to = "metric",
               values_drop_na = TRUE) %>% # make the metrics into one column
  select(-name) %>% # get rid of the name column since it's just the titles of the metrics which are already there
  select(country, metric, everything()) %>% # reorder things
  replace_with_na_all(condition = ~.x == "..") %>% # swap .. with NAs
  pivot_longer(cols = 3:ncol(.), names_to = "year",
               values_to = "tourism_arrivals_ct") %>% # make the years not columns anymore
  pivot_wider(names_from = metric, values_from = tourism_arrivals_ct) %>%
  mutate(overnights = as.numeric(`Overnights visitors (tourists)`), 
         same_day = as.numeric(`Same-day visitors (excursionists)`), 
         total_arrivals = as.numeric(`Total arrivals`),
         tourism_arrivals_ct = as.numeric(NA)) %>% # rename metrics so easier to work with, make numeric, and add a new column to fill with the new calculated values later
  select(country, year, overnights, same_day, total_arrivals, tourism_arrivals_ct) %>% # select columns needed for analysis (cruise passengers seem to be included in same-day)
  group_by(country, year) %>% # group by county and year
  mutate(
    tourism_arrivals_ct = case_when(
      !is.na(overnights) ~ overnights, # if there is a value, dont gapfill
      is.na(overnights) & !is.na(same_day) & !is.na(total_arrivals) ~ total_arrivals - same_day, # gapfill, when there is no data on overnights, fill with total_arrivals - same day
      TRUE ~ tourism_arrivals_ct # otherwise, NA
    ), # there were 0 situations like this in v2024
    total_arrivals = case_when(
      !is.na(total_arrivals) ~ total_arrivals, 
      is.na(total_arrivals) & !is.na(same_day) & !is.na(overnights) ~ overnights + same_day,
      TRUE ~ total_arrivals
    )
  ) %>% # v2024: overnights has 1036 NAs out of 6021
  # v2024: same_day has 3131 NAs out of 6021
  # v2024: total_arrivals has 2363 NAs
  mutate(arrivals_method = ifelse(is.na(overnights) & !is.na(same_day) & !is.na(total_arrivals), "UNWTO - subtraction", NA)) %>%
  mutate(arrivals_gapfilled = ifelse(arrivals_method == "UNWTO - subtraction", "gapfilled", NA)) %>% # prepare a "gapfilled" column to indicate "gapfilled" or NA
  ungroup() %>% # ungroup since not needed anymore
  select(country, year, tourism_arrivals_ct, total_arrivals, arrivals_method, arrivals_gapfilled) %>% # select only needed columns
  mutate(country = str_to_title(country), # make countries look nice
         tourism_arrivals_ct = round(as.numeric(tourism_arrivals_ct) * 1000),
         total_arrivals = round(as.numeric(total_arrivals)*1000)) # since the units were in thousands

# Run name_2_rgn() to ensure the UNWTO data has OHI region names
unwto_clean_names_int <- name_2_rgn(df_in = unwto_clean, # do this just for Bonaire since it is the only region not matching above
                                        fld_name = 'country',
                                        # flds_unique = c('year'),
                                        keep_fld_name = TRUE) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_ct, arrivals_method, arrivals_gapfilled) # v2024: not losing the USA!! Yay (compare to v2023).  Also, total_arrivals was dropped at this point because tourism_arrivals_ct has already been gapfilled using totals if same-day was present (Overnights = total - same-day), and it feels like if total was used to gapfill places that don't have any same-day data, it may be overinflating the arrivals for that region.  This can be revisited.

# ---- determine if there were any duplicates due to name_2_rgn ----

colSums(is.na(unwto_clean_names_int)) # v2024: 789 NAs in tourism_arrivals_ct
length(unwto_clean_names_int$tourism_arrivals_ct) # within 4833 observations for `tourism_arrivals_ct`

# this shows rows that have at least one duplicate, including both the first and subsequent occurrences of each duplicated row. using the tidyverse allows us to see all instances of duplicated data, not just the second and subsequent occurrences (which is what just a duplicated() call would give you).
duplicates_int <- unwto_clean_names_int %>%
  group_by(rgn_id, rgn_name, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates_int$rgn_name)
# v2024:
# [1] "China"                                              
# [2] "Guadeloupe and Martinique"                          
# [3] "Northern Mariana Islands and Guam"                  
# [4] "Montenegro"                                         
# [5] "Puerto Rico and Virgin Islands of the United States"

## v2024: this code to fix duplicates was not working, and actually removed the duplicates from above from the data
# unwto_dupe_fix <- unwto_clean_names_int %>%
#   group_by(rgn_id, year, arrivals_method, arrivals_gapfilled) %>%
#   summarize(sum_fix = ifelse(all(is.na(tourism_arrivals_ct)), NA, sum(tourism_arrivals_ct, na.rm = TRUE)),
#             sum_fix_2 = ifelse(all(is.na(total_arrivals)), NA, sum(total_arrivals, na.rm = TRUE))) %>%
#   mutate(arrivals_method = ifelse(is.na(arrivals_method) & !is.na(sum_fix), "UNWTO", arrivals_method)) %>%
#   rename(tourism_arrivals_ct = sum_fix,
#          total_arrivals = sum_fix_2)

# instead, make a function to avoid losing NAs (they turn to 0s) when there are values to aggregate with
sum_with_na <- function(x) {
  if (all(is.na(x))) { # for x, if all of the values are NA, leave as NA.  
    return(NA)
  } else { 
    return(sum(x, na.rm = TRUE)) # if there are values for that grouped dataframe, then sum them according to the grouping (in this case, by region and year)
  }
}

# use the function within a summarize to aggregate for all regions that have duplicate years
international_clean_names_agg <- unwto_clean_names_int %>% 
  group_by(rgn_id, rgn_name, year) %>% 
  summarize(
    tourism_arrivals_ct = sum_with_na(tourism_arrivals_ct), # put each column in the function, so no NAs are lost
    data_source = first(arrivals_method, na_rm = TRUE),
    arrivals_gf = first(arrivals_gapfilled, na_rm = TRUE),
    .groups = "drop" # to ungroup, fixes warning from summarise to use reframe (when it is not applicable here because we want to reduce each group down to a single row)
  ) %>% 
  mutate(
    data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "UNWTO", data_source) # so that if there is data from UNWTO that was not gapfilled, the data_source is accurate
  )

# double check to test
duplicates_int_2 <- international_clean_names_agg %>%
  group_by(rgn_id, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates_int_2$rgn_name) # v2024: 0! yay -- and China/other duplicates are preserved here!!


# check out things so far
summary(international_clean_names_agg) 
# v2023: 828 NAs in arrivals (before filtering the years down and gapfilling), 1708 in `total_arrivals`
# v2024: 757 NAs in arrivals (before filtering the years down and gapfilling), `total_arrivals` removed already

# see how many regions have values
check_international <- international_clean_names_agg %>% 
  filter(!is.na(tourism_arrivals_ct))
length(unique(check_international$rgn_id)) # v2024: 161 regions present!
```

### Gapfilling International Arrivals

```{r}
# gapfill for international arrivals!
# downfill then upfill missing values using a linear model of the average increase per years across all years of data for 1995-2019
# for 2020 and 2021, use the global average proportion increase or decrease and add to the previous years value

# reminder: 
# `tourism_arrivals_ct` is the tourism arrivals count from the INTERNATIONAL data, which consists of overnights, and was gapfilled with (total_arrivals - same_day).

test_int_2021 <- international_clean_names_agg %>%
  filter(year %in% c(2020, 2021)) %>%
  filter(!is.na(tourism_arrivals_ct)) %>%
  pivot_wider(names_from = year, 
              values_from = tourism_arrivals_ct,
              names_prefix = "int_tourism_") %>%
  mutate(tourism_ct_diff = (int_tourism_2021 - int_tourism_2020)/int_tourism_2020)

gf_2021_tourism_int <- mean(test_int_2021$tourism_ct_diff, na.rm = TRUE) # global average increase for 2021 tourism_arrivals_ct
# v2024: 0.1290837

# ok, lets use these values to gapfill 2021 if it is NA and 2020 exists. So increase by X proportion (0.129)

test_int_2020 <- international_clean_names_agg %>%
  filter(year %in% c(2019, 2020)) %>%
  filter(!is.na(tourism_arrivals_ct)) %>%
  pivot_wider(names_from = year, 
              values_from = tourism_arrivals_ct,
              names_prefix = "int_tourism_") %>%
  mutate(tourism_ct_diff = (int_tourism_2020 - int_tourism_2019)/int_tourism_2019)
 
gf_2020_tourism_int <- mean(test_int_2020$tourism_ct_diff, na.rm = TRUE) # global average decrease for 2020 toursism
# v2024: -0.7118338

# ok, lets use these values to gapfill 2020 if it is NA and 2019 exists. So decrease by X proportion ~70%

# upfill earlier years that do not have data with nearby year data
unwto_int_upfill <-  international_clean_names_agg %>%
  filter(year < 2020) %>%
  group_by(rgn_id) %>%
  arrange(rgn_id, year) %>%
  tidyr::fill(tourism_arrivals_ct, .direction = "up") %>% # fill in any values that are empty from early years with values from the nearest year. Doing this because doesn't make sense to add earlier years based on a trend 
  mutate(
    arrivals_gf = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "gapfilled", arrivals_gf),
    data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "UNWTO - nearby year", data_source)
         )

## calculate regional average increase or decrease in number of tourism arrivals
lm_coef_data_tourism_int <- international_clean_names_agg %>%
  filter(!(year %in% c(2020, 2021))) %>%
  group_by(rgn_id) %>%
  filter(!is.na(tourism_arrivals_ct)) %>%
  summarize(lm_coef_tourism = lm(tourism_arrivals_ct ~ year)$coefficients[2])

# Initialize a flag to check if there are still NAs
na_flag <- TRUE

# filter out any regions with all nas for each year, as these can't be gapfilled
all_nas_tourism_int <- unwto_int_upfill %>%
  group_by(rgn_id) %>%
  filter(all(is.na(tourism_arrivals_ct))) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_ct, data_source, arrivals_gf)

unwto_gapfill_lm_2019_tourism_int <- unwto_int_upfill %>%
  left_join(lm_coef_data_tourism_int) %>% 
  ungroup() %>%
  filter(!(rgn_id %in% c(all_nas_tourism_int$rgn_id))) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_ct, data_source, arrivals_gf, lm_coef_tourism)


## now lets fill in any values down with the linear model average increase per year for tourists
while(na_flag) {
  
  unwto_gapfill_lm_2019_tourism_int <- unwto_gapfill_lm_2019_tourism_int %>%
    group_by(rgn_id) %>%
    arrange(year) %>%
    mutate(
      tourism_arrivals_ct = case_when(
        is.na(tourism_arrivals_ct) & !is.na(lag(tourism_arrivals_ct)) ~ lag(tourism_arrivals_ct) + lm_coef_tourism, # if there is a previous value, but now current value, gapfill using the lag plut the lm_coef_tourism
        TRUE ~ tourism_arrivals_ct
      )
    ) %>%
    mutate(data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "UNWTO - linear model", data_source)) %>%
    mutate(arrivals_gf = ifelse(data_source == "linear model", "gapfilled", arrivals_gf)) %>%
    ungroup()
  
  # Check if there are still NAs left in either column
  na_flag <- any(is.na(unwto_gapfill_lm_2019_tourism_int$tourism_arrivals_ct))
}

# combine back with any regions that had all NAs
unwto_gapfill_lm_2019_tourism_all_int <- unwto_gapfill_lm_2019_tourism_int %>%
  dplyr::select(-lm_coef_tourism) %>%
  rbind(all_nas_tourism_int)

# ==== gapfill for 2020 and 2021 now ==== #
unwto_2020_2021_int <- international_clean_names_agg %>%
  filter(year > 2019) 

unwto_all_gf_int <- unwto_gapfill_lm_2019_tourism_all_int %>%
  rbind(unwto_2020_2021_int) %>%
  group_by(rgn_id) %>%
  arrange(rgn_id, year) %>%
  # apply global average proportional increase or decrease for 2020 and 2021, because of covid pandemic messing up trends...
  mutate(tourism_arrivals_ct = ifelse(year == 2020 & is.na(tourism_arrivals_ct), lag(tourism_arrivals_ct, n = 1) + lag(tourism_arrivals_ct, n = 1)*gf_2020_tourism_int, tourism_arrivals_ct)) %>%
  mutate(tourism_arrivals_ct = ifelse(year == 2021 & is.na(tourism_arrivals_ct), lag(tourism_arrivals_ct, n = 1) + lag(tourism_arrivals_ct, n = 1)*gf_2021_tourism_int, tourism_arrivals_ct)) %>%
  mutate(data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "UNWTO - 2020 and 2021 gapfill method", data_source)) %>%
  mutate(arrivals_gf = ifelse(data_source == "2020 and 2021 gapfill method", "gapfilled", arrivals_gf)) %>%
  filter(year >= 2008) %>% # get only the year we need and beyond
  drop_na(tourism_arrivals_ct) # remove any remaining NAs (any remaining have all NAs for that region)

length(unique(unwto_all_gf_int$rgn_name)) # now only 157 regions after gapfilling and removing regions with only NAs!
```


## Domestic UNWTO Inbound Tourism Arrivals

```{r}
file_path_unwto_domestic_trips <- here::here(unwto_dir, "unwto-domestic-trips-data.xlsx")
unwto_arrivals_dom_trips <- readxl::read_xlsx(file_path_unwto_domestic_trips, skip = 4) # read in the raw domestic trips data

unwto_dom_trips_clean <- unwto_arrivals_dom_trips %>% 
  select(country = `Basic data and indicators`, total_trips = `...6`, subdivision_1 = `...7`, subdivision_2 = `...8`, `1995`:`2021`) %>% # select relevant columns
  fill(country, .direction = "down") %>% # add country name to all data associated with that country
  pivot_longer(cols = c("total_trips", "subdivision_1", "subdivision_2"),
               values_to = "metric",
               values_drop_na = TRUE) %>% # make the metrics into one column
  select(-name) %>% # get rid of the name column since it's just the titles of the metrics which are already there
  select(country, metric, everything()) %>% # reorder things
  replace_with_na_all(condition = ~.x == "..") %>% # swap .. with NAs
  pivot_longer(cols = 3:ncol(.), names_to = "year",
               values_to = "tourism_arrivals_trips") %>% # make the years not columns anymore
  pivot_wider(names_from = metric, values_from = tourism_arrivals_trips) %>%
  mutate(overnights = as.numeric(`Overnights visitors (tourists)`), 
         same_day = as.numeric(`Same-day visitors (excursionists)`), 
         total_trips = as.numeric(`Total trips`),
         tourism_arrivals_trips = as.numeric(NA)) %>% # rename metrics so easier to work with, make numeric, and add a new column to fill with the new calculated values later
  select(country, year, overnights, same_day, total_trips, tourism_arrivals_trips) %>% # select columns needed for analysis (cruise passengers seem to be included in same-day)
  group_by(country, year) %>% # group by county and year
  mutate(
    tourism_arrivals_trips = case_when(
      !is.na(overnights) ~ overnights, # if there is a value, dont gapfill
      is.na(overnights) & !is.na(same_day) & !is.na(total_trips) ~ total_trips - same_day, # gapfill, when there is no data on overnights, fill with total_trips - same day
      TRUE ~ tourism_arrivals_trips # otherwise, NA
    ), # there were 0 situations like this in v2024
    total_trips = case_when(
      !is.na(total_trips) ~ total_trips, 
      is.na(total_trips) & !is.na(same_day) & !is.na(overnights) ~ overnights + same_day,
      TRUE ~ total_trips
    )
  ) %>% # v2024: overnights has 5145 NAs out of 6021
  # v2024: same_day has 5507 NAs out of 6021
  # v2024: total_trips has 5294 NAs
  # v2024: total_trips has 5145 NAs
  mutate(arrivals_method = ifelse(is.na(overnights) & !is.na(same_day) & !is.na(total_trips), "UNWTO - subtraction", NA)) %>%
  mutate(arrivals_gapfilled = ifelse(arrivals_method == "UNWTO - subtraction", "gapfilled", NA)) %>% # prepare a "gapfilled" column to indicate "gapfilled" or NA
  ungroup() %>% # ungroup since not needed anymore
  select(country, year, tourism_arrivals_trips, total_trips, arrivals_method, arrivals_gapfilled) %>% # select only needed columns
  mutate(country = str_to_title(country), # make countries look nice
         tourism_arrivals_trips = round(as.numeric(tourism_arrivals_trips) * 1000),
         total_trips = round(as.numeric(total_trips) * 1000)) # since the units were in thousands

kableExtra::kable(colSums(is.na(unwto_dom_trips_clean))) # to see how many NAs are present within the data
# because we have so many NAs here, this can be used for gapfilling the accomodations data.
```

Now we have the domestic overnight trips data, which can be used to gapfill the domestic accommodation data!

```{r}
# bring in domestic accommodation data from UNWTO
file_path_unwto_domestic_acc <- here::here(unwto_dir, "unwto-domestic-accommodation-data.xlsx")
unwto_arrivals_dom_acc <- readxl::read_xlsx(file_path_unwto_domestic_acc, skip = 4) # read in the raw domestic accomodations data

# clean up the raw xlsx!
unwto_clean_dom <- unwto_arrivals_dom_acc %>% 
  select(country = `Basic data and indicators`, source_arrivals = `...6`, subdivision_1 = `...7`, subdivision_2 = `...8`, `1995`:`2021`) %>% # select relevant columns
  fill(country, .direction = "down") %>% # add country name to all data associated with that country
  fill(source_arrivals, .direction = "down") %>% 
  filter(subdivision_1 %in% "Overnights" | subdivision_1 %in% "Guests") %>% 
  select(-subdivision_2) %>% # get rid of the NA column
  replace_with_na_all(condition = ~.x == "..") %>% # swap ".." with NAs
  mutate(source_arrivals_all = paste(source_arrivals, subdivision_1, sep = ":")) %>% # unite the columns that define the source for easier cleaning and pivoting
  select(country, source_arrivals_all, 4:ncol(.)) %>% # use the . to specify the whole dataframe
  pivot_longer(cols = 3:ncol(.), names_to = "year",
               values_to = "tourism_arrivals_ct") %>% # make the years not columns anymore
  pivot_wider(names_from = source_arrivals_all, values_from = tourism_arrivals_ct) %>%
  select(-c("Total:Guests","Hotels and similar establishments:Guests")) %>% # to keep it consistent with the international data, we only want number of overnights
  mutate(hotel_overnights = as.numeric(`Hotels and similar establishments:Overnights`), 
         total_overnights = as.numeric(`Total:Overnights`),
         tourism_arrivals_ct = as.numeric(NA)) %>% # rename metrics so easier to work with, make numeric, and add a new column to fill with the new calculated values later
  select(country, year, hotel_overnights, total_overnights, tourism_arrivals_ct) %>% # select columns needed for analysis
  group_by(country, year) %>% # group by county and year
  mutate(
    tourism_arrivals_ct = case_when(
      !is.na(hotel_overnights) ~ hotel_overnights, # if there is a value, dont gapfill
      is.na(hotel_overnights) & !is.na(total_overnights) ~ total_overnights, # gapfill, when there is no data on hotel overnights, fill with total overnights
      TRUE ~ tourism_arrivals_ct # otherwise, NA
    )) %>% 
  # v2024: hotel_overnights has 3835 NAs out of 6021 (colSums(is.na(unwto_clean_dom)))
  # v2024: total_overnights has 4614 NAs out of 6021
  # v2024: tourism_arrivals_ct has 3719 NAs out of 6021 
  mutate(arrivals_method = ifelse(is.na(hotel_overnights) & !is.na(total_overnights), "UNWTO - total", NA)) %>%
  mutate(arrivals_gapfilled = ifelse(arrivals_method == "UNWTO - total", "gapfilled", NA)) %>% # prepare a "gapfilled" column to indicate "gapfilled" or NA
  ungroup() %>% # ungroup since not needed anymore
  select(country, year, tourism_arrivals_ct, arrivals_method, arrivals_gapfilled) %>% # select only needed columns
  mutate(country = str_to_title(country), # make countries look nice
         tourism_arrivals_ct = round(as.numeric(tourism_arrivals_ct) * 1000)) # since the units were in thousands

# ---- join with trips data to gapfill countries that do not have domestic overnight accomodations data ---
join_dom_acc_trips <- left_join(unwto_clean_dom, unwto_dom_trips_clean, by = c("country", "year", "arrivals_method","arrivals_gapfilled")) %>% 
  mutate(
    tourism_arrivals_all = case_when(
      !is.na(tourism_arrivals_ct) ~ tourism_arrivals_ct, # if there is a value, dont gapfill
      is.na(tourism_arrivals_ct) & !is.na(tourism_arrivals_trips) ~ tourism_arrivals_trips, # gapfill, when there is no data on hotel overnights from the accommodations data, fill with trip overnights
      is.na(tourism_arrivals_ct) & is.na(tourism_arrivals_trips) & !is.na(total_trips) ~ total_trips, # if the overnight data is not present for either, use total trips to gf
      TRUE ~ tourism_arrivals_ct # otherwise, keep the accommodation overnight data
    )) %>% 
  mutate(data_source = ifelse(is.na(tourism_arrivals_ct) & (!is.na(tourism_arrivals_trips) | !is.na(total_trips)), "UNWTO - trips", arrivals_method)) %>%
  mutate(arrivals_gf = ifelse(data_source == "UNWTO - trips", "gapfilled", arrivals_gapfilled)) %>% 
  select(country, year, tourism_arrivals_all, data_source, arrivals_gf) 

# run name_2_rgn on the domestic data so it has OHI region ids
unwto_clean_names_dom <- name_2_rgn(df_in = join_dom_acc_trips, 
                                        fld_name = 'country',
                                        keep_fld_name = TRUE) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_all, data_source, arrivals_gf) # lost many land locked regions, duplicates were also found so we need to look into that

region_data() # to get rgns_eez

# see whether there are regions that were not properly converted to ohi rgns
setdiff(unwto_clean_names_dom$rgn_name, rgns_eez$rgn_name) # v2024: Kiribati, but it is OHI region 212 and is preserved in `unwto_clean_names_dom`

# ---- determine if there were any duplicates due to name_2_rgn ----

colSums(is.na(unwto_clean_names_dom)) # v2024: 2852 NAs in tourism_arrivals_all
length(unwto_clean_names_dom$tourism_arrivals_all) # within 4833 observations for `tourism_arrivals_all`

# this shows rows that have at least one duplicate, including both the first and subsequent occurrences of each duplicated row. using the tidyverse allows us to see all instances of duplicated data, not just the second and subsequent occurrences (which is what just a duplicated() call would give you).
duplicates <- unwto_clean_names_dom %>%
  group_by(rgn_id, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates$rgn_name)
# v2024: 
# [1] "China" # only one set of values, the duplicates are NA                                             
# [2] "Guadeloupe and Martinique" # all NAs                         
# [3] "Northern Mariana Islands and Guam"  # all NAs                
# [4] "Montenegro" # values dont overlap, can aggregate                                        
# [5] "Puerto Rico and Virgin Islands of the United States" # all NAs

# therefore, move forward by aggregating by region id and year
## make a function to avoid losing NAs (they turn to 0s) when there are values to aggregate with
sum_with_na <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm = TRUE))
  }
}

# now use the function to aggreggate by region and year!
domestic_clean_names_agg <- unwto_clean_names_dom %>% 
  group_by(rgn_id, rgn_name, year) %>% 
  summarize(
    tourism_arrivals_all = sum_with_na(tourism_arrivals_all), # put each column in the function, so no NAs are lost
    data_source = first(data_source, na_rm = TRUE),
    arrivals_gf = first(arrivals_gf, na_rm = TRUE),
    .groups = "drop" # to ungroup, fixes warning from summarise to use reframe (when it is not applicable here)
  ) %>% 
  mutate(
    data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_all), "UNWTO", data_source) # so that if there is data from UNWTO that was not gapfilled, the data_source is accurate
  )

# double check to test
duplicates_2 <- domestic_clean_names_agg %>%
  group_by(rgn_id, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates_2$rgn_name) # v2024: 0! yay

# see which/how many regions have values
check_domestic <- domestic_clean_names_agg %>% 
  filter(!is.na(tourism_arrivals_all))
length(unique(check_domestic$rgn_id)) # v2024: 101 regions present
```

### Gapfilling Domestic Arrivals

```{r}
# gapfill for domestic arrivals!
# downfill then upfill missing values using a linear model of the average increase per years across all years of data for 1995-2019
# for 2020 and 2021, use the global average proportion increase or decrease and add to the previous years value

# reminder: 
# `tourism_arrivals_all` is the tourism arrivals count from the DOMESTIC data, which is primarily hotel overnights (accommodation) data, and was gapfilled using domestic trip data also from UNWTO.

test_2021_dom <- domestic_clean_names_agg %>%
  filter(year %in% c(2020, 2021)) %>%
  filter(!is.na(tourism_arrivals_all)) %>%
  pivot_wider(names_from = year, 
              values_from = tourism_arrivals_all,
              names_prefix = "int_tourism_") %>%
  mutate(tourism_arrivals_all_diff = (int_tourism_2021 - int_tourism_2020)/int_tourism_2020)

gf_2021_tourism_dom <- mean(test_2021_dom$tourism_arrivals_all_diff, na.rm = TRUE) # global average increase for 2021 tourism_arrivals_ct
# v2024: 0.2941438

# ok, lets use these values to gapfill 2021 if it is NA and 2020 exists. So increase by X proportion (0.294)

test_2020_dom <- domestic_clean_names_agg %>%
  filter(year %in% c(2019, 2020)) %>%
  filter(!is.na(tourism_arrivals_all)) %>%
  pivot_wider(names_from = year, 
              values_from = tourism_arrivals_all,
              names_prefix = "int_tourism_") %>%
  mutate(tourism_arrivals_all_diff = (int_tourism_2020 - int_tourism_2019)/int_tourism_2019)
 
gf_2020_tourism_dom <- mean(test_2020_dom$tourism_arrivals_all_diff, na.rm = TRUE) # global average decrease for 2020 toursism
# v2024: -0.3473281

# ok, lets use these values to gapfill 2020 if it is NA and 2019 exists. So decrease by X proportion ~35%

# upfill earlier years that do not have data with nearby year data
unwto_upfill_dom <-  domestic_clean_names_agg %>%
  filter(year < 2020) %>%
  group_by(rgn_id) %>%
  arrange(rgn_id, year) %>%
  tidyr::fill(tourism_arrivals_all, .direction = "up") %>% # fill in any values that are empty from early years with values from the nearest year. Doing this because doesn't make sense to add earlier years based on a trend 
  mutate(
    arrivals_gf = ifelse(is.na(data_source) & !is.na(tourism_arrivals_all), "gapfilled", arrivals_gf),
    data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_all), "UNWTO - nearby year", data_source)
  )

## calculate regional average increase or decrease in number of tourism arrivals
lm_coef_data_tourism_dom <- domestic_clean_names_agg %>%
  filter(!(year %in% c(2020, 2021))) %>%
  group_by(rgn_id) %>%
  filter(!is.na(tourism_arrivals_all)) %>%
  filter(!rgn_id %in% c("136","205")) %>% # Guatemala and Myanmar only have one value, will need to be dropped!!
  summarize(lm_coef_tourism = lm(tourism_arrivals_all ~ year)$coefficients[2])

# Initialize a flag to check if there are still NAs
na_flag <- TRUE

# filter out any regions with all nas for each year, as these can't be gapfilled
all_nas_tourism_dom <- unwto_upfill_dom %>%
  group_by(rgn_id) %>%
  filter(all(is.na(tourism_arrivals_all))) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_all, data_source, arrivals_gf)

unwto_gapfill_lm_2019_tourism_dom <- unwto_upfill_dom %>%
  left_join(lm_coef_data_tourism_dom) %>% 
  ungroup() %>%
  filter(!(rgn_id %in% c(all_nas_tourism_dom$rgn_id))) %>%
  dplyr::select(rgn_id, rgn_name, year, tourism_arrivals_all, data_source, arrivals_gf, lm_coef_tourism)


## now lets fill in any values down with the linear model average increase per year for tourists
while(na_flag) {
  
  unwto_gapfill_lm_2019_tourism_dom <- unwto_gapfill_lm_2019_tourism_dom %>%
    group_by(rgn_id) %>%
    arrange(year) %>%
    mutate(
      tourism_arrivals_all = case_when(
        is.na(tourism_arrivals_all) & !is.na(lag(tourism_arrivals_all)) ~ lag(tourism_arrivals_all) + lm_coef_tourism, # if there is a previous value, but now current value, gapfill using the lag plut the lm_coef_tourism
        TRUE ~ tourism_arrivals_all
      )
    ) %>%
    mutate(data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_all), "UNWTO - linear model", data_source)) %>%
    mutate(arrivals_gf = ifelse(data_source == "linear model", "gapfilled", arrivals_gf)) %>%
    ungroup()
  
  # Check if there are still NAs left in either column
  na_flag <- any(is.na(unwto_gapfill_lm_2019_tourism_dom$tourism_arrivals_all))
} ######stopped here

# combine back with any regions that had all NAs
unwto_gapfill_lm_2019_tourism_all_int <- unwto_gapfill_lm_2019_tourism_int %>%
  dplyr::select(-lm_coef_tourism) %>%
  rbind(all_nas_tourism_int)

# ==== gapfill for 2020 and 2021 now ==== #
unwto_2020_2021_int <- international_clean_names_agg %>%
  filter(year > 2019) 

unwto_all_gf_int <- unwto_gapfill_lm_2019_tourism_all_int %>%
  rbind(unwto_2020_2021_int) %>%
  group_by(rgn_id) %>%
  arrange(rgn_id, year) %>%
  # apply global average proportional increase or decrease for 2020 and 2021, because of covid pandemic messing up trends...
  mutate(tourism_arrivals_ct = ifelse(year == 2020 & is.na(tourism_arrivals_ct), lag(tourism_arrivals_ct, n = 1) + lag(tourism_arrivals_ct, n = 1)*gf_2020_tourism_int, tourism_arrivals_ct)) %>%
  mutate(tourism_arrivals_ct = ifelse(year == 2021 & is.na(tourism_arrivals_ct), lag(tourism_arrivals_ct, n = 1) + lag(tourism_arrivals_ct, n = 1)*gf_2021_tourism_int, tourism_arrivals_ct)) %>%
  mutate(data_source = ifelse(is.na(data_source) & !is.na(tourism_arrivals_ct), "UNWTO - 2020 and 2021 gapfill method", data_source)) %>%
  mutate(arrivals_gf = ifelse(data_source == "2020 and 2021 gapfill method", "gapfilled", arrivals_gf)) %>%
  filter(year >= 2008) %>% # get only the year we need and beyond
  drop_na(tourism_arrivals_ct) # remove any remaining NAs (any remaining have all NAs for that region)

length(unique(unwto_all_gf_int$rgn_name)) # now only 157 regions after gapfilling and removing regions with only NAs!
```


## Aggregating International and Domestic Values

Let's use `setdiff()` and see which regions that have values overlap between the international and domestic data.  If it is a reasonable amount, I am considering aggregating international and domestic and then for regions that have one or the other, leaving it as such.

```{r}
domestic_rgns <- unique(check_domestic$rgn_id) # domestic OHI regions with data
international_rgns <- unique(check_international$rgn_id) # international OHI regions with data

# regions w values in domestic but not in international
only_in_domestic <- setdiff(domestic_ids, international_ids) 
# [1]  14  45  84 137 195 210

# regions w values in international but not in domestic
only_in_international <- setdiff(international_ids, domestic_ids)
#  [1]   4   6   7   8   9  11  13  17
#  [9]  19  26  28  32  37  39  46  47
# [17]  48  49  51  55  57  58  64  65
# [25]  78  90  96 103 108 110 111 113
# [33] 114 115 116 117 118 119 120 121
# [41] 122 123 124 125 129 130 133 147
# [49] 148 151 152 153 154 155 157 166
# [57] 167 168 169 190 193 194 198 204
# [65] 208 212 220 237 244 245 247 250

# regions that have values in both datasets
in_both <- intersect(domestic_ids, international_ids)
#  [1]   5  15  16  18  20  24  25  31
#  [9]  40  41  42  43  50  52  53  54
# [17]  56  59  61  62  66  67  68  69
# [25]  70  71  72  73  74  75  76  77
# [33]  79  80  81  82  98  99 100 101
# [41] 102 106 112 126 127 131 132 134
# [49] 135 136 138 139 143 162 163 164
# [57] 171 172 173 174 175 176 177 178
# [65] 179 180 181 182 183 184 185 186
# [73] 187 188 189 191 196 197 199 200
# [81] 202 203 205 206 207 209 214 215
# [89] 216 218 222 223 224 231 232

cat("Regions only in domestic data:", length(only_in_domestic), "\n") # Regions only in domestic data: 6 
cat("Regions only in international data:", length(only_in_international), "\n") # Regions only in international data: 72
cat("Regions in both datasets:", length(in_both), "\n") # Regions in both datasets: 95 
cat("Total unique regions:", length(union(domestic_ids, international_ids)), "\n") # Total unique regions: 173
```

Let's move on, as we decided that it makes sense to aggregate the international and domestic data! This is because we are creating a score for each region, and only comparing across regions once that score has been calculated (versus ranking regions and comparing them to each other based on tourism density, not the rescaled score).  So the more accurate the data, the better!


```{r}
# aggregating international and domestic data by region and year
arrivals_aggregated <- left_join(international_clean_names_agg, domestic_clean_names_agg, by = c("rgn_id","rgn_name","year"))
```

Now, lets bring in our tourism data and calculate the coastal proportion! The coastal proportion will then be multiplied by the arrivals value per region and year. Finally, the coastal area will be 

```{r, eval=FALSE}
# ------- coastal population data -----------
# In 2024, the coastal population data was fixed! This is great.  We will now use `globalprep/mar_prs_population/v2024/output/mar_pop_25mi.csv`.
coastal_pop_data <- read_csv(here("globalprep","mar_prs_population","v2024","output","mar_pop_25mi.csv")) ## read in coastal population data from other data layer

coastal_pop_data_fill <- coastal_pop_data %>%
  filter(year == 2020) %>%
  mutate(year = 2021) %>%
  rbind(coastal_pop_data) # add 2021 data year, just using year 2020

length(unique(coastal_pop_data_fill$rgn_id)) # v2024: 220 regions available! Any region without data is uninhabited.

# bring in World Bank Global population data from `/home/shares/ohi/git-annex/globalprep/_raw_data/WorldBank/d2024/WorldBank_global_country_population_1960-2023`
total_population_wb <- read_csv(here(raw_data_dir, "WorldBank", data_dir_version_year, "WorldBank_global_country_population_1960-2023","API_SP.POP.TOTL_DS2_en_csv_v2_1584446.csv"), skip = 3) %>% 
  pivot_longer(cols = 5:ncol(.), names_to = "year",
               values_to = "total_pop") %>% 
  janitor::clean_names()
```


```{r, eval=FALSE}
# ------- total global population data -----------
# give ohi rgns to world bank data
total_population_rgns <- name_2_rgn(df_in = total_population_wb, 
                                        fld_name = 'country_name',
                                        keep_fld_name = TRUE)

# drop Macao and Hong Kong, since we only need mainland China
# Northern Mariana Islands and Guam, Puerto Rico and Virgin Islands of the United States: aggregate!
total_population_final <- total_population_rgns %>%
  filter(country_name != "Hong Kong SAR, China" & country_name != "Macao SAR, China") %>% # only keep mainland, not city specific data
  group_by(rgn_id, rgn_name, year) %>% 
  summarise(
    total_pop = sum(total_pop, na.rm = TRUE), # aggregate by rgn and year
    .groups = "drop" # ungroup
  )

length(unique(total_population_final$rgn_name)) # v2024: 166 regions available of total population data!
```


```{r, eval=FALSE}
# ------- coastal area data -----------
# read in area of coastline data
inland_filepath <- here::here("globalprep", "lsp", version_year, "int", "area_protected_1km.csv")
inland_data <- read_csv(inland_filepath)
offshore_filepath <- here::here("globalprep", "lsp", version_year, "int", "area_protected_3nm.csv")
offshore_data <- read_csv(offshore_filepath)

# get combined value of inland and offshore for each ohi region
inland_offshore <- inland_data %>%
  left_join(offshore_data, by = join_by(rgn_id, year, rgn_name)) %>%
  select(rgn_id, year, a_tot_km2.x, a_tot_km2.y) %>%
  group_by(rgn_id, year) %>%
  mutate(total_inland_offshore_area = a_tot_km2.x + a_tot_km2.y,
         year = as.numeric(year)) %>%
  select(rgn_id, year, total_inland_offshore_area)
```


```{r, eval=FALSE}
# ============= join the tourism arrival, coastal population, and total population data =================== #

options(scipen = 999) # to change options and avoid scientific notation

arrivals_prop_df <- left_join(arrivals_aggregated, total_population_final, by = c("rgn_id","rgn_name","year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  left_join(., coastal_pop_data_fill, by = c("rgn_id","year")) %>% 
  left_join(., inland_offshore, by = c("rgn_id","year")) %>% 
  filter(year %in% 2000:2021) %>% # to keep consistent with the coastal pop data
  mutate(coastal_prop = (popsum/total_pop)) %>% # values that are greater than 1 are because of inconsistencies between World Bank and the coastal population data.  However, those with a value greater than 1 are mostly islands or small regions with a large coastal area.  In that case, we could say 100% of the tourism could be coastally related.
  mutate(coastal_prop = ifelse(coastal_prop > 1, 1, coastal_prop)) # for rgns and years in which the total_population is greater than 1, we cap it at 1 

ap_final <- arrivals_prop_df %>% 
  mutate(ap = (tourism_arrivals_ct * coastal_prop)/(total_inland_offshore_area)) # this final `Ap` (arrivals proportion) is in the units of (tourists/km2).  Aka, a higher proportion indicates that there is a larger number of tourists per km2 of coastal area, and there is therefore a higher tourism density.  This number is now ready to be divided by `Sr`, or the sustainability of tourism.
```

```{r}
# interactive plot of the arrival proportions
ap_plot <- ap_final %>%
  plot_ly(x = ~year, y = ~ap, color = ~rgn_name, 
          type = "scatter", mode = "lines") %>%
  layout(title = "Global Arrival proportions (Ap) by Region",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Arrival proportion (Ap)"))
ap_plot
```


<!-- ### Divide tourist arrival count by total arrival count -->

<!-- -   In 2023, we considered rescaling by coastal area and/or coastal population, however, we decided to scrap this, as it was unduly harming scores for larger regions. We will just use the tourist arrival/total arrivals metric. -->
<!-- -   Because of some gapfilling and uncertainty in the data, some regions reported higher tourist arrivals than total arrivals. Due to this, we sometimes have ratios \>1. To deal with this, we just give those regions a ratio of 1. -->
<!-- -   Some regions reported only tourist arrivals, and no total arrivals. Because of this, we gapfilled any that didn't have total arrivals with the tourist arrivals. This isn't ideal, but was the best we could do. -->

<!-- ```{r, eval=FALSE} -->
<!-- # divide the number of tourist arrivals by area of coastline and population to get the proportion -->
<!-- tourism_props <- unwto_all_gf %>% -->
<!--     mutate(year = as.numeric(year)) %>% -->
<!--  # left_join(offshore_data, by = c("rgn_id", "year")) %>% -->
<!--   left_join(coastal_pop_data_fill, by = c("rgn_id", "year")) %>% -->
<!--   # mutate(Ap = ((tourism_arrivals_ct/total_inland_offshore_area)/popsum)) %>% -->
<!--   mutate(total = ifelse(is.na(total), tourism_arrivals_ct, total)) %>% -->
<!--   mutate(Ap = tourism_arrivals_ct/total) %>% -->
<!--   filter(year >= 2008) %>% # filter to the years we are interested in for any data that isn't filtered yet -->
<!--   filter(popsum > 0) %>% # filter out no pop regions -->
<!--   left_join(rgns_eez) %>% -->
<!--   mutate(Ap = ifelse(Ap > 1, 1, Ap)) # if greater than 1 make 1 -->

<!-- # check out things so far -->
<!-- summary(tourism_props) # should be 0 NAs for Ap if gapfilling worked -->
<!-- hist(tourism_props$Ap) -->

<!-- ``` -->

<!-- ```{r, eval=FALSE} -->

<!-- # make rescaled column the actual value column -->
<!-- tourism_props_rescaled <- tourism_props %>% -->
<!--   dplyr::select(rgn_id, year, arrivals_method, arrivals_gapfilled, Ap) -->

<!-- ``` -->

### Removing low population / uninhabited regions

```{r, eval=FALSE}
### after gap-filling, make sure low/uninhabited regions are NA
# create df for unpopulated/low populated regions
low_pop()
low_pop <- low_pop %>%
  filter(est_population < 3000 | is.na(est_population)) %>%  # filter out regions that have populations > or equal to 3000 and keep NA values 
  rename(rgn_label = rgn_nam)

summary(tourism_props_rescaled)
# v2020 371 NAs
# v2022 114 NAs
# v2023 0 NAs (because of gapfilling)

# make sure all the NAs are uninhabited regions
tourism_props_nas <- tourism_props_rescaled %>% 
  filter(is.na(Ap)) %>% 
  select(rgn_id, year) %>% # v2023: was rgn_id, year, r1_label, r2_label, rgn_label but we did not gapfill by georegion so did not have the last 3
  left_join(low_pop, by = "rgn_id") # v2023: was by = c("rgn_id", "rgn_label") but we did not gapfill by georegion so did not have rgn_label

tourism_props_nas %>% 
  filter(Inhabited == 0 & !is.na(est_population)) %>% 
  nrow() # 0 ✓

max(tourism_props_nas$est_population, na.rm=TRUE) < 3000 # should be true

# make sure all the uninhabited regions are NA and then drop them (all NA regions get added back as a last step just in case its relevant)
tourism_props_rescaled <- tourism_props_rescaled %>% 
  mutate(Ap = ifelse(rgn_id %in% low_pop$rgn_id, NA, Ap)) 


# check NAs once more 
summary(tourism_props_rescaled)
# v2019: Adding the low pop df identifies 13 additional regions that should be NA instead of gapfilled, taking the total number of NAs in the data set from 245 to 700
# v2020: Adding the low pop df takes the total number of NAs in the data set from 371 to 832
# v2022: Adding the low pop df takes the total number of NAs in the data set from 14 to 40
# v2023: Adding the low pop df takes the total number of NAs in the data set from 0 to 14

# after checking NAs, get rid of them
tourism_props_rescaled <- tourism_props_rescaled %>% 
  drop_na(Ap)
```

### Write output files

```{r, eval=FALSE}
# we want to make sure all OHI regions are present in the data, even if could not calculate Ap (this will add back in what we removed above as well as any region not present in the data)
year_range <- unique(tourism_props_rescaled$year) # get the year range of Ap
year_range_df <- data.frame(year = year_range) # make it a dataframe

# save gapfill info
tourism_props_gf_to_write <- tourism_props_rescaled %>%
  select(-Ap) # don't need actual values for the gapfill information

write_csv(tourism_props_gf_to_write, here(paste0("globalprep/tr/v", version_year, "/output/tr_arrivals_props_tourism_gf.csv")))

# save gap-filled data
tourism_props_to_write <- tourism_props_rescaled %>%
  select(rgn_id, year, Ap) # don't need gf info here, just the values

write_csv(tourism_props_to_write, here(paste0("globalprep/tr/v", version_year, "/output/tr_arrivals_props_tourism.csv")))
```

### Look at changes in recent years

We would expect for tourism jobs to decrease across the board from 2019 and 2020 given the pandemic, and likely see a rebound to some extent between 2020 and 2021 — let's make sure that's reflected in our results.

```{r, eval=FALSE}
tourism_props_compare <- tourism_props_to_write %>%
  mutate(year = as.numeric(as.character(year))) %>%
  filter(year >= 2019) %>%
  pivot_wider(names_from = year, values_from = Ap)

# compare 2019 and 2020
plot(tourism_props_compare$"2019", tourism_props_compare$"2020",
     xlab = "v2023 2019 Arrivals Proportion", ylab = "v2023 2020 Arrivals Proportion")
abline(0, 1) # more data below the line

# compare 2020 and 2021
plot(tourism_props_compare$"2020", tourism_props_compare$"2021",
     xlab = "v2023 2020 Arrivals Proportion", ylab = "v2023 2021 Arrivals Proportion")
abline(0, 1) # more data above the line
```

Everything looks reasonable.

### Look at changes vs. previous data source (v2023)

```{r, eval=FALSE}
new_data <- read_csv(paste0("globalprep/tr/v", version_year, "/output/tr_arrivals_props_tourism.csv"))
old_data <- read_csv(paste0("globalprep/tr/v", prev_ver_yr, "/output/tr_jobs_pct_tourism.csv"))

compare_common_data <- new_data %>%
  left_join(old_data, by = c("rgn_id", "year")) %>%
  drop_na()

plot(compare_common_data$Ep, compare_common_data$Ap,
     xlab = "v2022 Employment Proportion", ylab = "v2023 Arrivals Proportion")
abline(0, 1)




compare_common_data_2021 <- new_data %>%
  left_join(old_data, by = c("rgn_id", "year")) %>%
  drop_na() %>%
  filter(year == 2021)

plot(compare_common_data_2021$Ap, compare_common_data_2021$Ep,
     xlab = "v2023 Arrivals Proportion", ylab = "v2022 Employment Proportion")
abline(0, 1)

compare_common_data_2020 <- new_data %>%
  left_join(old_data, by = c("rgn_id", "year")) %>%
  drop_na() %>%
  filter(year == 2020)

plot(compare_common_data_2020$Ap, compare_common_data_2020$Ep,
     xlab = "v2023 Arrivals Proportion", ylab = "v2022 Employment Proportion")
abline(0, 1)

compare_common_data_2019 <- new_data %>%
  left_join(old_data, by = c("rgn_id", "year")) %>%
  drop_na() %>%
  filter(year == 2019)

plot(compare_common_data_2019$Ap, compare_common_data_2019$Ep,
     xlab = "v2023 Arrivals Proportion", ylab = "v2022 Employment Proportion")
abline(0, 1)

compare_common_data_2015 <- new_data %>%
  left_join(old_data, by = c("rgn_id", "year")) %>%
  drop_na() %>%
  filter(year == 2015)

plot(compare_common_data_2015$Ap, compare_common_data_2015$Ep,
     xlab = "v2023 Arrivals Proportion", ylab = "v2022 Employment Proportion")
abline(0, 1)
```

### Check out some specific countries (v2023) -- this was for exploring changes in methodology, can skip or use/edit parts in future years

```{r, eval=FALSE}
# check some countries that changed a lot in v2023's first push to global
check_countries_graph <- tourism_props %>% 
  filter(rgn_id == 24 | rgn_id == 51 | rgn_id == 189 | rgn_id == 118 | rgn_id == 31) %>%
  mutate(rgn_id_plot = as.factor(rgn_id),
         year_plot = year)

Ap_graph <- ggplot(check_countries_graph, aes(x = year_plot, y = Ap, color = rgn_id_plot)) +
  geom_line() +
  theme_minimal() +
  labs(x = "",
       color = "Region ID")


Ap_graph

countries_in_2021 <- check_countries_graph %>%
  filter(year == "2021") %>%
  left_join(rgns_eez, by = "rgn_id") %>%
  select(-year_plot)

library(kableExtra)
kable(countries_in_2021) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

# Ts: Tourism sustainability

These data are from the World Economic Forum's "Travel and Tourism Development Index" (<https://www.weforum.org/publications/travel-tourism-development-index-2024/>) Downloaded 08/20/2024.

See mazu: \_raw_data/WEF-Economics/ for more details and the raw data.

The TTDI was formerly the TTCI which was a similar index, but unfortunately not comparable. v2024's updated TTDI uses data up through 2023 to calculate their sustainability scores.

These data are gapfilled using gdppcppp and UN georegion information (see next section for obtaining and preparing these data).

```{r WEF processing, eval=FALSE}
# bring in TTDI pdf! Download it to your computer, place it in Mazu
ttdi_file <- "WEF_Travel_and_Tourism_Development_Index_2024.pdf" # the file name
ttdi_raw <- pdf_text(here::here(raw_data_dir, "WEF-Economics", data_dir_version_year, ttdi_file)) # using the package `pdftools`, read in using pdf_text()

sr_scores_table <- ttdi_raw[[11]] %>% # page number of the pdf that has the data we want (aka the sustainability scores)
  as.data.frame() # make the page as a data frame.  Be warned, the data is all there! You just can't see it yet.

columns = c("rank","economy","score","rank_change","score_change","diff_avg","rank_2","economy_2","score_2","rank_change_2","score_change_2","diff_avg_2", "rank_3","economy_3","score_3","rank_change_3","score_change_3","diff_avg_3") # mirror the pdf and create clean column names

sr_scores <- sr_scores_table %>% 
  separate_rows('.', sep = "\n") %>%  # separate the rows: \n separates by moving down to the next line without returning to the beginning of the line
  rename(main_column = ".") %>% # rename the only column as main_column instead of just the period
  slice(16:55) %>% # choose the rows that have the table in them (disregard the page title etc)
  separate_wider_position(
    main_column, # the issue here is that there is not a common delimiter, and the three columns blur together. So, we can use `separate_wider_position()`, which allows you to instead define column breaks by the number of characters, which includes spaces
    widths = c(rank = 2, economy = 24, score = 4, rank_change = 3, score_change = 5, diff_avg = 19,
               rank_2 = 7, economy_2 = 24, score_2 = 5, rank_change_2 = 3, score_change_2 = 7, diff_avg_2 = 19,
               rank_3 = 9, economy_3 = 26, score_3 = 8, rank_change_3 = 1, score_change_3 = 1, diff_avg_3 = 1), # these numbers were found through trial and error, but are always at least the length of the longest character string
    names_sep = NULL, # default, to avoid renaming the column names
    names_repair = "check_unique", #the default), no name repair, but checks they are unique
    too_many = "debug", # will add additional columns to the output to help you locate and resolve the underlying problem
    too_few = "debug", # adds additional columns to the output to help you locate and resolve the underlying problem... these will not remain in final code
    cols_remove = FALSE # always FALSE if too_few or too_many are set to "debug"
  ) %>% 
  select(rank, economy, score,  rank_2, economy_2, score_2, rank_3, economy_3, score_3) %>% # to get rid of debugging columns and rankings
  mutate(
    economy = str_trim(economy, side = "both"), # remove the whitespace before and after the string
    rank_2 = str_trim(rank_2, side = "both"),
    economy_3 = str_trim(economy_3, side = "both"),
    rank_3 = str_trim(rank_3, side = "both"),
    score_3 = str_trim(score_3, side = "both"),
    economy_3 = str_replace(economy_3, pattern = "Trinidad and Tobago 3.5", replacement = "Trinidad and Tobago"), # this line specifically had issues with spacing, and no matter what it always had bleeding from the score.  Therefore, the economy name and score were fixed using str_replace.
    score_3 = str_replace(score_3, pattern = "2 -5", replacement = "3.52") 
    )

# now, we have a dataframe that contains three columns we would like to make into one. let's subset and create three dfs that each have the same name (they couldnt have the same name beforehand because they were in the same df), then rbind them together to have one df of all regions and their scores.
sr_scores_sub1 <- sr_scores %>% 
  select(rank, economy, score)

sr_scores_sub2 <- sr_scores %>% 
  select(rank_2, economy_2, score_2) %>% 
  rename(
    rank = rank_2,
    economy = economy_2,
    score = score_2
  )

sr_scores_sub3 <- sr_scores %>% 
  select(rank_3, economy_3, score_3) %>% 
    rename(
    rank = rank_3,
    economy = economy_3,
    score = score_3
  )

# rbind them together and clean up -- we do not need the column "rank", it was just preserved until now as an easy way to ensure a clean bind
sr_scores_clean <- rbind(sr_scores_sub1, sr_scores_sub2, sr_scores_sub3) %>% 
  select(-rank) %>% # remove rank
  filter(!is.na(score)) %>%  # remove NA, leaves 119 regions, good
  mutate(
    score = as.numeric(score), # ensure it is numeric
    sr = ((score - 1)/5) # rescale the score between 0 and 1
         ) 


# run name_2_gn so each economy receives an OHI rgn name and id  
ttdi_rgn <- name_2_rgn(df_in = sr_scores_clean, 
                       fld_name = 'economy',
                       keep_fld_name = TRUE) %>% # v2024: no duplicates! looks good
  select(rgn_id, rgn_name, score, sr)

### Save TTDI data file
write_csv(ttdi_rgn, here(int_dir,"wef_ttdi.csv")) # save to int folder
```

## Preparing the gdppcppp data:

These data are used to gapfill missing values in tourism sustainability. Most of the data are from the World Bank, but CIA data fill some gaps (CIA data is available for only the most recent year).

The Artisanal Opportunities goal uses gdppcppp data, so we will get the data that was processed for that goal.

```{r worldbank, eval=FALSE}
wb <- read_csv(here::here("globalprep","ao", version_year, "intermediate","gdppcppp_ohi.csv")) %>%
  dplyr::select(rgn_id, year, value)

```

# COME BACK AFTER AGGREGATING
CIA data are used to fill in missing gaps in the gdppcppp data (<https://www.cia.gov/the-world-factbook/field/real-gdp-per-capita/country-comparison>)

Downloaded: 07/05/2022

See README on the raw folder for instructions on how to download this data.

The following code is used to prepare these data for OHI:

```{r cia gdp, eval=FALSE}

cia_gdp <- read.csv(here(paste0("globalprep/tr/v", version_year, "/raw/cia_gdp_pc_ppp.csv"))) %>% 
  # remove dollar signs and commas and convert to numeric
  mutate(value = as.numeric(gsub("[$,]", "", value))) %>% 
  select(name, value) %>% 
  rename(country = name, pcgdp_cia = value)

 ## Data reported in a lower resolution than OHI regions
splits <- data.frame(country = "Saint Helena, Ascension, and Tristan da Cunha", 
                     country2 = c("Saint Helena", "Ascension","Tristan da Cunha"))

cia_gdp <- cia_gdp %>%
  left_join(splits, by='country') %>%
  mutate(country2 = ifelse(is.na(country2), country, country2)) %>%
  select(country = country2, pcgdp_cia)

cia_gdp_rgn <- name_2_rgn(df_in = cia_gdp, 
                       fld_name='country')

### Duplicated regions: Collapse regions after weighting by population (regions we include as a single region) - 

population_weights <- data.frame(country = c("Virgin Islands", "Puerto Rico",
                                             "China", "Hong Kong", "Macau",
                                             "Guam", "Northern Mariana Islands"),
                                 # from world bank - updated v2022
                                 population = c(105870, 3263584, 1412360000,
                                                7413100, 658391, 170184, 57910))

cia_gdp_rgn <- cia_gdp_rgn %>%
  left_join(population_weights, by="country") %>%
  mutate(population = ifelse(is.na(population), 1, population)) %>%
  group_by(rgn_id) %>%
  summarize(pcgdp_cia = weighted.mean(pcgdp_cia, population)) %>%
  ungroup() %>%
  filter(rgn_id <= 250) %>%
  select(rgn_id, pcgdp_cia)

write.csv(cia_gdp_rgn, here(paste0("globalprep/tr/v", version_year, "/intermediate/wb_rgn_cia_GDPPCPPP.csv")), row.names=FALSE)

```

The following code combines the two gdp datasets and gapfills missing regions using UN georegions.

If there is no World Bank gdppcppp data (pcgdp), the CIA data is used (pcgdp_cia). The pcgdp2 variable includes both the World Bank and CIA data (with CIA data only used if there is not World Bank data). The remaining data are estimated using UN geopolitical regions. Ideally, the mean gdppcppp value is calculated at the r2 scale (gdp_pred_r2) using regions within each class with gdppcppp data. If there were not enough regions with data at the r2 scale, the average at the r1 scale was used (gdp_pred_r1). The gdp_all variable combines all estimates using the following heirarchy: World Bank -\> CIA -\> estimated using mean from r2 UN geopolitical regions -\> estimated using mean from r1 UN geopolitical regions.

```{r gapfill gdp, eval=FALSE}

### world bank gdp data
gdppcppp <- wb %>%
  select(rgn_id, year, pcgdp = value)

### cia gdp data
gdppcppp2 <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/wb_rgn_cia_GDPPCPPP.csv")))


### Use WB data, but if missing, use pcgdp_cia.
### combine with UN georegion data
years <- data.frame(year = min(gdppcppp$year):max(gdppcppp$year))

georegions <- ohicore::georegions

regions <- georegions %>%
  left_join(georegion_labels, by = 'rgn_id')

gdp_raw <- merge(years, regions, by=NULL) %>%
   left_join(gdppcppp, by = c('rgn_id', 'year')) %>%
  left_join(gdppcppp2, by = c("rgn_id")) 

## quick compare to make sure the CIA and World Bank data are compatible
plot(gdp_raw$pcgdp[gdp_raw$year==2021], gdp_raw$pcgdp_cia[gdp_raw$year==2021])
abline(0,1, col="red")
# a few minor outliers but overall looks good

gdp_raw <- gdp_raw %>%
  mutate(pcgdp2 = ifelse(is.na(pcgdp), pcgdp_cia, pcgdp))

## Calculating the means across different geopolitical levels (e.g. r2, r1)
gdp_raw <- gdp_raw %>%
  group_by(r2, year) %>%
  mutate(gdp_pred_r2 = mean(pcgdp2, na.rm=TRUE)) %>%
  ungroup() %>%
  group_by(r1, year) %>%
  mutate(gdp_pred_r1 = mean(pcgdp2, na.rm=TRUE)) %>%
  ungroup() 

gdp_raw_gf <- gdp_raw %>%
  mutate(gdp_all = ifelse(is.na(pcgdp2), gdp_pred_r2, pcgdp2)) %>%
  mutate(gdp_all = ifelse(is.na(gdp_all), gdp_pred_r1, gdp_all)) %>%
  mutate(gapfilled = ifelse(is.na(pcgdp2) & !is.na(gdp_all), "gapfilled", NA)) %>%
  mutate(method = ifelse(is.na(pcgdp2) & !is.na(gdp_pred_r2), "UN georegion (r2)", NA)) %>%
  mutate(method = ifelse(is.na(pcgdp2) & is.na(gdp_pred_r2) & !is.na(gdp_pred_r1), "UN georegion (r1)", method)) 

write_csv(gdp_raw_gf, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_raw_gf.csv")))

gdp_data_gf <- gdp_raw_gf %>%
  select(rgn_id, year, gapfilled, method) 

write_csv(gdp_data_gf, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_gf.csv")))

gdp_data <- gdp_raw_gf %>%
  select(rgn_id, year, pcgdp = gdp_all)

write_csv(gdp_data, here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp.csv")))

```

The final step is gapfilling the Sustainability data using a linear model with gdppcppp and UN geopolitical regions as predictor variables.

```{r, eval=FALSE}

sust <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/wef_ttdi.csv")), stringsAsFactors = FALSE)

### don't need to gapfill data without tourism data:
## Most recent tourism data is 2019.  

ap_gf <- read.csv(here(paste0("globalprep/tr/v", version_year, "/output/tr_arrivals_props_tourism.csv"))) %>%
  # filter(year == 2021) %>%
  select(rgn_id, Ap, year) %>%
  filter(!is.na(Ap))

# gdp dataframe prepared above (World Bank, CIA, and gapfilled gdp data)
gdp_raw_gf <- read.csv(here(paste0("globalprep/tr/v", version_year, "/intermediate/gdp_raw_gf.csv")), stringsAsFactors = FALSE) %>% 
  # filter(year == 2021) %>%
  select(rgn_id, r0_label, r1_label, r2_label, rgn_label,
         pcgdp, pcgdp_cia, pcgdp2, gdp_all, year) 

tr_sust <- gdp_raw_gf %>%
           left_join(sust, by = c("rgn_id", "year")) %>%
          left_join(ap_gf, by = c("rgn_id", "year")) %>%
          rename(S_score = score) %>%
          filter(rgn_id != 213)

### Add gapfill flag variable 
## Reminder:
## pcgdp2: includes both the World Bank and CIA data (with CIA data only used if there is not World Bank data)
## Ep: Proportion of workforce directly employed in tourism
## S_score: tourism sustainability score

tr_sust_gf <- tr_sust %>%
  mutate(gapfilled = ifelse(is.na(S_score) & !is.na(Ap), "gapfilled", NA)) %>%
  mutate(method = ifelse(is.na(S_score) & !is.na(Ap) & is.na(pcgdp2), "lm georegion + gdppcppp, with est. gdppcppp", NA)) %>%
  mutate(method = ifelse(is.na(S_score) & !is.na(Ap) & !is.na(pcgdp2), "lm georegion + gdppcppp", method)) %>%
  select(rgn_id, gapfilled, method, year)

write.csv(tr_sust_gf, here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability_gf.csv")), row.names=FALSE)

```

### Gapfilling

Linear models using gdppcppp and UN geopolitical regions as predictor variables. However if there is no gdppc data we estimate the gdppc using the UN georegions and then used in the linear model to gapfill the sustainability score.

```{r, eval=FALSE}

### Gapfill S using r1 and/or r2 regional data and PPP-adjusted per-capita GDP
### Looked at models with a year variable, but wasn't significant and decided to exclude

mod3 <- lm(S_score ~ as.factor(r2_label) + gdp_all, data=tr_sust, na.action = na.exclude)
summary(mod3)
anova(mod3)

mod4 <- lm(S_score ~ as.factor(r1_label) + gdp_all, data=tr_sust, na.action = na.exclude)
summary(mod4)
anova(mod4)

plot(predict(mod3), tr_sust$S_score)
abline(0,1)
plot(predict(mod4), tr_sust$S_score)
abline(0,1)


## Estimate missing data and gapfill
# Some of the r1 levels do not have data and consequently causes a fail. This chunk of code drops these levels so an NA is returned

# Select only r2 column
new_data <- tr_sust %>% 
  dplyr::select(r2_label, gdp_all)

unique(tr_sust$r2_label)

r2_w_data <- unique(tr_sust$r2_label[!is.na(tr_sust$S_score)])
  
new_data_r2 <- new_data %>%
  mutate(r2_label = ifelse(r2_label %in% r2_w_data, r2_label, NA))

# Predict sustainability scores using linear model 3 (using r2 data)
tr_sust <- tr_sust %>% 
  dplyr::mutate(S_score_pred_r2 = predict(mod3, newdata = new_data_r2))


# Select only r1 column
new_data <- tr_sust %>% 
  dplyr::select(r1_label, gdp_all)

unique(tr_sust$r1_label)

r1_w_data <- unique(tr_sust$r1_label[!is.na(tr_sust$S_score)])

new_data_r1 <- new_data %>%
  mutate(r1_label = ifelse(r1_label %in% r1_w_data, r1_label, NA))

# Predict sustainability scores using linear model 4 (using r1 data)
tr_sust <- tr_sust %>% 
  dplyr::mutate(S_score_pred_r1 = predict(mod4, newdata = new_data_r1))



## some are missing the r1 predictions, but none of these have Ep scores, so not relevant
View(filter(tr_sust, is.na(S_score_pred_r1)))

tr_sust <- tr_sust %>%
  mutate(S_score_2 = ifelse(is.na(S_score), S_score_pred_r2, S_score)) %>%
  mutate(S_score_2 = ifelse(is.na(S_score_2), S_score_pred_r1, S_score_2)) %>%
  filter(year %in% c(2019, 2021)) %>%
  select(rgn_id, year, S_score=S_score_2)

summary(tr_sust)

write_csv(tr_sust, here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability.csv")))
```

## Compare with previous year of data

```{r, eval=FALSE}
tr_sust <- read_csv(here(paste0("globalprep/tr/v", version_year, "/output/tr_sustainability.csv")))

prev_year <- (as.numeric(version_year) - 1) %>% 
  as.character()

compare <- tr_sust %>% 
  pivot_wider(names_from = year, values_from = S_score)

# current vs previous year of data
plot(compare$"2021", compare$"2019")
abline(0, 1, col="red")
# looks good

```

# Tw: Travel warnings

-   Travel warnings were deleted from the v2020 assessment.
