---
title: 'OHI `r format(Sys.Date(), "%Y")` - Tourism and Recreation '
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html' 
  pdf_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

[REFERENCE RMD FILE](http://ohi-science.org/ohiprep_v2024/globalprep/tr/v2024/tr_data_prep.html)

# Summary

This document describes the steps for obtaining the data used to calculate the tourism and recreation goal for the 2024 global assessment.

The general calculation is: tr = Ap \* Sr and Xtr = tr/90th quantile across regions

-   Ap = Proportion of overnight tourist arrivals to total arrivals
-   Sr = (S-1)/5; Sustainability of tourism

*The following data are used:*

-   Numbers of tourist arrivals, used in the calculation of the proportion of arrivals to area of coastline to population: obtained through the [UNWTO](https://www.unwto.org/tourism-statistics/key-tourism-statistics) (in the form of thousands of people). More info on tourism terms [here](https://www.unwto.org/glossary-tourism-terms). Range: 1995-2022 (only 2008-2022 is used)
-   Tourism sustainability: World Economic Forum. The Travel & Tourism Development Index 2024 (version May 2024). 2024. [TTDI](https://www3.weforum.org/docs/WEF_Travel_and_Tourism_Development_Index_2024.pdf)
-   Per capita GDP: (World Bank with gaps filled using CIA data), used to gapfill missing values in Tourism sustainability
-   World Bank (brought in from AO intermediate, \~/globalprep/ao/v2024/intermediate/gdppcppp_ohi.csv)
-   CIA (<https://www.cia.gov/the-world-factbook/field/real-gdp-per-capita/country-comparison/>)

**Ap = multiply arrivals by coastal population/total population, then divide by coastal area (sum of inland (1 km) and offshore (3 nm) area)**

-   Rescale using the 90th quantile as the reference point, or 80th (reevaluate when we get to this step)
-   Consider dropping low population areas, like they did in v2023
-   Calculate `Sr = (S-1)/5`; Sustainability of tourism
-   Finally, use the equation: `tr = Ap * Sr` for final scores before updating ohi-global

# Setup

```{r setup, message=FALSE, warning=FALSE, results="hide"}
# library(devtools)
# devtools::install_github("ohi-science/ohicore@dev") # dont worry about devtools
library(ohicore)
library(tidyverse)
library(stringr)
library(WDI)
library(here)
library(janitor)
library(plotly)
library(readxl)
library(naniar)
library(countrycode)
library(pdftools)
library(tictoc)

# ---- sources! ----
source(here("workflow", "R", "common.R")) # file creates objects to process data

region_data() # for rgns_all and rgns_eez

regions_shape() # returns spatial shape object named 'regions' which includes land, eez, highseas, and antarctica regions

#source(here(paste0("globalprep/tr/v", version_year, "/R/tr_fxns.R"))) # not used presently

# ---- set year and file path info ----
current_year <- 2024 # Update this in the future!!
version_year <- paste0("v",current_year)
data_dir_version_year <- paste0("d", current_year)
prev_ver_yr <- paste0("d", (current_year - 1))

# ---- data directories ----

# Raw data directory (on Mazu)
raw_data_dir <- here::here(dir_M, "git-annex", "globalprep", "_raw_data")

# UNWTO (UN World Tourism) raw data directory
unwto_dir <- here(raw_data_dir, "UNWTO", data_dir_version_year)

# intermediate data outputs dir
int_dir <- here("globalprep","tr", version_year, "intermediate")

# final output dir
output_dir <- here("globalprep","tr", version_year, "output")

# final figs dir
figs_dir <- here("globalprep","tr", version_year, "figs")
```

# Calculate Ap

*International data:* Click Inbound Arrivals, scroll down to "Total arrivals", click download data.

*Domestic data:* Click Domestic Tourism, scroll down to "Total trips", click download. Then download from "Accommodations".

## International UNWTO Inbound Tourism Arrivals
```{r}

```

## Domestic UNWTO Inbound Tourism Accomodation and Trips

Psuedocode:

- Bring in Accommodations
  - Filter to Total:Guests
    - see how many NAs
  - Quick lm across all other variables
  
- Bring in Trips
  - Filter to Overnights
  
- lm of Accommodations:Total:Guests to Trips:Overnights
  - plot them, see if 1:1.  Also plot it logged
  
- If the lm looks okay, and if they are correlated, use Trips:Overnights to fill in data for Accommodations:Total:Guests
  
- Next, see if there is a correlation between Accommodations:Total:Guests and Accommodations:Hotel:Guests
  - If it looks good (1:1 ish, also plot logged) then you can fill in Accommodations:Hotel:Guests data for Accommodations:Total:Guests
  
- Use name_2_rgn to convert UNWTO's "country" into "rgn_name" so that we can use georegions to approximate `mean_stay`
  - Ensure there are no duplicates, if there are then aggregate them accordingly
  
- If within the Accommodations data, there are countries which have Total:Guests and Total:Overnights, divide them to see how many nights guests stay on average for each country (call it `mean_stay`), use that to infer any gaps in Total:Guests data where there are only Overnights.
  - If possible, take lm(mean_stay ~ r2) or lm(mean_stay ~ r1) and find that mean_stay = ____*r1 to find a value for mean_stay for each georegion grouping. 
  - Or, after looking and seeing if they are all very similar values across georegions, take the mean and use that number to calculate guests using overnight values
  
- Ensure that the end product is all in the units of guests

### Clean Accommodations data

```{r}
# bring in domestic accommodation data from UNWTO
file_path_unwto_domestic_acc <- here::here(unwto_dir, "unwto-domestic-accommodation-data.xlsx")
unwto_arrivals_dom_acc <- readxl::read_xlsx(file_path_unwto_domestic_acc, skip = 4) # read in the raw domestic accomodations data

# clean up the raw xlsx! We want to have the country, the source (Total:Guests, Total:Overnights, Hotel:Guests, Hotel:Overnights), then the values
unwto_clean_dom <- unwto_arrivals_dom_acc %>% 
  select(country = `Basic data and indicators`, source_arrivals = `...6`, subdivision_1 = `...7`, subdivision_2 = `...8`, `1995`:`2021`) %>% # select relevant columns
  fill(country, .direction = "down") %>% # add country name to all data associated with that country
  fill(source_arrivals, .direction = "down") %>% 
  filter(subdivision_1 %in% "Overnights" | subdivision_1 %in% "Guests") %>% 
  select(-subdivision_2) %>% # get rid of the NA column
  replace_with_na_all(condition = ~.x == "..") %>% # swap ".." with NAs
  mutate(source_arrivals_all = paste(source_arrivals, subdivision_1, sep = ":")) %>% # unite the columns that define the source for easier cleaning and pivoting
  select(country, source_arrivals_all, 4:ncol(.)) # use the . to specify the whole dataframe

# tidy up the data to be usable
dom_acc_guests <- unwto_clean_dom %>% 
  pivot_longer(cols = 3:ncol(.), names_to = "year",
               values_to = "tourism_arrivals_ct") %>% # make the years not columns anymore
  pivot_wider(names_from = source_arrivals_all, values_from = tourism_arrivals_ct) %>%
  mutate(
    total_guests = as.numeric(`Total:Guests`),
    total_overnights = as.numeric(`Total:Overnights`),
    hotel_guests = as.numeric(`Hotels and similar establishments:Guests`), 
    hotel_overnights = as.numeric(`Hotels and similar establishments:Overnights`),
    tourism_arrivals_ct = as.numeric(NA)
    ) %>% # rename metrics so easier to work with, make numeric, and add a new column to fill with the new calculated values later
  select(country, year, total_guests, total_overnights, hotel_guests, hotel_overnights, tourism_arrivals_ct) %>% # select columns needed for future raw gapfilling 
  mutate(country = str_to_title(country)) # make countries look nice
```


```{r}
# see how many NAs are within the raw data for Total:Guests
summary(dom_acc_guests)
                                      
 #  total_guests      total_overnights
 # Min.   :     1.0   Min.   :     2  
 # 1st Qu.:   753.8   1st Qu.:  1532  
 # Median :  5226.0   Median : 13165  
 # Mean   : 19478.2   Mean   : 53824  
 # 3rd Qu.: 15603.5   3rd Qu.: 39173  
 # Max.   :371590.0   Max.   :576370  
 # NA's   :4947       NA's   :4614    
 #  hotel_guests      hotel_overnights  
 # Min.   :     0.2   Min.   :     0.2  
 # 1st Qu.:   160.0   1st Qu.:   412.2  
 # Median :  1326.0   Median :  2981.5  
 # Mean   : 11091.4   Mean   : 21297.9  
 # 3rd Qu.:  7737.2   3rd Qu.: 13999.5  
 # Max.   :342883.0   Max.   :440121.0  
 # NA's   :4129       NA's   :3835  
length(unique(dom_acc_guests$country)) # 223, which includes regions that have all NA values (UNWTO always has datasets with those 223 countries)

dom_acc_guests_noNA <- dom_acc_guests %>% 
  filter(!is.na(total_guests)) # remove NAs to see how many regions have total_guests values

length(unique(dom_acc_guests_noNA$country)) # 57 unique countries, not that many. let's move on to raw gapfilling and see if we can increase those numbers using other columns.

summary(lm(total_guests ~ as.numeric(year), data = dom_acc_guests)) # significant model and slope, but very low Multiple R-squared:  0.01165.  Not the best...
```

So, it looks like there are about 4947 NAs within the Total:Guests column.  However, we want to preserve guests/individuals as the unit (rather than overnights), so in the Raw Gapfilling section, let's see if we can gapfill with either other columns (hotel_guests) or the "Trips" data.

### Clean Trips data

```{r}
file_path_unwto_domestic_trips <- here::here(unwto_dir, "unwto-domestic-trips-data.xlsx")
unwto_arrivals_dom_trips <- readxl::read_xlsx(file_path_unwto_domestic_trips, skip = 4) # read in the raw domestic trips data

# clean up the Trips data!
unwto_dom_trips_clean <- unwto_arrivals_dom_trips %>% 
  select(country = `Basic data and indicators`, total_trips = `...6`, subdivision_1 = `...7`, subdivision_2 = `...8`, `1995`:`2021`) %>% # select relevant columns
  fill(country, .direction = "down") %>% # add country name to all data associated with that country
  pivot_longer(cols = c("total_trips", "subdivision_1", "subdivision_2"),
               values_to = "metric",
               values_drop_na = TRUE) %>% # make the metrics into one column
  select(-name) %>% # get rid of the name column since it's just the titles of the metrics which are already there
  select(country, metric, everything()) %>% # reorder things
  replace_with_na_all(condition = ~.x == "..") %>% # swap .. with NAs
  pivot_longer(cols = 3:ncol(.), names_to = "year",
               values_to = "tourism_arrivals_trips") %>% # make the years not columns anymore
  mutate(country = str_to_title(country)) %>% # make the country names look nicer
  pivot_wider(names_from = metric, values_from = tourism_arrivals_trips) %>%
  mutate(trips_overnights = as.numeric(`Overnights visitors (tourists)`), 
         trips_same_day = as.numeric(`Same-day visitors (excursionists)`), 
         trips_total = as.numeric(`Total trips`)) %>% # rename metrics so easier to work with, make numeric, and add a new column to fill with the new calculated values later
  select(country, year, trips_overnights, trips_same_day, trips_total) # select only the columns we will be working with
```

```{r}
# see how many NAs are within the raw data for Trips:Overnights
summary(unwto_dom_trips_clean)
                                      
 #   country              year           trips_overnights 
 # Length:6021        Length:6021        Min.   :      3  
 # Class :character   Class :character   1st Qu.:   2918  
 # Mode  :character   Mode  :character   Median :  14376  
 #                                       Mean   :  60263  
 #                                       3rd Qu.:  50884  
 #                                       Max.   :2321983  
 #                                       NA's   :5145     
 # trips_same_day       trips_total     
 # Min.   :      0.1   Min.   :      3  
 # 1st Qu.:   4799.2   1st Qu.:   8438  
 # Median :  21174.0   Median :  37058  
 # Mean   :  89189.9   Mean   : 237510  
 # 3rd Qu.:  98575.5   3rd Qu.: 201468  
 # Max.   :1834200.0   Max.   :6005852  
 # NA's   :5507        NA's   :5305

length(unique(unwto_dom_trips_clean$country)) # 223, which includes regions that have all NA values (UNWTO always has datasets with those 223 countries)

unwto_dom_trips_clean_noNA <- unwto_dom_trips_clean %>% 
  filter(!is.na(trips_overnights)) # remove NAs to see how many regions have total_guests values

length(unique(unwto_dom_trips_clean_noNA$country)) # 82 unique countries, not bad! Let's see if there is an association between trips_overnights and total_guests, to see if we can use the trips_overnights values for total_guests etc!
```


### Raw Gapfilling

#### Accommodations:Hotel:Guests used to fill in data for Accommodations:Total:Guests

```{r}
# ------- total_guests ~ hotel_guests ---------
mod1 <- lm(total_guests ~ hotel_guests, data = dom_acc_guests)
summary(mod1)
# significant model!
# total_guests = 3113.231504 + 1.143579*hotel_guests

# plot the data to compare
ggplot(dom_acc_guests, aes(x = hotel_guests, y = total_guests)) +
  geom_point(alpha = 0.5) +  # Add some transparency to points
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +  # Use a minimal theme
  labs(title = "Hotel:Guests vs Total:Guests within Accommodation",
       x = "Hotel Guests",
       y = "Total Guests") # does not look awful, the slope is slightly more than 1

# log scale
ggplot(dom_acc_guests, aes(x = log(hotel_guests + 1), y = log(total_guests + 1))) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(title = "Hotel:Guests vs Total:Guests (Log Scale) within Accommodation",
       x = "Log(Hotel Guests + 1)",
       y = "Log(Total Guests + 1)") # looks a lot better!

# let's look at their correlation. see ?`cor` if you are interested in learning more about how this is calculated (`cor` computes the variance of x and the covariance or correlation of x and y).  Kendall correlation is preferred when there are small samples or some outliers (https://datascience.stackexchange.com/questions/64260/pearson-vs-spearman-vs-kendall)
cor(dom_acc_guests$total_guests, dom_acc_guests$hotel_guests, method = "kendall") # NA

sd(dom_acc_guests$total_guests) # NA
sd(dom_acc_guests$hotel_guests) # NA

# lets try a different cor method to see if that is why we are getting NAs
cor(dom_acc_guests_cor$total_guests, dom_acc_guests_cor$hotel_guests, method = "pearson") # NA

sum(!is.finite(dom_acc_guests$total_guests)) # 0 
sum(!is.finite(dom_acc_guests$hotel_guests)) # 0, no infinite values, which is expected and good.  so what is the issue...

range(dom_acc_guests$total_guests, na.rm = TRUE) #  1 371590
range(dom_acc_guests$hotel_guests, na.rm = TRUE) # 0.18 342883.00
sum(is.na(dom_acc_guests$total_guests)) # 4947 NAs
sum(is.na(dom_acc_guests$hotel_guests)) # 4129 NAs
mean(!is.na(dom_acc_guests$total_guests)) # 0.1783757, aka 18% NA/total obs
mean(!is.na(dom_acc_guests$hotel_guests)) # 0.3142335, aka 32% NA/total obs

# significant amt of NAs, so we need to remove them before finding the correlation:
clean_dom_acc_guests <- dom_acc_guests[complete.cases(dom_acc_guests[c("total_guests", "hotel_guests")]), ] # filtering the dom_acc_guests data frame to keep only the rows where both "total_guests" and "hotel_guests" have non-missing values

# now lets try to find the correlation:
cor(clean_dom_acc_guests$total_guests, clean_dom_acc_guests$hotel_guests, method = "kendall") # 0.8984606, great!! This means that there is a pretty high correlation between the two variables (cor is close to 1).  Correlation measures the strength of association between two variables and the direction of the relationship: so in this case, it is positive and they have high association.
sd(clean_dom_acc_guests$total_guests) # 43242.82 guests
sd(clean_dom_acc_guests$hotel_guests) # 36884.91 guests, fairly similar to each other

# therefore, go ahead and do raw gapfilling in the next chunk using this data.
```

```{r}
# ------- gapfill using hotel_guests when total_guests = NA ---------
dom_acc_guests_gf1 <- dom_acc_guests %>% 
  group_by(country, year) %>% # group by county and year
  mutate(
    tourism_arrivals_ct = case_when(
      !is.na(total_guests) ~ total_guests, # if there is a value, dont gapfill
      is.na(total_guests) & !is.na(hotel_guests) ~ hotel_guests, # gapfill, when there is no data on total guests, fill with hotel guests
      TRUE ~ tourism_arrivals_ct # otherwise, NA
    )) %>% 
  mutate(arrivals_method = ifelse(is.na(total_guests) & !is.na(hotel_guests), "UNWTO - Hotel:Guests", NA)) %>%
  mutate(arrivals_gapfilled = ifelse(arrivals_method == "UNWTO - Hotel:Guests", "gapfilled", NA)) %>% # prepare a "gapfilled" column to indicate "gapfilled" or NA
  ungroup() # ungroup since not needed anymore

# take a look at how that changed the number of NAs:
summary(dom_acc_guests_gf1) # now, `tourism_arrivals_ct` has 4102 NAs, versus 4947 NAs in `total_guests`

dom_acc_guests_gf1_noNA <- dom_acc_guests_gf1 %>% 
  filter(!is.na(tourism_arrivals_ct)) # remove NAs to see how many regions have total_guests values

length(unique(dom_acc_guests_gf1_noNA$country)) # we now have 95 unique countries, that is so great!! Let's move on and see if we can gapfill further.
```

#### Trips:Overnights used to fill in data for Accommodations:Total:Guests

**Join Accomodations and Trips data frames first.**

```{r}
dom_acc_trips_join <- left_join(dom_acc_guests_gf1, unwto_dom_trips_clean, by = c("country","year")) # now it has all variables within one data frame
```

Now, evaluate if gapfilling is appropriate: 

```{r}
# ------- total_guests ~ trips_overnights ---------
mod2 <- lm(total_guests ~ trips_overnights, data = dom_acc_trips_join)
summary(mod2)
# significant model and slope! allllmost sig intercept, but that is okay
# total_guests = -2569.9 + 0.70848*trips_overnights

# plot the data to compare
ggplot(dom_acc_trips_join, aes(x = total_guests, y = trips_overnights)) +
  geom_point(alpha = 0.5) +  # Add some transparency to points
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +  # Use a minimal theme
  labs(title = "Accomodation:Total:Guests vs Trips:Overnight",
       x = "Trips Overnight Visitors (from Trips)",
       y = "Total Guests (from Accomodation)") # does not look that bad, the slope is slightly less than 1

# log scale
ggplot(dom_acc_trips_join, aes(x = log(total_guests + 1), y = log(trips_overnights + 1))) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(title = "Accomodation:Total:Guests vs Trips:Overnight (Log Scale)",
       x = "Log(Trips Overnight Visitors (from Trips) + 1)",
       y = "Log(Total Guests (from Accomodation) + 1)") # mediocre, but not awful.

# let's look at their correlation. see ?`cor` if you are interested in learning more about how this is calculated (`cor` computes the variance of x and the covariance or correlation of x and y).  Kendall correlation is preferred when there are small samples or some outliers (https://datascience.stackexchange.com/questions/64260/pearson-vs-spearman-vs-kendall)
range(dom_acc_trips_join$total_guests, na.rm = TRUE) #  1 371590
range(dom_acc_trips_join$trips_overnights, na.rm = TRUE) #   3 2321983, much larger range than total_guests, but it does include India and the US now
sum(is.na(dom_acc_trips_join$total_guests)) # 4947 NAs
sum(is.na(dom_acc_trips_join$trips_overnights)) # 5145 NAs
mean(!is.na(dom_acc_trips_join$total_guests)) # 0.1783757, aka 18% NA/total obs
mean(!is.na(dom_acc_trips_join$trips_overnights)) # 0.1454908, aka 15% NA/total obs

# significant amt of NAs, so we need to remove them before finding the correlation:
clean_dom_acc_trips_join <- dom_acc_trips_join[complete.cases(dom_acc_trips_join[c("total_guests", "trips_overnights")]), ] #filtering the dom_acc_guests data frame to keep only the rows where both "total_guests" and "hotel_guests" have non-missing values

# now lets try to find the correlation:
cor(clean_dom_acc_trips_join$total_guests, clean_dom_acc_trips_join$trips_overnights, method = "kendall") # 0.7645997, not bad! This means that there is a positive correlation between the two variables (cor is relatively close to 1).  Correlation measures the strength of association between two variables and the direction of the relationship: so in this case, it is positive and they have an association.
sd(clean_dom_acc_trips_join$total_guests) # 54458.93 guests
sd(clean_dom_acc_trips_join$trips_overnights) # 67727.62 guests, larger standard deviation than total_guests

# go ahead and do raw gapfilling in the next chunk using this data.
```

Let's gf!

```{r}
# -------- gf using trips_overnights for NA values of tourism_arrivals_ct -----
dom_acc_guests_gf2 <- dom_acc_trips_join %>% 
  group_by(country, year) %>% # group by county and year
  mutate(
    tourism_arrivals_ct = case_when(
      !is.na(tourism_arrivals_ct) ~ tourism_arrivals_ct, # if there is a value, dont gapfill
      is.na(tourism_arrivals_ct) & !is.na(trips_overnights) ~ trips_overnights, # gapfill, when there is no data on total guests or hotel guests, fill with trips_overnights visitors
      TRUE ~ tourism_arrivals_ct # otherwise, NA
    )) %>% 
  mutate(arrivals_method = ifelse(is.na(total_guests) & is.na(hotel_guests) & !is.na(hotel_guests), "UNWTO - Trips:Overnights", NA)) %>%
  mutate(arrivals_gapfilled = ifelse(arrivals_method == "UNWTO - Trips:Overnights", "gapfilled", NA)) %>% # fill the "gapfilled" column to indicate "gapfilled" or NA
  ungroup() # ungroup since not needed anymore

# take a look at how that changed the number of NAs:
summary(dom_acc_guests_gf2) # now, `tourism_arrivals_ct` has 3913 NAs, versus 4947 NAs in `total_guests`

dom_acc_guests_gf2_noNA <- dom_acc_guests_gf2 %>% 
  filter(!is.na(tourism_arrivals_ct)) # remove NAs to see how many regions have total_guests values

length(unique(dom_acc_guests_gf2_noNA$country)) # we now have 115 unique countries, that is so great!! Let's move on and see if we can gapfill even further!
```


#### Mean Stay as a multiplier to convert Total:Overnights into Total:Guests

So, there are some instances in which there is data for Total:Overnights, but not Total:Guests.  Based on looking at the methodology (https://www.e-unwto.org/doi/epdf/10.18111/9789284424160?role=tab), it seems that guests are the individuals coming to a specific place, versus overnights includes the same people over again if they stay for multiple nights. 

Essentially, Guests = Overnights / Mean Stay (in days).

So, we need to find mean stays! As a reminder, this would need to be calculated using the raw `total_guests` and `total_overnights` variables.  

**First, we have to convert to OHI Regions!!**

```{r dom_name_2_rgn}
# run name_2_rgn on the domestic data so it has OHI region ids
rgns_dom_acc_guests_gf2 <- name_2_rgn(df_in = dom_acc_guests_gf2, 
                                        fld_name = 'country',
                                        keep_fld_name = TRUE) %>% 
  select(rgn_id, rgn_name, country, year, total_guests, total_overnights, tourism_arrivals_ct, arrivals_method, arrivals_gapfilled) # lots of landlocked countries were removed, and there are duplicates.  Let's take a look.

# we only need mainland China data, not necessarily Hong Kong or Macao.
duplicates <- rgns_dom_acc_guests_gf2 %>%
  group_by(rgn_id, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates$rgn_name)
# [1] "China"                                              
# [2] "Guadeloupe and Martinique"                          
# [3] "Northern Mariana Islands and Guam"                  
# [4] "Montenegro"                                         
# [5] "Puerto Rico and Virgin Islands of the United States"

# see which ones even have data -- if none of the columns have a value, then we should drop these regions anyways.
duplicates_no_all_NAs <- duplicates %>%
  filter(if_any(c(total_guests, total_overnights, tourism_arrivals_ct), ~!is.na(.))) # only keeps the rows where at least one of the columns has a value

unique(duplicates_no_all_NAs$rgn_name) # only [1] "Montenegro" and [2] "Puerto Rico and Virgin Islands of the United States" need to be aggregated and fixed... the rest did not have data anyway.

# looking at the data, it looks like the country "Serbia and Montenegro" does not have data past 2001... so it is not usable anyway. So we can just drop that country but leave the country "Montenegro".  Then we can aggregate puerto rico and virgin islands `tourism_arrivals_ct` values per year.

# therefore, move forward by aggregating by region id and year for puerto rico and virgin islands.
## make a function to avoid losing NAs (they turn to 0s) when there are values to aggregate with
sum_with_na <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm = TRUE))
  }
}

# now use the function to aggregate by region and year!
dom_guests_agg <- rgns_dom_acc_guests_gf2 %>% 
  filter(!(rgn_id %in% c(209, 140, 14, 13))) %>% # removing duplicates that are all NAs
  filter(!(country %in% "Serbia And Montenegro")) %>% # removing a duplicate that does not have good data anyway
  group_by(rgn_id, rgn_name, year) %>% 
  summarize(
    tourism_arrivals_ct = sum_with_na(tourism_arrivals_ct), # put each column in the function, so no NAs are lost
    total_guests = first(total_guests, na_rm = TRUE),
    total_overnights = first(total_overnights, na_rm = TRUE),
    arrivals_method = first(arrivals_method, na_rm = TRUE),
    arrivals_gapfilled = first(arrivals_gapfilled, na_rm = TRUE),
    .groups = "drop" # to ungroup, fixes warning from summarise to use reframe (when it is not applicable here)
  ) #### wait why does it not say gapfilled for arrivals method for puerto rico?????? come back to this!!!  It loses arrivals_gapfilled and methods data

# double check to test
duplicates_2 <- dom_guests_agg %>%
  group_by(rgn_id, year) %>% 
  filter(n() > 1) %>%
  ungroup()
unique(duplicates_2$rgn_name) # v2024: 0! yay

# see which/how many regions have values
check_domestic <- dom_guests_agg %>% 
  filter(!is.na(tourism_arrivals_ct))
length(unique(check_domestic$rgn_id)) # v2024: 81 regions present
```

Next, **calculate `mean_stay` for each region**, if possible. Using an lm, see if `mean_stay` could be correlated to the georegion (r2 ideally because it is more specific, but r1 works as well).  If so, take the mean for that georegion, and use it as a way to convert Accommodation:Total:Overnights into Accommodation:Total:Guests.

```{r lm_mean_stay}
dom_mean_stay <- dom_guests_agg %>% 
  group_by(rgn_id, rgn_name, year) %>% 
  mutate(mean_stay = total_overnights/total_guests) # by referring to UNWTO documentation

range(dom_mean_stay$mean_stay, na.rm = TRUE) # 0.9666667 8.4101144
sd(dom_mean_stay$mean_stay, na.rm = TRUE) # 1.234238

georegions <- ohicore::georegions %>% 
  left_join(rgns_eez, by = "rgn_id") %>% 
  select(-c(Notes)) # not needed

# join mean_stay df with georegions! so we can see if there can be an association by georgn...
dom_mean_stay_georgns <- dom_mean_stay %>% 
  left_join(georegions, by = c("rgn_id","rgn_name"))

mod3 <- lm(mean_stay ~ r2, data = dom_mean_stay_georgns)
summary(mod3)
# not significant :(, so we have to move forward with the more general and less specific r1 groupings

mod4 <- lm(mean_stay ~ r1, data = dom_mean_stay_georgns)
summary(mod4)
# significant model!! HOWEVER, Multiple R-squared is 0.01365..... NOT GOOD.
# mean_stay = 3.2328514 + -0.0022696*r1
```


```{r lm_mean_stay}
# lets do a quick plot to compare mean stays and r1:
# plot the data to compare
ggplot(dom_mean_stay_georgns, aes(x = as.numeric(r1), y = mean_stay)) +
  geom_point(alpha = 0.5) +  # Add some transparency to points
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +  # Use a minimal theme
  labs(title = "Georegions vs Mean Stay",
       x = "Georegion (r1)",
       y = "Mean Stay (Days)") 

# log scale
ggplot(dom_mean_stay_georgns, aes(x = log(r1 + 1), y = log(mean_stay + 1))) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(title = "Georegions vs Mean Stay",
       x = "Log(Georegion (r1) + 1)",
       y = "Log(Mean Stay (Days) + 1)")

# let's look at their correlation. see ?`cor` if you are interested in learning more about how this is calculated (`cor` computes the variance of x and the covariance or correlation of x and y).  Kendall correlation is preferred when there are small samples or some outliers (https://datascience.stackexchange.com/questions/64260/pearson-vs-spearman-vs-kendall)
range(dom_mean_stay_georgns$r1, na.rm = TRUE) #   2 419, interesting to know
sum(is.na(dom_mean_stay_georgns$mean_stay)) # 3775 NAs
sum(is.na(dom_mean_stay_georgns$r1)) # 27 NAs: high seas regions
mean(!is.na(dom_mean_stay_georgns$mean_stay)) # 0.1726934, aka 17% NA/total obs
mean(!is.na(dom_mean_stay_georgns$r1)) # 0.9940828, aka 99% NA/total obs, great!

# significant amt of NAs, so we need to remove them before finding the correlation:
clean_dom_mean_stay_georgns <- dom_mean_stay_georgns %>% 
  filter(!is.na(mean_stay)) #filtering the dom_acc_guests data frame to keep only the rows where mean_stay has non-missing values

# now lets try to find the correlation:
cor(clean_dom_mean_stay_georgns$mean_stay, clean_dom_mean_stay_georgns$r1, method = "kendall") # 0.08158826, VERY LITTLE correlation... to be expected.
```

I would not move forward using georegions to group by and take the mean. Instead, I would either use the global mean to gapfill for mean_stays, or consider using a moving average.

What I will do for now is take the mean by region, assuming that there are some instances in which there may be some years that have overnights and guest values, but in other years it could have only overnights values. In that case, it would make the most sense to use the value found within the region if possible.

For regions that only have overnights data, we will use the global mean of mean_stays as our value to divide Overnights by.

**As a reminder:**

Mean Stay = Overnights/Guests (to see how many nights they stayed on average)

Therefore, if we have Overnights, we can find Guests by dividing by the Mean Stay. (Guests = Overnights / Mean Stay)

```{r mean_stay_multiplier}
dom_mean_stay_georgns_gf <- dom_mean_stay_georgns %>%
  group_by(rgn_id, rgn_name) %>% 
  mutate(
    mean_stay_filled = mean(mean_stay, na.rm = TRUE),
    mean_stay_filled = if_else(is.nan(mean_stay_filled), NA, mean_stay_filled)
    )

# global mean multiplier (technically, divider lol)
global_mean_stay <- mean(dom_mean_stay_georgns$mean_stay, na.rm = TRUE) # v2024: 2.893876 days

dom_mean_stay_georgns_filled <- dom_mean_stay_georgns_gf %>%
  mutate(mean_stay_filled = ifelse(is.na(mean_stay_filled), global_mean_stay, mean_stay_filled)) # if the mean_stay was already found regionally, leave it, otherwise replace with global mean_stay value
```

```{r}
# ------ use mean_stay to gapfill values --------
dom_acc_guests_gf3 <- dom_mean_stay_georgns_filled %>% 
  mutate(
    tourism_arrivals_ct = case_when(
      is.na(tourism_arrivals_ct) & !is.na(total_overnights) ~ total_overnights/mean_stay_filled,
      TRUE ~ tourism_arrivals_ct
    )
  ) %>%  # it worked!! Angola 1998 is a good example of where it may have worked.  There are others.
  select(rgn_id, rgn_name, year, tourism_arrivals_ct, arrivals_method, arrivals_gapfilled) ### LATER, CHECK ON GAPFILLED COLUMNS and add an if else here to show we gapfilled using this method

# take a look at how that changed the number of NAs:
summary(dom_acc_guests_gf3) # now, `tourism_arrivals_ct` has 2890 NAs, versus 4947 NAs in `total_guests`

dom_acc_guests_gf3_noNA <- dom_acc_guests_gf3 %>% 
  filter(!is.na(tourism_arrivals_ct)) # remove NAs to see how many regions have total_guests values

length(unique(dom_acc_guests_gf3_noNA$rgn_name)) # we now have 86 it added only 5 more regions --- it is still worth it!!
```









