---
title: 'OHI `r format(Sys.time(), "%Y")`: Natural Products: Ornamentals Data Prep'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    toc: true
    number_sections: false
    toc_float: yes
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../../workflow/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

[REFERENCE RMD FILE: <http://ohi-science.org/ohiprep_v2024/globalprep/np/v2024/STEP1a_np_ornamentals_prep.html>]

# Summary

This analysis calculates the score for ornamentals per region based on FAO commodities data. This calculation process includes creating a sustainability layer based off of risk and exposure for ornamental fishing.

# Updates from previous assessment

### v2024
*   Complete overhaul of methodology from previous years. Original data prep file: `STEP1a_np_ornamental_prep.Rmd`
*   New year of FAO data (2022)

------------------------------------------------------------------------

# Data Source

**Reference**:\
<https://www.fao.org/fishery/en/statistics/software/fishstatj> App release date: March 2024 FAO raw commodities quantity 1976_2021 FAO raw commodities value 1976_2022 [FAO metadata](http://www.fao.org/fishery/statistics/global-commodities-production/en)

- Global aquatic trade - All partners aggregated - Quantities and Values - 1976-2022 (Release date: July 2024)


**Downloaded**: 2024-08-07

**Description**: Quantity (tonnes) and value (USD) of raw commodities (Exports only) for each country, taxa, year. The FAO data is subset to include commodities in these categories: ornamental fish, fish oil, seaweed and plants (see: raw/commodities2products_weighting.csv for details).

**Time range**: 1976-2022

Files can be found on Mazu: `home/shares/ohi/git-annex/globalprep/_raw_data/FAO_commodities/d2024` accompanied by README.md with detailed download instructions. 

------------------------------------------------------------------------

# Methods



Although we do not know actual sustainable levels of harvest for ornamentals, $S_p$, in each region, we estimated it based on exposure and risk components:

$$S_c = 1 - average(E_c + R_c)$$

where $E_c$ is the exposure term and $R_c$ is the risk term for ornamentals.

The exposure term, $E_c$, is the ln-transformed intensity of harvest for ornamental fish calculated as tonnes of harvest per km2 of coral and rocky reef, relative to the global maximum. We ln transformed the harvest intensity scores because the distribution of values was highly skewed; because we do not know the true threshold of sustainable harvest, nearly all values would be considered highly sustainable without the log transformation. To estimate rocky reef extent area (km2) we used data from Halpern et al. (2008) ([Halpern et al. 2008](https://ohi-science.org/ohi-methods/goals/goal-models-data.html#ref-halpern2008global)), which assumes rocky reef habitat exists in all cells within 1 km of shore. Coral extent area (km2) are from UNEP-WCMC et al. (2018) ([UNEP-WCMC et al. 2018](https://ohi-science.org/ohi-methods/goals/goal-models-data.html#ref-unep-wcmc2018global)).



The risk term, $R_c$, is based on whether ornamental fishing has unsustainable harvest practices. In specific, we used the intensity of cyanide and dynamite fishing as a proxy. Risk for ornamental fish was set based on assessments of cyanide or dynamite fishing by [Reefs at Risk Revisited](www.wri.org/publication/reefs-at-risk-revisited) under the assumption that most ornamental fishes are harvested from coral reefs.

## Setup

```{r}

knitr::opts_chunk$set(eval = FALSE)

# ======= Load packages ============
if (!require(librarian)) {
  install.packages("librarian")
  library(librarian)
}
librarian::shelf(
  ohicore, #devtools::install_github('ohi-science/ohicore@dev') #if relevant: restart session after reinstalling
  dplyr,
  stringr,
  tidyr,
  here,
  tidyverse,
  zoo,
  ggplot2,
  plotly,
  tictoc,
  RColorBrewer
)

# ======= Set directories ===========
# Update scenario year, set up programmatic scenario year updates
scen_year_number <- 2024
scen_year <- as.character(scen_year_number)
prev_scen_year <- as.character(scen_year_number - 1)
data_dir_year <- paste0("d", scen_year)
prev_data_dir_year <- paste0("d", prev_scen_year)
v_scen_year <- paste0("v", scen_year)

current_np_dir <- here::here("globalprep", "np", paste0(v_scen_year, "_new"))

# ======== Source functions ===========
source(here::here("workflow", "R", "common.R")) # directory locations
# Load FAO-specific user-defined functions
source(here::here("workflow", "R", "fao_fxn.R")) # function for cleaning FAO files
# function for handling FAO commodity data specific to NP
source(here::here(current_np_dir, "R", "np_fxn.R")) 
```



STEP 2: Ornamentals
1.    Using data A, filter the tonnes and USD for ornamentals.  
2.    Calculate exposure and risk
3.    then use these to estimate sustainability: mutate(sustainability = 1 - mean(c(exposure, risk), na.rm = TRUE))
4.    Calculate sustainable tonnes: tonnes * sustainability
5.    Rolling average of sustainable tonnes
6.    Identify max value from rolling average of sustainable tonnes (maybe limit to most recent 10 years of data to ID max). 
7.    get relative tonnes by dividing the rolling average of sustainable tonnes by max(rolling average of sustainable tonnes)


So you can then save an intermediate dataset that looks like this:

`rgn_id`\
`year`,\
`product`, (ornamentals, in this case)\
`tonnes`: from gapfilled FAO commodity data\
`usd`: from gapfilled FAO commodity data\
`sustainability`: `1 - mean(c(exposure, risk), na.rm = TRUE)`\
`sustainable_tonnes`: `tonnes` $*$ `sustainability`\
`max_sustainable_tonnes`: `max(sustainable_tonnes)`\
`rel_sust_tonnes`: `tonnes_rollmean` / `peak_sust_tonnes`

Then, in final data, save `rgn_id`, `product`, `year`, `rel_sust_tonnes` 


```{r}
# read in FAO commodities usd tonnes int data
np_harvest_usd_tonnes <- read_csv(here(current_np_dir, "int", "np_harvest_usd_tonnes_gf.csv"))

# filter FAO commodities int data to ornamentals
np_ornamentals_base <- np_harvest_usd_tonnes %>% 
  dplyr::filter(product %in% c("ornamentals"))

# check
table(np_ornamentals_base$product)
# 3938 -- v2024

```



## Calculate Exposure

Calculate ornamentals exposure based on habitats


Returns the first input data frame with a new column for exposure:
[`rgn_id` `year` `product` `exposure`]

```{r}
    
# ---- Subset np harvest ornamentals ----
np_harvest_orn <- np_ornamentals_base %>%
  dplyr::select(rgn_id, year, product, tonnes) %>% 
  arrange(rgn_id)

# ====== Determine Habitat Areas for Exposure ========
# rocky reef habitat area
hab_rocky <- readr::read_csv(here::here("globalprep", "hab_rockyreef", "v2012", # hasn't been updated since 2012
                           "data", "habitat_extent_rocky_reef_updated.csv")) %>%
  dplyr::select(rgn_id, km2) %>%
  dplyr::filter(km2 > 0)

# coral habitat area
hab_coral <- readr::read_csv(here::here("globalprep", "hab_coral", "v2021", # hasn't been updated since 2021
                           "data", "habitat_extent_coral_updated.csv")) %>%
  dplyr::select(rgn_id, km2) %>%
  dplyr::filter(km2 > 0)

# find total area in coral and rocky reef habitats
area_dual_hab <- np_harvest_orn %>%
  dplyr::left_join(hab_coral %>%
                     dplyr::rename(coral_km2 = km2),
                   by = c('rgn_id')) %>%
  dplyr::left_join(hab_rocky %>%
                     dplyr::rename(rocky_km2 = km2),
                   by = c('rgn_id')) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(km2 = sum(c(rocky_km2, coral_km2), na.rm = TRUE)) %>%
  dplyr::filter(km2 > 0) %>%
  dplyr::select(-coral_km2, -rocky_km2) %>% 
  dplyr::arrange(rgn_id)
  

# ======== Determine Exposure ========
### exposure: get tonnes / area, find max in past 10 years, use max to rescale with log transform

# calculate raw exposure (tonnes / area)
np_exp_raw <- area_dual_hab %>%
  # find regions and years with harvest in rocky reef and coral habitat
  dplyr::mutate(expos_raw = ifelse(tonnes > 0 & km2 > 0,
                                   (tonnes / km2), # raw exposure: harvest tonnes / habitat area
                                   0)) %>% # 0 exposure if no harvest or no special habitat
  dplyr::group_by(product) 
 
# define most recent 10 year range 
max_exp_year <- max(np_exp_raw$year)
exp_decade_range <- seq((max_exp_year - 10), max_exp_year, 1)

## find max raw exposure value in the past 10 years
np_exp_max <- np_exp_raw %>%
  filter(year %in% c(exp_decade_range)) %>% 
  group_by(rgn_id) %>% # not sure why they didn't group by rgn_id before... 
  dplyr::mutate(expos_prod_max = max(expos_raw, na.rm = TRUE)) %>%
#  dplyr::ungroup() %>% 
  dplyr::select(rgn_id, expos_prod_max) %>% 
  # keep 1 row with max value per region (values are the same across year range per region)
  dplyr::distinct()


# find 95th quantile of the past 10 years ----

# subset of past 10 years
np_exp_recent_decade <- np_exp_raw %>% 
  dplyr::filter(year %in% c(exp_decade_range))

global_95th_quantile <- stats::quantile(np_exp_recent_decade$expos_raw, probs = c(0.95))
# v2024: 95th is 0.1511856; 99th is 45.90351 


# completely new method: exposure = raw exposure / global 95th quantile
# calculate rescaled exposure ----
np_exposure_new <- np_exp_raw %>% 
  dplyr::mutate(global_95th = global_95th_quantile) %>% 
  dplyr::mutate(exposure = expos_raw / global_95th)

 
# log-transform: ----
np_exp_log <- np_exp_raw %>% 
  dplyr::mutate(log_exp = log(expos_raw))

log_exp_recent_decade <- np_exp_log %>% 
  dplyr::filter(year %in% c(exp_decade_range))

log_global_95th_exp <- stats::quantile(log_exp_recent_decade$log_exp, probs = c(0.95))
# -1.889

np_log_exposure <- np_exp_log %>% 
  dplyr::mutate(log_global_95th = log_global_95th_exp) %>% 
  dplyr::mutate(log_exposure = log_exp / log_global_95th)


np_log_exposure_2 <- np_exp_log %>% 
  dplyr::mutate(log_global_95th = log_global_95th_exp) %>% 
  dplyr::mutate(log_exposure = (log_exp + 1) / (log_global_95th + 1))



## calculate rescaled exposure [old]
np_exp_orn_rescale <- np_exp_raw %>%
  # bring in max (of the past 10 yrs) value, 
  # fill single value for each region across full time range
  dplyr::left_join(np_exp_max, by = "rgn_id") %>% 
  # rescale exposure using max value
  dplyr::mutate(exposure = (log(expos_raw + 1) / log(expos_prod_max + 1)),
                exposure = ifelse(exposure > 1, 1, exposure)) %>%
  dplyr::select(-km2, -expos_raw, -expos_prod_max)


# ======== Save intermediate data products ==========
# clean up columns
np_exp_gf <- np_exp_orn_rescale %>%
  dplyr::mutate(gapfilled = ifelse(is.na(exposure), 1, 0)) %>%
  dplyr::mutate(method = ifelse(is.na(exposure), "prod_average", NA)) %>%
  dplyr::select(rgn_id = rgn_id, product, year, gapfilled, method)

# write out to int folder (directory)
readr::write_csv(np_exp_gf, here(current_np_dir, 
                                 "int", # used to be output, moved to int
                                 # because exposure is no longer an output
                                 "np_exposure_ornamentals_gf_record.csv"))

# add exposure for countries with (habitat extent == NA)
np_exp_orn <- np_exp_orn_rescale %>%
  dplyr::group_by(product) %>% # not sure why they didn't have some georegion gapfill instead of gapfilling with mean global exposure??
  dplyr::mutate(mean_exp = mean(exposure, na.rm = TRUE)) %>%
  dplyr::mutate(exposure = ifelse(is.na(exposure), mean_exp, exposure)) %>%
  dplyr::select(-mean_exp) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(product = as.character(product)) %>%
  dplyr::select(rgn_id, year, product, exposure)

# save ornamentals exposure to int 
readr::write_csv(np_exp_orn, here(current_np_dir,
                                  "int", # used to be output, moved to int
                                 # because exposure is no longer an output
                                  "np_exposure_ornamentals.csv"))

# potentially filter to past 20 years (or hardcode from 2004 onward)

```



```{r}
ggplot(data = np_exp_orn %>% filter(rgn_id == 163), aes(x = year, y = exposure)) + 
  geom_line() +
  theme_bw()
```




```{r}
np_exp_raw

ggplot(data = np_exp_raw %>% filter(rgn_id == 163), aes(x = year, y = expos_raw)) + 
  geom_line() +
  theme_bw()


```


global-scaled exposure test:


```{r}
np_global_max <- np_exp_raw %>%
  filter(year %in% c(exp_decade_range)) %>% 
 # group_by(rgn_id) %>% # not sure why they didn't group by rgn_id before... 
  dplyr::mutate(expos_prod_max = max(expos_raw, na.rm = TRUE)) %>% # 143.385 for all...
  dplyr::ungroup() %>% # without this, "ornamentals" column remains
  dplyr::select(rgn_id, expos_prod_max) %>% 
  # keep 1 row with max value per region (values are the same across year range per region)
  dplyr::distinct()

# calculate rescaled exposure
global_exp_rescale <- np_exp_raw %>%
  # bring in max (of the past 10 yrs) value, 
  # fill single value for each region across full time range
  dplyr::left_join(np_global_max, by = "rgn_id") %>% 
  # rescale exposure using max value
  dplyr::mutate(exposure = (log(expos_raw + 1) / log(expos_prod_max + 1)),
                exposure = ifelse(exposure > 1, 1, exposure)) %>%
  dplyr::select(-km2, -expos_raw, -expos_prod_max)


np_exp_orn_global <- global_exp_rescale %>%
  dplyr::group_by(product) %>% # not sure why they didn't have some georegion gapfill instead of gapfilling with mean global exposure??
  dplyr::mutate(mean_exp = mean(exposure, na.rm = TRUE)) %>%
  dplyr::mutate(exposure = ifelse(is.na(exposure), mean_exp, exposure)) %>%
  dplyr::select(-mean_exp) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(product = as.character(product)) %>%
  dplyr::select(rgn_id, year, product, exposure)


ggplot(data = np_exp_orn_global %>% filter(rgn_id == 163), aes(x = year, y = exposure)) + 
  geom_line() +
  theme_bw()


# exposure scale is much smaller now, as values are being compared to global max exposure, not regional...
```



## Calculate Risk

```{r}

### calculates NP risk based on:
###   ornamentals:      risk = 1 if blast or cyanide fishing
### Returns a data frame of risk, by product, by region:
###
#########################################.


### Determine Risk

r_cyanide <- 
  read_csv(here("globalprep", "np_prs_poison_blast_fishing", "v2013", "data", "gl_thr_poison_3nm_rgn2013.csv")) %>%
  #AlignDataYears(layer_nm = "np_cyanide", layers_obj = layers) %>%
  dplyr::filter(!is.na(score) & score > 0) %>%
  dplyr::select(rgn_id,
                #year = scenario_year,
                cyanide = score)

r_blast <-
  read_csv(here("globalprep", "np_prs_poison_blast_fishing", "v2013", "data", "gl_thr_blast_3nm_rgn2013.csv")) %>%
  #AlignDataYears(layer_nm = "np_blast", layers_obj = layers)  %>%
  filter(!is.na(score) & score > 0) %>%
  dplyr::select(rgn_id,
         #year = scenario_year,
         blast = score)


### risk for ornamentals set to 1 if blast or cyanide fishing present, based on Nature 2012 code
###  despite Nature 2012 Suppl saying Risk for ornamental fish is set to the "relative intensity of cyanide fishing"
risk_orn <- r_cyanide %>%
  full_join(r_blast, by = c("rgn_id")) %>%
  mutate(ornamentals = 1) %>%
  dplyr::select(rgn_id, ornamentals)

### risk as binary
np_risk_orn <-
  expand.grid(
    rgn_id  = unique(np_harvest_orn$rgn_id),
    year = unique(np_harvest_orn$year)
  ) %>%
  ### ornamentals
  left_join(risk_orn, by = c("rgn_id"))  %>%
  mutate(ornamentals = ifelse(is.na(ornamentals), 0, ornamentals)) %>%
  gather(product, risk, -rgn_id, -year) %>%
  mutate(product = as.character(product))

np_risk_gf <- np_risk_orn %>%
  mutate(gapfilled = 0, method = NA) %>%
  dplyr::select(-risk)

readr::write_csv(np_risk_gf, here(current_np_dir, "output", "np_risk_ornamentals_gf.csv"))

## write this as a risk layer 
readr::write_csv(np_risk_orn, here(current_np_dir, "output", "np_risk_ornamentals.csv"))
```


## Calculate Sustainability

1 - mean(c(exposure, risk), na.rm = TRUE)


## Calculate Sustainable Harvest

tonnes * sustainability



## Rolling Average of Sustainable Tonnes



## Identify Max Sustainable Tonnes

max(sustainable_tonnes)


## Relative Sustainable Tonnes


tonnes_rollmean / peak_sust_tonnes




