---
title: "CRS Prep & Review"
output: html_document
date: "2024-08-05"
---

```{r setup, include=FALSE}

library(terra)
library(here)

```

Data describing commercial landings (tonnes) in 2017.
```{r}
fish <- rast(here("Reference", "CRS", "commercial_landings_2017.tif"))

## Take a look
plot(fish)
# lets log so we can visualize this a little better
plot(log(fish+1))
```

How many fish were captured?
```{r}
global(fish, "sum", na.rm = TRUE)
# 86.3 million
```
Check against Sea Around Us to make sure this seems generally reasonable.

-   seems quite low <https://www.seaaroundus.org/data/#/global?chart=catch-chart&dimension=sector&measure=tonnage&limit=10>


What is the area of each cell?
```{r}
cell_area <- cellSize(fish)
plot(cell_area, box = TRUE)
```
Describe what the plotted image is saying?

Given this, what is kind of deceptive about these plots?
```{r}
plot(fish)
# lets log so we can visualize this a little better
plot(log(fish+1))

```

-   the cells in the northern and southern extremes have a much gerater area than those concentrated in the center
-   this may hide higher commercial fishing values in the center band, as those pixels will appear smaller when plotted
-   more coarse resolution at the poles can lead to higher catch shown over a greater area than in reality
-   when we sum values across cells, the area of that cell is relevant when we do a global summation:
    -   From the documentation on `cellSize()`
    
> Compute the area covered by individual raster cells.
> Computing the surface area of raster cells is especially relevant for longitude/latitude rasters.

> But note that for both angular (longitude/latitude) and for planar (projected) coordinate reference systems raster cells sizes are generally not constant, unless you are using an equal-area coordinate reference system.

> For planar CRSs, the area is therefore not computed based on the linear units of the coordinate reference system, but on the *actual* area by transforming cells to longitude/latitude. If you do not want that correction, you can use transform=FALSE or init(x, prod(res(x)))


Let's convert these data to another coordinate reference system:
```{r}
moll <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
fish_moll <- terra::project(fish, crs(moll))

# take a look
par(mar = c(5, 4, 4, 0)) # set margins: bottom, left, top, right
plot(fish_moll)
plot(log(fish_moll + 1), main = "Transformed CRS: Mollweide")
plot(log(fish+1), main = "Original CRS: WGS 84 EPSG 4326")

## Count the total tonnes
global(fish_moll, "sum", na.rm = TRUE)
# 389.3 million

# calculate cell area
cell_area_moll <- cellSize(fish_moll)

#par(mai = c(1, 1, 3, 5)) # set margins: bottom, left, top, right
#par(oma = c(1, 1, 3, 4))
#par(mar = c(1,1,1,1))
plot(cell_area_moll, main = "Mollweide projection cell size")

plot(cell_area, main = "Original projection cell size")
```

```{r}
(cell_area_moll$area)
# min 0
# max is 547 million
cell_area$area
# min: 122442386 (122 million)
# max: 3077249667 (3 billion)
```


Compare to the original tonnes? What is happening? How can you improve this?
-   original tons: 86331119
-   Mollweide tons: 389271065
-   the cell size area is now relatively consistent across the globe -- old cell sizes have stretched etc., to fit a standardized cell size (I tried to extend the margins to see the full scalebar label values but struggled)
-     perform global calculations before reprojecting

```{r}
View(fish)

summary(fish)

fish_polygons <- terra::as.polygons(fish)

fish_df <- terra::values(fish_polygons)
```

```{r}
# Get the cell numbers
cells <- 1:terra::ncell(fish)

# Get the x and y coordinates for each cell
xy <- terra::xyFromCell(fish, cells)

# Extract the values
values <- terra::values(fish)

# Combine everything into a data frame
fish_df <- data.frame(
  x = xy[,1],
  y = xy[,2],
  value = values
)

fish_no_na <- fish_df %>% 
  na.omit()
```

-   

See what happens when we make multiple CRS conversions:

```{r}
# Robinson projection
rob <- "+proj=robin +datum=WGS84 +units=m +no_defs"

fish_moll_rob <- terra::project(fish_moll, crs(rob))
plot(log(fish_moll_rob + 1))

# now back to lat long:
fish_moll_rob_latlon <- project(fish_moll_rob, fish)

plot(log(fish_moll_rob_latlon + 1))

# check to see if it exactly matches the start
check <- fish_moll_rob_latlon - fish
plot(log(check + 1))
check


```

What is this telling us?


-   the values have changed notably between projections