---
title: "CRS Prep & Review"
date: "2024-08-01"
#date: "`r format(Sys.Date(), '%Y')`"
#author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
author: "Anna Ramji"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '../../workflow/templates/ohi_hdr.html'
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
# ---- Load packages ----
library(terra)
library(here)
library(tidyverse)
```


Data describing commercial landings (tonnes) in 2017.

```{r}
# Read in commercial landings data
fish <- terra::rast(here("Reference", "CRS", "commercial_landings_2017.tif"))

# print out geospatial data info (class, dimensions, resolution, extent, CRS, etc.)
fish # equivalent to (fish)

# print out summary of value layer
summary(fish)

## Take a look
terra::plot(fish)
terra::plot(fish, mar = c(0, 1.5, 0, 5))


# lets log so we can visualize this a little better
terra::plot(log(fish+1), mar=c(0,1.5,0,3.5))
```


Trying to save image

```{r}
# fish_log_plot <- terra::plot(log(fish+1))
# 
# fish_log <- log(fish$commercial_landings_2017 + 1)

# png('file.png', height=nrow(r), width=ncol(r)) 
# plot(r, maxpixels=ncell(r))
# dev.off()
#terra::writeRaster(fish_log, here::here("Reference/CRS/figs/fish_log_plot.png"), overwrite=TRUE)
```


```{r}
# fish_df <- as.data.frame(fish)
# fish_log_plot <- ggplot2::ggplot() +
#   geom_raster(data = log(fish_df + 1), aes(x = commercial_landings_2017))
# fish_log_plot
```


```{r}
# # Set the output file name and path
# output_file <- here::here("Reference", "CRS", "figs", "commercial_landings_2017_log.png")
# 
# # Open a PNG device
# png(filename = output_file, 
#     width = 800, height = 600, 
#     units = "px", res = 100)
# 
# # Create your plot
# plot(log(fish+1), 
#      main="Log-transformed Commercial Landings (2017)", 
#      col=hcl.colors(100, "viridis"),
#      legend.args=list(text="Log(landings + 1)", side=4, line=2.5, cex=0.8))
# 
# # Close the PNG device
# dev.off()
```


```{r}
# library(terra)
# 
# # Set the output file name and path
# output_file <- here::here("Reference", "CRS", "figs", "commercial_landings_2017_log.png")
# 
# # Ensure the directory exists
# dir.create((output_file), showWarnings = TRUE, recursive = TRUE)
# 
# # Open a PNG device
# png(filename = output_file, width = 800, height = 600, units = "px", res = 100)
# 
# # Create your plot
# plot(log(fish+1), 
#      main="Log-transformed Commercial Landings (2017)", 
#      col=hcl.colors(100, "viridis"),
#      legend.args=list(text="Log(landings + 1)", side=4, line=2.5, cex=0.8))
# 
# # Close the PNG device
# dev.off()
```


How many fish were captured?

```{r}
global(fish, "sum", na.rm = TRUE)
# 86.3 million tonnes
```
Check against Sea Around Us to make sure this seems generally reasonable.

-   seems a little low but generally makes sense (close enough) -- [Sea Around Us](https://www.seaaroundus.org/data/#/global?chart=catch-chart&dimension=sector&measure=tonnage&limit=10) says 87 million 
-   around 100 million <https://www.seaaroundus.org/data/#/global?chart=catch-chart&dimension=sector&measure=tonnage&limit=10>


What is the area of each cell?

```{r}
cell_area <- terra::cellSize(fish)
plot(cell_area, box = TRUE, mar = c(1.5, 0, 0, 3))

print(cell_area)
```

Describe what the plotted image is saying?

- cell size is much smaller at the poles, increases as you head in towards the equator

Given this, what is kind of deceptive about these plots?

-   if density of fish landings in tonnes, pole values would appear higher than they are -- apparent tonnes per cell would be higher 
-   if count, same thing (seems higher than in reality)

```{r}
plot(fish)
# lets log so we can visualize this a little better
plot(log(fish+1))

```

-   the cells in the northern and southern extremes have a much smaller area than those concentrated in the center
-   this may hide higher commercial fishing values in the center band, as those pixels will appear smaller when plotted
-   value associated w the cell is changing when you convert -- resampling by diff method (nearest neighbor, linear interpolation, bilinear)
-   poles may appear to have higher values than they actually do -- same value over smaller area = higher density value
-   more coarse resolution at the poles can lead to higher catch shown over a greater area than in reality
-   when we sum values across cells, the area of that cell is relevant when we do a global summation:
    -   From the documentation on `cellSize()`
    
> Compute the area covered by individual raster cells.
> Computing the surface area of raster cells is especially relevant for longitude/latitude rasters.

> But note that for both angular (longitude/latitude) and for planar (projected) coordinate reference systems raster cells sizes are generally not constant, unless you are using an equal-area coordinate reference system.

> For planar CRSs, the area is therefore not computed based on the linear units of the coordinate reference system, but on the *actual* area by transforming cells to longitude/latitude. If you do not want that correction, you can use transform=FALSE or init(x, prod(res(x)))


Let's convert these data to another coordinate reference system:
```{r}
moll <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
fish_moll <- terra::project(fish, crs(moll))

# take a look
par(mar = c(5, 4, 4, 0)) # set margins: bottom, left, top, right
plot(fish_moll)
plot(log(fish_moll + 1), main = "Transformed CRS: Mollweide")
plot(log(fish+1), main = "Original CRS: WGS 84 EPSG 4326")

## Count the total tonnes
global(fish_moll, "sum", na.rm = TRUE)
# 389.3 million

# calculate cell area
cell_area_moll <- cellSize(fish_moll)

#par(mai = c(1, 1, 3, 5)) # set margins: bottom, left, top, right
#par(oma = c(1, 1, 3, 4))
#par(mar = c(1,1,1,1))
plot(cell_area_moll, main = "Mollweide projection cell size")

plot(cell_area, main = "Original projection cell size")
```

-   389 million tonnes is crazy

```{r}
(cell_area_moll$area)
# min 0
# max is 547 million
cell_area$area
# min: 122442386 (122 million)
# max: 3077249667 (3 billion)
```


Compare to the original tonnes? What is happening? How can you improve this?
-   original tons: 86331119
-   Mollweide tons: 389271065
-   the cell size area is now relatively consistent across the globe -- old cell sizes have stretched etc., to fit a standardized cell size (I tried to extend the margins to see the full scalebar label values but struggled)
-     perform global calculations before reprojecting



-   could take raster of cell size, multiply by original (if density) raster, then sum values 
-   ^ divide if count

-   tonnes / area = density, then project
-   take new cell size, multiply by density raster
-   sum at end to check again -- might need to reduce resolution, summary before projecting data


```{r}
# get density (tonnes/area)
fish_density <- fish / cell_area

plot(log(fish_density))

# reproject to Mollweide
fish_density_moll <- terra::project(fish_density, crs(moll))
fish_density_moll

plot(log(fish_density_moll))

# get area of cells
new_cell_size <- cellSize(fish_density_moll)
plot(new_cell_size)

# multiply density by area to get count (tonnes)
new_tonnes <- fish_density_moll * new_cell_size
plot(log(new_tonnes + 1))

# sum to check
global(new_tonnes, "sum", na.rm = TRUE)
# 85 million!
```

-   close enough :)
-   hopefully you don't care about Fiji data too much (gets distorted with Mollweide projection)


```{r}
#View(fish)

summary(fish)

fish_polygons <- terra::as.polygons(fish)

fish_df <- terra::values(fish_polygons)
```

```{r}
# Get the cell numbers
cells <- 1:terra::ncell(fish)

# Get the x and y coordinates for each cell
xy <- terra::xyFromCell(fish, cells)

# Extract the values
values <- terra::values(fish)

# Combine everything into a data frame
fish_df <- data.frame(
  x = xy[,1],
  y = xy[,2],
  value = values
)

fish_no_na <- fish_df %>% 
  na.omit()
```

-   

See what happens when we make multiple CRS conversions:

```{r}
# Robinson projection
rob <- "+proj=robin +datum=WGS84 +units=m +no_defs"

fish_moll_rob <- terra::project(fish_moll, crs(rob))
plot(log(fish_moll_rob + 1))

# now back to lat long:
fish_moll_rob_latlon <- project(fish_moll_rob, fish)

plot(log(fish_moll_rob_latlon + 1))
plot(log(fish + 1))

# check to see if it exactly matches the start
check <- fish_moll_rob_latlon - fish
plot(log(check + 1))
check


```

What is this telling us?


-   the values have changed notably between projections

-   some are highly inflated (large positive numbers), other deflated (large negative in new projection)

-   adding error when we flip between resolutions

-   they're all just approximations (by nature)


```{r}
fish_moll_rob

fish_moll_rob_latlon <- project(fish_moll_rob, fish) # match template (convert crs and resolution)
fish_moll_rob_latlon

fish_moll_rob_latlon_v2 <- project(fish_moll_rob, crs(fish)) # match crs (just convert crs)
fish_moll_rob_latlon_v2
```

